### 2020-02-06 paper/src/paper.tex
Background chapter - V.1

\documentclass[11pt,twoside,a4paper]{report}

\usepackage[ruled, vlined, linesnumbered]{algorithm2e}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{changepage}
\usepackage{fancyhdr} % Header

\pagenumbering{arabic}

\begin{document}

\renewcommand{\headrulewidth}{0pt}
\newcommand{\ie}{\textit{i.e.} }
\newcommand{\nats}{\mathbb{N} }
\newcommand{\horizontalline}{\newline\vspace{.3cm}\hfill\makebox[.5\linewidth]{\rule{.5\textwidth}{0.4pt}}\hfill\vspace{.1cm}}
% title
\title{Implementing \& Evaluating Space Efficient Algoirthms for Detecting Large Neighbourhoods in Graph Streams}
\author{Dom Hutchinson}
\date{\today}
\maketitle

% Header
\pagestyle{fancy}
\fancyhead[L]{Dom Hutchinson (1701111)}
\fancyhead[C]{}
\fancyhead[R]{\today}

\chapter{Background}

\section{Motivation}
\par For most algorithms time efficiency is viewed with primary importance, rather than space efficiency. However if the memory used for an algorithm exceeds the size of the computer's RAM then virtual memory would have be used. Virtual memory has a much longer read/write than RAM meaning that the algorithm's run-time will be much greater in this scenario. Thus for algorithms which are designed to work with large data sets space efficiency is more important.
\par In the early 2000s companies began to realise that data could be used as a commodity, and with this the amount of data being collected has surged in the years since. In this same time period we have nots seen the same surge in the amount of RAM typical computers have. Thus people wishing to capitalise on this increase in available data need to give greather thought to the space efficiency of the algorithms they are using. Otherwise run-times could become unusuable.
\par Social networks are good sources of data about people. This data can leveraged in many ways, notably within advertising. Behaviourly targeted adverts have been shown to have more than twice the click through rate of standard online advertiesments \cite{targetedAds}. The last decade has seen developments in the space of targeted advertising, one of which is the rise of social media influencers. These are people who are deemed to have a large online following, which they are able to influence. Advertisers will often approach these influencers with deals by which the influencer will promote a product to their audience, hopefully driving sales.
\par The question becomes, ``Who is an influencer \& how do we find them?''. This question can be generalised as the $\mathtt{Neighborhood\ Detection\ Problem}$. Which is the problem I shall discuss in this paper.
%TODO
Although in my description of the Neighbourhood Detection Problem I validate its relevance by referring to find social media influencers, it is applicable to many other scenarios.
\begin{itemize}
	\item[-] TODO
\end{itemize}
\section{Neighbourhood Detection Problem}
%TODO Describe Graph streams, insertion-only & insertion-deletion
\par Graph streams are a dynamic way of representing graphs. They are simply a list of instructions for how to construct a graph and in this paper I shall discuss two types: insertion-only streams \& insertion-deletion streams. There are other types which allow for updating vertex values, among other properties, but they are not relevant to the problem we are considering in this paper.
\par Insertion-only streams are a list of pairs of vertices, representing endpoints of edges. Each time we read an entry from an insertion-only stream we are adding the edge it describes to the graph. We apply a limitation that no duplicate edegs appear in an insertion-only stream and thus the order of the stream has no affect on the validity of the algorithms which shall be discussed.
\par Entries in an insertion-deletion list specify a pair of vertices, representing an edge, and a boolean which states whether this is an insertion or deletion entry. If it is an insertion entry then we are adding the described edge to the graph; if it is a deletion entry then we are removing it. Again we add a limitation that no edges are repeated \& that we can not recieve a request to delete an edge from the graph which does not currently exist in the graph. This does however mean that the order of an insertion-deletion is important as the deletion of an edge must come after it.
\par Despite these limitations it is still possible to construct almost any type of graph from a graph stream. Suppose you wish to analyse which devices in a network are making the most request. Here vertices represent devices \& edges represent requests. In this scenario you would want to allow for multiple edges (say $n>2)$ between the same pair of devices (say $\{x,y\}$). One solution would be to add $n$ new vertices $\{v_{x,y,1},\dots,v_{x,y,n}\}$ each with a label stating they are just indetermiate nodes and not full devices. Creating edges from $x$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ \& from $y$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ allows for an interpretation of multiple edges.
\par If a graph stream is ordered by time (which is common in real life applications) then it represents a dynamic-evolution of the graph over time. Allow the user to roll back the graph should they wish to make assessments during particular time preiods, rather than just assess the final graph.
\par The algorithms which shall be discussed in this paper are designed to work with just a single pass of a graph stream. This is not only time efficient as it does not require multiple reads of the stream, but means that should more information be gained after the algorithm is run it is easy to incorporate it into algorithm if we store the final state of the data structures used. Single pass algorithms mean we don't necessarily have to store the graph stream, we could simply send entries from the system being modelled straight into the algorithm, without being stored in a file. This is useful when a graph stream potentially could take up multiple terra-bytes, but is unnecessary in most applications.
\horizontalline
% Build to problem
\par The $\mathtt{High\ Degree\ Detection\ Probem}$ is a problem in graph theory where we seek to find a vertex of sufficiently high degree, in a given graph. This node does not necessarily have to have the greatest degree in the graph, and often we cannot be sure whether it does as the greatest degree is unknown beforehand. This problem can be formally stated as
\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
\textbf{Problem 1} $\texttt{High Degree Detection}$.\\
Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A,B$, where $|A|=n$ and $|B|=\text{poly }n$ , and edge set $E$. Set a restriction that at least one vertex in $A$ has degree, at least, $d$.\\
In $\texttt{High Degree Detection Detection}(G,d)$ we are tasked with outputting a vertex from $A$ with degree, at least, $d$.
}}\end{adjustwidth}\vspace{.3cm}
\par This problem can be solved efficiently for both insertion-only \& insertion-deletion graph streams with a fairly simple algorithm which simply runs through all the edges, incrementing or decrementing (depending on whether it is an insertion or deletion edge)  a count for the degree of each vertex as it goes. Returning the first vertex it finds with degree equal to $d$.\\
%TODO Do I have to justify the algorithm
\begin{algorithm}[H]
\caption{Single Pass High Degree Detection}
\SetKwInOut{Require}{require}
\Require{Stream $\{(i_0,x_0,y_0)\dots(i_n,x_n,y_n)\}$, degree bond $d$}
$D\leftarrow\{\{\}\}\ \{\text{degree map}\}$\\
\For{$j\in[0\dots n]$} {
	\uIf{$x_j\in D$}{
		\lIf{$i_j$}{$D[x_j]-=1$}
		\lElse{$D[x_j]+=1$}
	}
	\lElse{$D[x_j]=1$}
	\lIf{$D[x_j]=d$}{\Return{$x_j$}}
	\textbf{repeat} with $y_j$
}
\Return{FAIL}
\end{algorithm}\vspace{.3cm}
In \textbf{Algorithm 1} $i$ is a boolean which is true if the edge is an insertion edge, and false if it is a deletion edge. $x\ \&\ y$ are the two endpoints of the edge. The algorithm makes the reasonable assumption that you will never recieve a deletion edge for an edge which is not in the graph, nor shall you recieve the same edge twice.
\par The $\mathtt{High\ Degree\ Detection\ Probem}$ is limited in its applications as it only returns a vertex with high degree \& gives you no information as to its neighbourhood. Solving this problem would not be very helpful when it comes to choosing influencers as it would only return people who have large followings, but not who their followers are. Thus an advertisier would not be able to evaluate whether their followers represented the audience they wished to market to.
\par The $\mathtt{Neighbourhood\ Detection\ Problem}$ is an extension of the $\mathtt{High\ Degree}$ $\mathtt{Detection\ Problem}$. The task is to return a vertex with sufficiently high degree \& a certain proportion that vertex's neighbourhood. This is formally stated as
\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
%TODO Shorten this, make it neater.
\textbf{Problem 2} $\texttt{Neighbourhood Detection}$.\\
Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A,B$, where $|A|=n$ and $|B|=\text{poly }n$, and edge set $E$.\\
In $\texttt{Neighbourhood Detection}(G,d,c)$ we are tasked with outputting a vertex from $A$ and at least $d/c$ of its neighbours in $B$.\\
Here $d$ is a threshold parameter \& $c$ is an approximation parameter.
}}\end{adjustwidth}
\par Solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ allows for a much more general assessment of the influence of the returned vertex. In the context of finding social media influences the returned neighbourhood can be assessed to determine whether the returned person is suitable for a given advertising campaign. When assessing the members of the neighbourhood you could look at the demographics of the members \& how active the members are (among other features) in order to determine whether they fit the target of the campaign. For cases where $c\neq 1$ you only get a fraction of a vertices whole neighbourhood but for sufficiently high $d$ you an implement many statistical tools to still be able to draw meaningful inferences. Later it shall be shown that there are meaningful advantages to increasing the value of $c$.

\section{Why Space Efficiency?}
Simply solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ is fairly trivial. \textbf{Algorithm 2} is an algorithm for solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ for insertion-only graph streams by recording the vertices in the neighbourhood of every $A$-vertex and then returning the first neighbourhood we enounter with $\frac{d}{c}$ members.\\
\begin{algorithm}[H]
\caption{Na\"ive Single-Pass Insertion-Streaming Algorithm for Neighbourhood Detection}
\SetKwInOut{Require}{require}
\Require{Stream $\{(s_0,t_0)\dots(s_n,t_n)\}$, degree bound $d$, precision bound $c$}
$N\leftarrow\{\{\}\}\ \{\text{neighbourhoods}\}$\\
\For{$i=0\dots n$} {
	append $t_i$ to $N[s_i]$\\
	\lIf{\text{size}($N[s_i])\geq \frac{d}{c}$} {
		\Return{($s_i$,$N[s_i]$)}
	}
}
\Return{FAIL}
\end{algorithm}
%TODO check space requirements
\par\textbf{Algorithm 2} requires $O\left(n^2\right)$ space. Thus the space used to run the algorithm grows very quickly which could easily lead to the computer's RAM overflowing, increasing execution time. This is undesirable, especially given the run time complexity of \textbf{Algorithm 2} is alread bad, $O(n^2)$. % nc/d different vertices with neighbourhoods of d/c-1. Reading to position d/c-1, nc/d times
For solutions which are space efficient run times become unmanagable for large graph streams. In 2014 in the UK, Facebook had 36.68 million users \cite{users16} each with an average of 155 friend connections \cite{friends16}. In order for this to be represented in a bi-partite graph for $\mathtt{neighbourhood\ detection}$ we would need a graph stream with ${\sim}5.6$ billion edges. \textbf{Algorithm 2} would not be suitable for analysing this graph.
\par Instead we must look for much more space efficient algorithms for\\
$\mathtt{neighbourhood\ detection}$. In the rest of this paper I shall discuss \& test an algorithm which can solve $\mathtt{neighbourhood\ detection}$ for insertion-only streams with space $O(n\log n+n^{\frac1c}d\log^2n)$; and, two approaches to solving for $\mathtt{neighbourhood\ detection}$ for insertion-deletion streams with space $\tilde O\left(\frac{xd}{c}\right)\ \&\ \tilde O\left(\frac{nd}{c}\left(\frac1x+\frac1x\right)\right)$ respectively.

\section{Plan}
In this paper I shall implement and test the two algorithms presented by Dr Chrisian Konrad in \cite{orig}. These algorithms solve the \textit{Neighbourhood Detection Problem} for insertion-only \& insertion-deletion graph streams in a space-efficient manner. Both algorithms require only a single pass of the stream.
\par I shall use the na\"ive algorithm described in \textbf{Algorithm 2} as a benchmark for both time \& space efficiency of these two algorithms. I shall use a four different graph streams, described in \textbf{Table 1.1}, for my tests. % probs should have more graphs
These graphs were acquired from the \textit{Stanford Network Analysis Project}, \cite{SNAP}.
\par The theoretical space requirements for these two algorithms both depend on the degree variable, $d$, and precision variable, $c$. I shall perform tests where I vary each of these in order to see what the space requirements are in practice, and then compare these results to the theory.

\begin{table}[h]
\caption{Test Graph Streams}
\begin{tabular}{|l|l|l|l|l|}
\hline
Name&\# Edges&\# Vertices&Max Degree&File Size, KB\\\hline
$\mathtt{facebook\_small}$&292&52&36&3\\
$\mathtt{facebook}$&60,050&747&586&587\\
$\mathtt{gplus}$&1,179,613&12,417&5,948&52,839\\
$\mathtt{gplus\_large}$&30,238,035&120,100&104,947&1,328,820\\\hline
\end{tabular}
\label{Tab:Tcr}
\end{table}

\begin{thebibliography}{9}
\bibitem{orig} Christian Konrad \textit{Streaming Frequent Items with Timestamps and Detecting Large Neighborhoods in Graph Streams.} November 2019
\bibitem{targetedAds} Howard Beales \textit{The Value of Behavioral Targeting}.
\bibitem{users16} https://www.statista.com/statistics/553538/predicted-number-of-facebook-users-in-the-united-kingdom-uk/
\bibitem{friends16} https://www.telegraph.co.uk/news/science/science-news/12108412/Facebook-users-have-155-friends-but-would-trust-just-four-in-a-crisis.html
\bibitem{SNAP}http://snap.stanford.edu/data/
\end{thebibliography}

\end{document}

### 2020-02-07 insertionStreams.cpp
Fixed size of resevoirs

@@ -8,34 +8,34 @@
 #include <fstream>
 #include <iostream>
 #include <map>
+#include <math.h>
 #include <random>
 #include <string>
 #include <vector>
 
 using namespace std;
 
-int BYTES=0; // space used atm
-int MAX_BYTES=0; // max space used at any time
+int BYTES; // space used atm
+int MAX_BYTES; // max space used at any time
 
 /*-----------------*
  * DATA STRUCTURES *
  *-----------------*/
 
 using vertex = string; // typemap vertex
+using time_point=chrono::high_resolution_clock::time_point;
 
 struct edge { // undirected edge
   vertex fst;
   vertex snd;
 };
 
-typedef chrono::high_resolution_clock::time_point time_point;
-
 /*-----------*
 * SIGNATURES *
 *------------*/
 
 // size = size of resevoir
-void single_pass_insertion_stream(int c, int d, int size, ifstream& stream, vector<vertex>& neighbourhood, vertex& root);
+void single_pass_insertion_stream(int c, int d, int n,ifstream& stream, vector<vertex>& neighbourhood, vertex& root);
 void update_resevoir(vertex n, int d1, int d2, int count, int size, vector<vertex>& resevoir, vector<edge>& edges);
 void parse_edge(string str, edge& e);
 
@@ -44,33 +44,40 @@ void parse_edge(string str, edge& e);
 *------*/
 
 int main() {
-  int c=2, d=40, s=2; // c=runs, d/c=d2, s=size
-  ifstream stream("data/gplus.edges"); // file to read
-  vector<vertex> neighbourhood; vertex root; // variables for returned values
+  int d=104947, n=120100; // c=runs, d/c=d2, n=# vertices, NOTE - set d=max degree
+  for (int c=2;c<50;c++) {
+    BYTES=0; MAX_BYTES=0;
+    ifstream stream("data/gplus_large.edges"); // file to read
+    vector<vertex> neighbourhood; vertex root; // variables for returned values
 
-  time_point before=chrono::high_resolution_clock::now(); // time before execution
-  single_pass_insertion_stream(c,d,s,stream,neighbourhood,root); // NB does not tell you whose neighbourhood it is
-  time_point after=chrono::high_resolution_clock::now(); // time after execution
[…]

### 2020-02-07 naive.cpp

@@ -34,7 +34,7 @@ void naive_neighbourhood_detection(int d, ifstream& stream, vector<vertex>& neig
   *------*/
 
 int main() {
-  int d=30;
+  int d=297;
   ifstream stream("data/gplus.edges");
   vector<vertex> neighbourhood; vertex root; // variables for returned values
 

### 2020-02-07 utility.cpp

@@ -45,12 +45,13 @@ void merge_directory(const char* path, string output_name);
  *------*/
 
 int main() {
-  int count;
-  count_edges("data/gplus_large.edges",count);
-  cout<<count<<endl;
+  generate_insertion_deletion("data/gplus_large");
   string greatest; int deg;
   greatest_degree("data/gplus_large.edges",greatest,deg);
-  cout<<greatest<<" "<<deg<<endl;
+  cout<<"Greatest Degree "<<greatest<<" "<<deg<<endl;
+  int count;
+  count_edges("data/gplus_large_deletion.edges",count);
+  cout<<"# Edges "<<count<<endl;
 
   return 0;
  }
@@ -101,8 +102,11 @@ void generate_insertion_deletion(string file_name) {
   float p=0.1; // Probability to delete an edge // VARY THIS
   default_random_engine generator;
   bernoulli_distribution d(p);
+  int i=0;
 
   while (getline(stream,line)) {
+    i+=1;
+    if (i%10000==0) cout<<i<<", ";
     parse_edge(line,e);
     edges.push_back(e);
     outfile<<"I "<<e.fst<<" "<<e.snd<<endl;
@@ -114,7 +118,7 @@ void generate_insertion_deletion(string file_name) {
       edges.erase(edges.begin()+to_delete);
     }
   }
-
+  cout<<endl;
   stream.close();
   outfile.close();
 }
@@ -180,13 +184,14 @@ void greatest_degree(string file_name, string& vertex, int& degree) {
 
   while (getline(stream,line)) {
     i+=1;
-    if (i%10000==0) cout<<i<<",";
+    if (i%100000==0) cout<<i<<",";
     parse_edge(line,e);
 
     if (degrees.count(e.fst)) {
       if (e.insertion) degrees[e.fst]+=1;
       else degrees[e.fst]-=1;
     } else {
+      // You cannot delete an edge which does not exist
       degrees[e.fst]=1;
     }
 

### 2020-02-10 paper/src/paper.tex
Background Chapter

@@
 \usepackage{changepage}
+\usepackage{enumerate}
+\usepackage{enumitem}
 \usepackage{fancyhdr} % Header
@@
 % title
-\title{Implementing \& Evaluating Space Efficient Algoirthms for Detecting Large Neighbourhoods in Graph Streams}
+\title{Implementing \& Evaluating Space Efficient Algorithms for Detecting Large Neighbourhoods in Graph Streams}
 \author{Dom Hutchinson}
@@
 \section{Motivation}
-\par For most algorithms time efficiency is viewed with primary importance, rather than space efficiency.
-However if the memory used for an algorithm exceeds the size of the computer's RAM then virtual memory would have be used.
-Virtual memory has a much longer read/write than RAM meaning that the algorithm's run-time will be much greater in this scenario.
-Thus for algorithms which are designed to work with large data sets space efficiency is more important.
-\par In the early 2000s companies began to realise that data could be used as a commodity, and with this the amount of data being collected has surged in the years since.
-In this same time period we have nots seen the same surge in the amount of RAM typical computers have.
-Thus people wishing to capitalise on this increase in available data need to give greather thought to the space efficiency of the algorithms they are using.
-Otherwise run-times could become unusuable.
+\par When designing algorithms time efficiency is generally viewed with primary importance, over space efficiency.
+However, if the memory used for an algorithm exceeds the RAM space allocated to it during execution then virtual memory would have to be used.
+Virtual memory has a much longer read/write time than RAM meaning that an algorithm's run-time will be much greater in this scenario.
+Thus, algorithms which are designed to work with large data sets space efficiency should be viewed as having primary importance.
+\par In the early 2000s companies began to realise that data could be used as a commodity, and with this, the amount of data being collected has surged in the years since.
+In this same time period, we have not seen the same surge in the amount of RAM typical computers have.
+Thus people wishing to capitalise on this increase in available data need to give greater thought to the space efficiency of the algorithms they are using.
+Otherwise, run-times could become unusable.
 \par Social networks are good sources of data about people.
-This data can leveraged in many ways, notably within advertising.
-Behaviourly targeted adverts have been shown to have more than twice the click through rate of standard online advertiesments \cite{targetedAds}.
-The last decade has seen developments in the space of targeted advertising, one of which is the rise of social media influencers.
-These are people who are deemed to have a large online following, which they are able to influence.
-Advertisers will often approach these influencers with deals by which the influencer will promote a product to their audience, hopefully driving sales.
-\par The question becomes, ``Who is an influencer \& how do we find them?''.
+This data can be leveraged in many ways, notably within advertising.
+Advertisements which are targeted at people with certain observed behaviours have been shown to have more than twice the click-through rate of standard online advertisements \cite{targetedAds}.
+This has motivated the development of techniques for assessing the behaviours of a person just from their online presence.
+\par One technique is to look at who a person is connected to and then to use observations about these connections to infer the behaviours of said person.
+In turn, we have seen the rise of the ``Social Media Influencer".
+These are people who are deemed to have a large online following and, moreover, a following which they are able to influence.
+Advertisers will approach influencers with deals by which the influencer will promote a product to their audience, hopefully driving sales.
+\par The question for advertisers now becomes ``Who is an influencer?".
+To answer this we need to indentify people with large neigbourhoods in a network \& members of this network.
+We need to find the members of the neighbourhood (or at least a subset of them) in order to be able to assess what the demographics of the influencers audience are.
 This question can be generalised as the $\mathtt{Neighborhood\ Detection\ Problem}$.
 Which is the problem I shall discuss in this paper.
-%TODO
-Although in my description of the Neighbourhood Detection Problem I validate its relevance by referring to find social media influencers, it is applicable to many other scenarios.
+\par Although in my description of the Neighbourhood Detection Problem I motivate its relevance by referring to finding social media influencers, it is applicable to many other scenarios.
 \begin{itemize}
-	\item[-] TODO
+	\item[-] Which items are most commonly bought \& which other items they are commonly bought with?
+	\item[-] Which resources in a network are being accessed most \& by whom?
 \end{itemize}
+%TODO More
+Answering these questions can be used to inform resource allocation in order to maximise a certain goal. Be that profit; or, network response speeds.
 \section{Neighbourhood Detection Problem}
-%TODO Describe Graph streams, insertion-only & insertion-deletion
-\par Graph streams are a dynamic way of representing graphs.
-They are simply a list of instructions for how to construct a graph and in this paper I shall discuss two types: insertion-only streams \& insertion-deletion streams.
-There are other types which allow for updating vertex values, among other properties, but they are not relevant to the problem we are considering in this paper.
-\par Insertion-only streams are a list of pairs of vertices, representing endpoints of edges.
-Each time we read an entry from an insertion-only stream we are adding the edge it describes to the graph.
-We apply a limitation that no duplicate edegs appear in an insertion-only stream and thus the order of the stream has no affect on the validity of the algorithms which shall be discussed.
-\par Entries in an insertion-deletion list specify a pair of vertices, representing an edge, and a boolean which states whether this is an insertion or deletion entry.
-If it is an insertion entry then we are adding the described edge to the graph; if it is a deletion entry then we are removing it.
-Again we add a limitation that no edges are repeated \& that we can not recieve a request to delete an edge from the graph which does not currently exist in the graph.
-This does however mean that the order of an insertion-deletion is important as the deletion of an edge must come after it.
-\par Despite these limitations it is still possible to construct almost any type of graph from a graph stream.
+% Describe Graph streams, insertion-only & insertion-deletion
+\par Graph Streams are a list of instructions for how to construct a graph and provide a dynamic way of representing graphs.
+In this paper, I shall discuss two types of graph streams: Insertion-Only Streams; and, Insertion-Deletion Streams.
+There are other types of graph streams that allow for additional operations such as updating vertex \& edge values, but these are not relevant to the problem being considering in this paper.
+\par Insertion-Only graph streams are a list of pairs of vertices, representing endpoints of unweighted-undirected edges.
+Each instruction describes a new edge to be inserted into the graph \& potential up to two new nodes which are in the graph.
+There is no way to remove or update an edge in an Insertion-Only graph stream.
+We apply a limitation that no duplicate edges appear in an insertion-only stream and thus the order of instructions in an Inserionly-Only stream has no effect on the validity of the algorithms which shall be discussed.
+\par Insertion-Deletion graph streams consist of a pair of vertices and a boolean.
+The vertices represent the endpoints of an unweighted-undirected edge \& the boolean describes whether this edge is being inserted or deleted from the graph.
+If it is an insertion instruction then we are adding the described edge to the graph; if it is a deletion instruction then we are removing it.
+Again we add a limitation that no edges are repeated \& that we will not receive a request to delete an edge from the graph which does not currently exist in the graph.
+This does, however, mean that the order of instructions in an Insertion-Deletion stream is important as the deletion of an edge must come after it's insertion.
+\par The limitations placed here often occur naturally in systems. On social networks you are not able to follow someone you are already following, nor unfollow someone you are not following.
+\par In spite of these limitations which are imposed it is still possible to construct almost any undirected-unweighted graph using the two graph streams described.
 Suppose you wish to analyse which devices in a network are making the most request.
 Here vertices represent devices \& edges represent requests.
-In this scenario you would want to allow for multiple edges (say $n>2)$ between the same pair of devices (say $\{x,y\}$).
-One solution would be to add $n$ new vertices $\{v_{x,y,1},\dots,v_{x,y,n}\}$ each with a label stating they are just indetermiate nodes and not full devices.
-Creating edges from $x$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ \& from $y$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ allows for an interpretation of multiple edges.
-\par If a graph stream is ordered by time (which is common in real life applications) then it represents a dynamic-evolution of the graph over time.
-Allow the user to roll back the graph should they wish to make assessments during particular time preiods, rather than just assess the final graph.
+For this scenario, you would want to allow for multiple edges (say $n>2$) between the same pair of devices (say $\{x,y\}$).
+One solution would be to add $n$ new vertices $\{v_{x,y,1},\dots,v_{x,y,n}\}$ each with a label stating they are just intermediate nodes and not full devices.
+Creating edges from $x$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ \& from $y$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ allows for an interpretation of multiple occurrences of the edge $(x,y)$.
+\par If a graph stream is ordered by time (which is common in real-life applications) then it represents a dynamic-evolution of the graph over time.
+Allowing the user to roll back the graph should they wish to make assessments during particular time periods, rather than just assess the final graph.
 \par The algorithms which shall be discussed in this paper are designed to work with just a single pass of a graph stream.
-This is not only time efficient as it does not require multiple reads of the stream, but means that should more information be gained after the algorithm is run it is easy to incorporate it into algorithm if we store the final state of the data structures used.
-Single pass algorithms mean we don't necessarily have to store the graph stream, we could simply send entries from the system being modelled straight into the algorithm, without being stored in a file.
-This is useful when a graph stream potentially could take up multiple terra-bytes, but is unnecessary in most applications.
+This is not only time-efficient as it does not require multiple reads of the stream, but means that should more information be gained after the algorithm is run it is easy to incorporate it into the algorithm if we store the final state of the data structures used.
+Single-pass algorithms mean we don't necessarily have to store the graph stream in a local file, we could simply send instructions from the system being modeled straight into the algorithm.
+This is useful when a graph stream potentially could take up multiple terra-bytes, but is unnecessary in most applications and can mean it is harder to verify results later.
 \horizontalline
@@
 \par The $\mathtt{High\ Degree\ Detection\ Probem}$ is a problem in graph theory where we seek to find a vertex of sufficiently high degree, in a given graph.
-This node does not necessarily have to have the greatest degree in the graph, and often we cannot be sure whether it does as the greatest degree is unknown beforehand.
+This node does not necessarily have the greatest degree in the graph, and often we cannot be sure whether it does as the greatest degree is unknown beforehand.
 This problem can be formally stated as
@@
 \textbf{Problem 1} $\texttt{High Degree Detection}$.\\
-Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A,B$, where $|A|=n$ and $|B|=\text{poly }n$ , and edge set $E$.
+Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A\ \&\ B$, where $|A|=n$ and $|B|=\text{poly }n$ , and edge set $E$.
 Set a restriction that at least one vertex in $A$ has degree, at least, $d$.\\
-In $\texttt{High Degree Detection Detection}(G,d)$ we are tasked with outputting a vertex from $A$ with degree, at least, $d$.
+In $\texttt{High Degree Detection}(G,d)$ we are tasked with outputting a vertex from $A$ with degree, at least, $d$.
 }}\end{adjustwidth}\vspace{.3cm}
-\par This problem can be solved efficiently for both insertion-only \& insertion-deletion graph streams with a fairly simple algorithm which simply runs through all the edges, incrementing or decrementing (depending on whether it is an insertion or deletion edge)  a count for the degree of each vertex as it goes.
+\par This problem can be solved efficiently for both insertion-only \& insertion-deletion graph streams with a fairly simple algorithm which simply runs through all the edges, incrementing or decrementing (depending on whether it is an insertion or deletion instruction)  a count for the degree of each vertex as it goes.
 Returning the first vertex it finds with degree equal to $d$.\\
@@
 	\uIf{$x_j\in D$}{
-		\lIf{$i_j$}{$D[x_j]-=1$}
-		\lElse{$D[x_j]+=1$}
+		\lIf{$i_j$}{$D[x_j]+=1$}
+		\lElse{$D[x_j]-=1$}
 	}
@@
 $x\ \&\ y$ are the two endpoints of the edge.
-The algorithm makes the reasonable assumption that you will never recieve a deletion edge for an edge which is not in the graph, nor shall you recieve the same edge twice.
+The algorithm makes the reasonable assumption that you will never recieve a deletion-instruction for an edge which is not in the graph, nor an insertion-instruction for an edge which is already in the graph.
 \par The $\mathtt{High\ Degree\ Detection\ Probem}$ is limited in its applications as it only returns a vertex with high degree \& gives you no information as to its neighbourhood.
-Solving this problem would not be very helpful when it comes to choosing influencers as it would only return people who have large followings, but not who their followers are.
-Thus an advertisier would not be able to evaluate whether their followers represented the audience they wished to market to.
+Solving this problem would not be very helpful when it comes to identifying influencers as it would only return people who have large followings, but not who their followers are.
+Thus an advertisier would not be able to evaluate whether the returned person's followers represented the audience they wished to target.
 \par The $\mathtt{Neighbourhood\ Detection\ Problem}$ is an extension of the $\mathtt{High\ Degree}$ $\mathtt{Detection\ Problem}$.
@@
 Here $d$ is a threshold parameter \& $c$ is an approximation parameter.
-}}\end{adjustwidth}
-\par Solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ allows for a much more general assessment of the influence of the returned vertex.
-In the context of finding social media influences the returned neighbourhood can be assessed to determine whether the returned person is suitable for a given advertising campaign.
-When assessing the members of the neighbourhood you could look at the demographics of the members \& how active the members are (among other features) in order to determine whether they fit the target of the campaign.
-For cases where $c\neq 1$ you only get a fraction of a vertices whole neighbourhood but for sufficiently high $d$ you an implement many statistical tools to still be able to draw meaningful inferences.
+}}\end{adjustwidth}\vspace{.3cm}
+\par Solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ allows for a more general assessment of the influence of the returned vertex.
+In the context of finding social media influencers the returned neighbourhood can be assessed to determine whether the returned person is suitable for a given advertising campaign.
+When assessing the members of the neighbourhood you could look at the demographics of the members \& how active the members are, among other features, in order to determine whether they fit the target audience of the campaign.
+For cases where $c\neq 1$ you only get a subset of a person's whole neighbourhood but for sufficiently high $d$ you an implement statistical tools to still be able to draw meaningful inferences.
 Later it shall be shown that there are meaningful advantages to increasing the value of $c$.
@@
 Simply solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ is fairly trivial.
-\textbf{Algorithm 2} is an algorithm for solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ for insertion-only graph streams by recording the vertices in the neighbourhood of every $A$-vertex and then returning the first neighbourhood we enounter with $\frac{d}{c}$ members.\\
+\textbf{Algorithm 2} is an algorithm for solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ for insertion-only graph streams by recording the vertices in the neighbourhood of every $A$-vertex and then returning the first encountered neighbourhood with $\frac{d}{c}$ members.\\
 \begin{algorithm}[H]
@@
 	append $t_i$ to $N[s_i]$\\
-	\lIf{\text{size}($N[s_i])\geq \frac{d}{c}$} {
+	\lIf{\text{size}($N[s_i])= \frac{d}{c}$} {
 		\Return{($s_i$,$N[s_i]$)}
@@
 Thus the space used to run the algorithm grows very quickly which could easily lead to the computer's RAM overflowing, increasing execution time.
-This is undesirable, especially given the run time complexity of \textbf{Algorithm 2} is alread bad, $O(n^2)$.
-% nc/d different vertices with neighbourhoods of d/c-1.
-Reading to position d/c-1, nc/d times
-For solutions which are space efficient run times become unmanagable for large graph streams.
+For solutions which are space inefficient run times become unmanagable for large graph streams.
 In 2014 in the UK, Facebook had 36.68 million users \cite{users16} each with an average of 155 friend connections \cite{friends16}.
-In order for this to be represented in a bi-partite graph for $\mathtt{neighbourhood\ detection}$ we would need a graph stream with ${\sim}5.6$ billion edges.
+In order for this to be represented in a bi-partite graph for $\mathtt{Neighbourhood\ Detection}$ we would need a graph stream with ${\sim}5.6$ billion edges.
 \textbf{Algorithm 2} would not be suitable for analysing this graph.
 \par Instead we must look for much more space efficient algorithms for\\
-$\mathtt{neighbourhood\ detection}$.
-In the rest of this paper I shall discuss \& test an algorithm which can solve $\mathtt{neighbourhood\ detection}$ for insertion-only streams with space $O(n\log n+n^{\frac1c}d\log^2n)$; and, two approaches to solving for $\mathtt{neighbourhood\ detection}$ for insertion-deletion streams with space $\tilde O\left(\frac{xd}{c}\right)\ \&\ \tilde O\left(\frac{nd}{c}\left(\frac1x+\frac1x\right)\right)$ respectively.
+$\mathtt{Neighbourhood\ Detection}$.
+In the rest of this paper I shall discuss \& test an algorithm which can solve $\mathtt{Neighbourhood\ Detection}$ for insertion-only streams with space $O(n\log n+n^{\frac1c}d\log^2n)$; and, two approaches to solving for $\mathtt{Neighbourhood\ Detection}$ for insertion-deletion streams with space $\tilde O\left(\frac{xd}{c}\right)\ \&\ \tilde O\left(\frac{nd}{c}\left(\frac1x+\frac1x\right)\right)$ respectively.
 
@@
 In this paper I shall implement and test the two algorithms presented by Dr Chrisian Konrad in \cite{orig}.
-These algorithms solve the \textit{Neighbourhood Detection Problem} for insertion-only \& insertion-deletion graph streams in a space-efficient manner.
+These algorithms solve the $\mathtt{Neighbourhood\ Detection\ Problem}$ for insertion-only \& insertion-deletion graph streams in a space-efficient manner.
 Both algorithms require only a single pass of the stream.

### 2020-02-10 insertionStreams.cpp
Execute test - output results to file

@@ -1,6 +1,7 @@
 /*-----------------*
  * TODO
- * Record time take & space used
+ * Tests
+ *   - Set parameters & output results to csv
  *-----------------*/
 
 #include <algorithm>
@@ -35,6 +36,7 @@ struct edge { // undirected edge
 *------------*/
 
 // size = size of resevoir
+void execute_test(int c_min, int c_max, int c_step, int d, int n, string file_name, string out_file);
 void single_pass_insertion_stream(int c, int d, int n,ifstream& stream, vector<vertex>& neighbourhood, vertex& root);
 void update_resevoir(vertex n, int d1, int d2, int count, int size, vector<vertex>& resevoir, vector<edge>& edges);
 void parse_edge(string str, edge& e);
@@ -44,35 +46,37 @@ void parse_edge(string str, edge& e);
 *------*/
 
 int main() {
-  int d=104947, n=120100; // c=runs, d/c=d2, n=# vertices, NOTE - set d=max degree
-  for (int c=2;c<50;c++) {
-    BYTES=0; MAX_BYTES=0;
-    ifstream stream("data/gplus_large.edges"); // file to read
-    vector<vertex> neighbourhood; vertex root; // variables for returned values
+  int d=586, n=747;
+  execute_test(2,20,1,d,n,"data/facebook.edges","results/facebook_results.csv");
+  d=5948; n=12417;
+  execute_test(2,20,1,d,n,"data/gplus.edges","results/gplus_results.csv");
+  d=104947; n=120100; // c=runs, d/c=d2, n=# vertices, NOTE - set d=max degree, n=number of vertices
+  execute_test(3,20,1,d,n,"data/gplus_large.edges","results/gplus_large_results.csv");
+  return 0;
+}
+
+// Runs algorithm multiple time, writing results to a csv file
+void execute_test(int c_min, int c_max, int c_step, int d, int n, string file_name, string out_file) {
+  ofstream outfile(out_file);
+  outfile<<"name,"<<file_name<<endl<<"n,"<<n<<endl<<"d,"<<d<<endl<<endl; // test details
+  outfile<<"c,time (milliseconds),space (bytes),succeed"<<endl; // headers
+  vector<vertex> neighbourhood; vertex root; // variables for returned values
+  for (int c=c_min;c<=c_max;c+=c_step) {
+    cout<<c<<"/"<<c_max<<endl; // output to terminal
[…]

### 2020-02-11 insertionStreams.cpp
Test now runs each c multiple times

@@ -36,7 +36,7 @@ struct edge { // undirected edge
 *------------*/
 
 // size = size of resevoir
-void execute_test(int c_min, int c_max, int c_step, int d, int n, string file_name, string out_file);
+void execute_test(int c_min, int c_max, int c_step, int reps, int d, int n, string file_name, string out_file); // for (int c=c_min;c<=c_max;c+=c_step). Reps is the number of times each c is tested, average is taken.
 void single_pass_insertion_stream(int c, int d, int n,ifstream& stream, vector<vertex>& neighbourhood, vertex& root);
 void update_resevoir(vertex n, int d1, int d2, int count, int size, vector<vertex>& resevoir, vector<edge>& edges);
 void parse_edge(string str, edge& e);
@@ -46,35 +46,40 @@ void parse_edge(string str, edge& e);
 *------*/
 
 int main() {
-  int d=586, n=747;
-  execute_test(2,20,1,d,n,"data/facebook.edges","results/facebook_results.csv");
-  d=5948; n=12417;
-  execute_test(2,20,1,d,n,"data/gplus.edges","results/gplus_results.csv");
-  d=104947; n=120100; // c=runs, d/c=d2, n=# vertices, NOTE - set d=max degree, n=number of vertices
-  execute_test(3,20,1,d,n,"data/gplus_large.edges","results/gplus_large_results.csv");
+  //int d=586, n=747, reps=10;
+  //execute_test(2,20,1,reps,d,n,"data/facebook.edges","results/facebook_results.csv");
+  //d=5948; n=12417;
+  //execute_test(3,20,1,reps,d,n,"data/gplus.edges","results/gplus_results.csv");
+  int d=104947, n=120100, reps=10; // c=runs, d/c=d2, n=# vertices, NOTE - set d=max degree, n=number of vertices
+  execute_test(3,20,1,reps,d,n,"data/gplus_large.edges","results/gplus_large_results.csv");
   return 0;
 }
 
 // Runs algorithm multiple time, writing results to a csv file
-void execute_test(int c_min, int c_max, int c_step, int d, int n, string file_name, string out_file) {
+void execute_test(int c_min, int c_max, int c_step, int reps, int d, int n, string file_name, string out_file) {
   ofstream outfile(out_file);
-  outfile<<"name,"<<file_name<<endl<<"n,"<<n<<endl<<"d,"<<d<<endl<<endl; // test details
-  outfile<<"c,time (milliseconds),space (bytes),succeed"<<endl; // headers
[…]

### 2020-02-26 l0Sampler.cpp
l0 samplers for streams

@@ -12,9 +12,9 @@
 using namespace std;
 int P=1073741789; // >2^30
 
- /*-----------------*
-  * DATA STRUCTURES *
-  *-----------------*/
+/*-----------------*
+ * DATA STRUCTURES *
+ *-----------------*/
 
 // parameters for hash function
  struct hash_params {
@@ -23,9 +23,9 @@ int P=1073741789; // >2^30
    unsigned long m;
  };
 
- /*------------*
-  * SIGNATURES *
-  *------------*/
+/*------------*
+ * SIGNATURES *
+ *------------*/
 
 // hashing
 hash_params generate_hash(int m);

### 2020-02-26 l0StreamSampleMany.cpp

@@ -1,0 +1,114 @@
+/*
+ *  MULTIPLE L0-SAMPLERs APPLIED TO A STREAM (SINGLE PASS)
+ *  uniformly samples a term from a stream
+ *  --------------------------
+ *  NOTES
+ *   See l0StreamSampler.cpp
+ *   The Hash function used hear (26/02) maps from [n] to [n^3]. Given n can be in the millions this breaks down due to bad choice of prime.
+ *    maybe just map [n] to [n^2]. More collisions, but at least it works
+ */
+
+#include <algorithm>
+#include <chrono>
+#include <iostream>
+#include <fstream>
+#include <random>
+#include <set>
+#include <vector>
+
+using namespace std;
+int P=1073741789; // >2^30-1
+
+/*-----------------*
+ * DATA STRUCTURES *
+ *-----------------*/
+
+// parameters for hash function
+struct hash_params {
+  unsigned long a;
+  unsigned long b;
+  unsigned long m;
+};
+
+/*------------*
+ * SIGNATURES *
+ *------------*/
+
+// hashing
+hash_params generate_hash(unsigned long m);
+unsigned long hash_function(unsigned long key, hash_params ps);
+
+/*------*
+ * BODY *
+ *------*/
+
+int main() {
+  string edge_file_path="data/facebook_deletion.edges";
+  int n=66643; // number of edges
+  int d=522; // max degree
+  int c=200; // accuracy
+  //int num_samplers=20*n*(int)(d/c)*log(n)*(((float)1/max((int)n/c,(int)sqrt(n)))+(float)(1/c));
+  int num_samplers=100; // NOTE RUNS LIKE poorly FOR LARGE num_samplers (many due to first while loop which can't be optimised, i think?)
+  cout<<num_samplers<<endl;
+
+  set<string> output; hash_params ps;
+  vector<string>* samples[num_samplers];
+  for (int i=0; i<num_samplers; i++) samples[i]=new vector<string>;
+
+  ifstream stream(edge_file_path);
+  int i=0; int j=1;
+  string line; unsigned long lim=n/pow(2,j);
+  while (getline(stream,line)) { // single pass of stream
+    for (int k=0; k<num_samplers; k++) {
+      ps=generate_hash(n);
+      unsigned long h=hash_function(i,ps);
+      if (h<=lim) samples[k]->push_back(line);
+    }
+    i+=1;
+  }
+
+  for (int k=0; k<num_samplers; k++) {
[…]

### 2020-02-26 l0StreamSampler.cpp

@@ -1,0 +1,103 @@
+/*
+ *  AN L0-SAMPLER FOR A STREAM
+ *  uniformly samples a term from a stream
+ *  --------------------------
+ *  NOTES
+ *    Probability of index i being sampled from vector |x| with x in R^n
+ *      depends on how many non-zero terms are in x, and not on which terms
+ *      are non-zero.
+ *   For a stream, all entries are considered non-zero so the probability
+ *      of index i being sampled depends on n (the length of the stream) only
+ */
+
+#include <chrono>
+#include <iostream>
+#include <fstream>
+#include <random>
+#include <vector>
+
+using namespace std;
+int P=1073741789; // >2^30-1
+
+/*-----------------*
+ * DATA STRUCTURES *
+ *-----------------*/
+
+ // parameters for hash function
+struct hash_params {
+  unsigned long a;
+  unsigned long b;
+  unsigned long m;
+};
+
+/*------------*
+ * SIGNATURES *
+ *------------*/
+
+// hashing
+hash_params generate_hash(unsigned long m);
+unsigned long hash_function(unsigned long key, hash_params ps);
+
+
+/*------*
+ * BODY *
+ *------*/
+
+int main() {
+  string edge_file_path="data/facebook_deletion.edges";
+  int n=66643; // number of edges in file
+  vector<string> sample;
+  hash_params ps=generate_hash(pow(n,3));
+
+  ifstream stream(edge_file_path);
+  int j=22; // level //TODO this is a bodge due to issue with hash function (FIND BETTER HASH FUNCTION)
+  int i=0; string line; unsigned long lim=pow(n,3)/pow(2,j);
+  while (getline(stream,line)) { // single pass of stream
+    unsigned long h=hash_function(i,ps);
+    if (h<=lim) sample.push_back(line);
+    i+=1;
+  }
+
+  while(sample.size()>1) {
+    ps=generate_hash(pow(n,3)); // change hash function
+    j+=1;
+    vector<string> new_sample;
+    lim=pow(n,3)/pow(2,j);
+    for (int i=0; i<sample.size(); i++) {
+      int h=hash_function(i,ps);
+      if (h<=lim) new_sample.push_back(sample[i]);
+    }
+    sample=new_sample;
+  }
+
+  if (sample.size()==1) {
+    cout<<"SUCCESS "<<sample[0]<<endl;
+  } else {
+    cout<<"FAIL"<<endl;
+  }
+
+  return 0;
+
+}
+
+/*---------*
+ * HASHING *
[…]

### 2020-02-28 naive.cpp
Restructured

@@ -1,138 +1,0 @@
-#include <algorithm>
-#include <chrono>
-#include <fstream>
-#include <iostream>
-#include <map>
-#include <string>
-#include <vector>
-
-using namespace std;
-
-int BYTES=0; // space used atm
-
-/*-----------------*
- * DATA STRUCTURES *
- *-----------------*/
-using time_point=chrono::high_resolution_clock::time_point;
-using vertex=string;
-
-struct edge {
-  vertex fst;
-  vertex snd;
-  bool insertion;
-};
-
-/*------------*
- * SIGNATURES *
- *------------*/
-
-void parse_edge(string str, edge& e);
-void naive_neighbourhood_detection(int d, ifstream& stream, vector<vertex>& neighbourhood, vertex& root);
-
- /*------*
-  * BODY *
-  *------*/
-
-int main() {
-  int d=297;
-  ifstream stream("data/gplus.edges");
-  vector<vertex> neighbourhood; vertex root; // variables for returned values
-
-  time_point before=chrono::high_resolution_clock::now(); // time before execution
-  naive_neighbourhood_detection(d,stream,neighbourhood,root);
-  time_point after=chrono::high_resolution_clock::now(); // time before execution
-
-  // Print out returned neighbourhood, if one exists
-  vertex* p=&root;
-  if (p==nullptr) cout<<"NO SUCCESSES"<<endl;
-  else {
-    cout<<"Neighbourhood for <"<<root<<">"<<endl;
-    cout<<"<";
-    for (vector<vertex>::iterator i=neighbourhood.begin(); i!=neighbourhood.end(); i++) cout<<*i<<",";
-    cout<<">"<<endl;
-  }
-  stream.close();
-
-  auto duration = chrono::duration_cast<chrono::microseconds>(after-before).count();
-  cout<<"Execution Time - "<<duration<<" microseconds ("<<duration/1000<<" milliseconds)"<<endl;
-  cout<<"Max space - "<<BYTES<<" bytes ("<<BYTES/1024<<" kb)"<<endl;
-  return 0;
-}
-
-// naive solution for insertion-only stream
-void naive_neighbourhood_detection(int d, ifstream& stream, vector<vertex>& neighbourhood, vertex& root) {
-  string line; edge e; map<vertex,vector<vertex> > neighbourhoods;
[…]

### 2020-02-28 src/insertionDeletionStreams/insertionDeletionStreams.cpp

@@ -72,14 +72,12 @@ bool* copy_arr(bool* arr, int size);
  *------*/
 
 int main() {
-  string edge_file_path="data/gplus_deletion.edges",vertex_file_path="data/gplus.vertices";
+  string edge_file_path="../../../data/gplus_deletion.edges",vertex_file_path="../../../data/gplus.vertices";
   int n=12417,c=200,d=4998;
   /*
-  string edge_file_path="data/gplus_large_deletion.edges", vertex_file_path="data/gplus.vertices";
+  string edge_file_path="../../../data/gplus_large_deletion.edges", vertex_file_path="../../../data/gplus.vertices";
   int n=102100, c=200, d=10000;
   */
-
-  cout<<"CCCCCCCCCCCCCCCCCCCCCCCCCCCC="<<c<<endl;
   vertex root; set<vertex> neighbourhood;
 
   time_point before=chrono::high_resolution_clock::now(); // time before execution
@@ -98,6 +96,15 @@ int main() {
   cout<<"Duration="<<duration/1000<<"ms"<<endl;
   if (BYTES>MAX_BYTES) MAX_BYTES=BYTES;
   cout<<"BYTES="<<BYTES/pow(1024,2)<<"mb, MAX_BYTES="<<MAX_BYTES/pow(1024,2)<<"mb"<<endl;
+
+  /*
+   *  EDGE SAMPLING
+   *  Approaches
+   *   1. Sampling a mix of insertion & deletion edges then reconstructing?
+   *      Seems dody as likely to have incorrect neighbourhood
+   *   2. Find all final edges & sample from these?
+   *      Seems worse than vertex sampling as need to know all edges
+   */
 
   return 0;
 }

### 2020-02-28 src/insertionDeletionStreams/l0Samplers/l0StreamSampleMany.cpp

@@ -43,7 +43,7 @@ unsigned long hash_function(unsigned long key, hash_params ps);
  *------*/
 
 int main() {
-  string edge_file_path="data/facebook_deletion.edges";
+  string edge_file_path="..././data/facebook_deletion.edges";
   int n=66643; // number of edges
   int d=522; // max degree
   int c=200; // accuracy
@@ -57,10 +57,10 @@ int main() {
 
   ifstream stream(edge_file_path);
   int i=0; int j=1;
-  string line; unsigned long lim=n/pow(2,j);
+  string line; unsigned long lim=(10*n)/pow(2,j);
   while (getline(stream,line)) { // single pass of stream
     for (int k=0; k<num_samplers; k++) {
-      ps=generate_hash(n);
+      ps=generate_hash(10*n);
       unsigned long h=hash_function(i,ps);
       if (h<=lim) samples[k]->push_back(line);
     }
@@ -71,10 +71,10 @@ int main() {
     vector<string>& sample=*samples[k];
     int j=1;
     while(sample.size()>1) {
-      ps=generate_hash(n);
+      ps=generate_hash(10*n);
       j+=1;
       vector<string> new_sample;
-      lim=n/pow(2,j);
+      lim=(10*n)/pow(2,j);
       for (int i=0; i<sample.size(); i++) {
         int h=hash_function(i,ps);
         if (h<=lim) new_sample.push_back(sample[i]);

### 2020-02-28 src/insertionDeletionStreams/l0Samplers/l0StreamSampler.cpp

@@ -44,14 +44,14 @@ unsigned long hash_function(unsigned long key, hash_params ps);
  *------*/
 
 int main() {
-  string edge_file_path="data/facebook_deletion.edges";
+  string edge_file_path="../../data/facebook_deletion.edges";
   int n=66643; // number of edges in file
   vector<string> sample;
-  hash_params ps=generate_hash(pow(n,3));
+  hash_params ps=generate_hash(10*n);
 
   ifstream stream(edge_file_path);
   int j=22; // level //TODO this is a bodge due to issue with hash function (FIND BETTER HASH FUNCTION)
-  int i=0; string line; unsigned long lim=pow(n,3)/pow(2,j);
+  int i=0; string line; unsigned long lim=(10*n)/pow(2,j);
   while (getline(stream,line)) { // single pass of stream
     unsigned long h=hash_function(i,ps);
     if (h<=lim) sample.push_back(line);
@@ -59,10 +59,10 @@ int main() {
   }
 
   while(sample.size()>1) {
-    ps=generate_hash(pow(n,3)); // change hash function
+    ps=generate_hash(10*n); // change hash function
     j+=1;
     vector<string> new_sample;
-    lim=pow(n,3)/pow(2,j);
+    lim=(10*n)/pow(2,j);
     for (int i=0; i<sample.size(); i++) {
       int h=hash_function(i,ps);
       if (h<=lim) new_sample.push_back(sample[i]);

### 2020-02-28 src/insertionStreams/insertionStreams.cpp

@@ -48,11 +48,11 @@ void parse_edge(string str, edge& e);
 
 int main() {
   //int d=586, n=747, reps=100;
-  //execute_test(2,20,1,reps,d,n,"data/facebook.edges","results/facebook_results.csv");
+  //execute_test(2,20,1,reps,d,n,"../../data/facebook.edges","results/facebook_results.csv");
   //int d=5948, n=12417, reps=20;
-  //execute_test(2,2,1,reps,d,n,"data/gplus.edges","results/gplus_results.csv");
-  d=104947, n=102100, reps=10; // c=runs, d/c=d2, n=# vertices, NOTE - set d=max degree, n=number of vertices
-  execute_test(3,20,1,reps,d,n,"data/gplus_large.edges","results/gplus_large_results.csv");
+  //execute_test(2,2,1,reps,d,n,"../../data/gplus.edges","results/gplus_results.csv");
+  //int d=104947, n=102100, reps=10; // c=runs, d/c=d2, n=# vertices, NOTE - set d=max degree, n=number of vertices
+  //execute_test(9,17,8,reps,d,n,"../../data/gplus_large.edges","results/gplus_large_results.csv");
   return 0;
 }
 

### 2020-02-28 paper/src/paper.tex
l0 algorithms

@@
 \fancyhead[R]{\today}
+
+\DontPrintSemicolon % Don't print semicolons in algorithm environment
 
@@
 
+\newpage
+\begin{algorithm}[H]
+\caption{Single $l_0$-Sampler for vector}
+\SetKwInOut{Require}{require}
+\Require{Vector $\textbf{a}=(a_1,\dots,a_n)$}
+$j=1$.\\
+$\textbf{a}_j=\textbf{a}$.\\
+\While{true}{
+	H=new hash function.\\
+	$a_{j+1}=\pmb0$.\\
+	$j=j+1$.\\
+	\tcp{Sampling}
+	\For{$i\in[1,n]$}{
+		$h=H(i)$.\\
+		\lIf{$h\leq\frac{n^3}{2^j}$}{$a_{j+1,i}=a_{j,i}$}
+		\lElse{$a_{j+1,i}=0$}
+	}
+	$\textbf{a}_{j}=\textbf{a}_{j+1}$
+	\;
+	\tcp{Attempt Recover}
+	$w_1:=\sum a_{j,i}$.\\
+	$w_2:=\sum ia_{j,i}$.\\
+	\;
+	\If{$w_2\equiv0$}{\Return$\mathtt{FAIL}$\tcp*{$a_j\equiv\pmb0$}}
+	\If{$w_1\equiv1$}{\Return$w_2$\tcp*{Index of non-zero value}}
+}
+\end{algorithm}
+Hash functions should be k-wise independent and map from $n$ to $n^3$ where $n$ is the number of vertices in the graph (\ie length of vector since it is from a graph matrix).
+
+\begin{algorithm}[H]
+\caption{Single $l_0$-Sampler for a stream}
+\SetKwInOut{Require}{require}
+\Require{Stream $\textbf{s}=\{s_1,\dots,s_n\}$ with $n$ known, Hash Function $H$}
+sampled=\{\}.\\
+\For{$i\in[1,n\}$}{
+$h=H(i)$.\\
+\lIf{$h\leq\frac{n^3}{2^1}$}{sampled.append($s_i$)}
+}
+$j=1$.\\
+\While{sampled.size()$>1$}{
+  $j=j+1$.\\
+  new\_sample=\{\}.\\
+  \For{$i\in[1,\text{sampled.size()}]$}{
+    $h=H(i)$.\tcp*{Consider changing hash function}
+    \lIf{$h\leq\frac{n^3}{2^j}$}{new\_sample.append(sampled[i])}
+  }
+  sampled=new\_sample.
+}
+\lIf{sampled.size()$\equiv1$}{\Return{Success,sampled[0]}}
+\Return{Fail,NULL}
+\end{algorithm}
+
+\begin{algorithm}[h]
+\caption{$m$ Parallel $l_0$-Samplers for a stream}
+\SetKwInOut{Require}{require}
+\Require{Stream $\textbf{s}=\{s_1,\dots,s_n\}$ with $n$ known, Hash Functions $\textbf{H}=\{H_1,\dots,H_m\}$}
+output=\{\}.\\
+sampled=\{\{\}\}.\\
+\For{$i\in[1,n]$}{
+\tcp{Single pass of stream}
+  \For{$j\in[1,m]$}{
+    $h=H_j(i)$.\\
+    \lIf{$h\leq\frac{n^3}{2^1}$}{sampled[j].append($s_i$)}
+  }
+}
+\For{$j\in[1,m]$}{
+  $k=1$.\\
+  \While{sampled[j].size()$>1$}{
+    $k=k+1$.\\
+    new\_sample=\{\}.\\
+    \For{$i\in[1,\text{sampled[j].size()}]$}{
+      $h=H_j(i)$.\tcp*{Consider changing hash function}
+      \lIf{$h\leq\frac{n^3}{2^k}$}{new\_sample[j].append(sampled[i])}
+    }
+    sampled[j]=new\_sample.
+  }
+  \lIf{sampled[j].size()$\equiv1$}{output.append(sampled[j][0])}
+}
+\Return{output}
+\end{algorithm}
+
 \begin{thebibliography}{9}

### 2020-02-28 src/insertionStreams/naive.cpp
Naive implementation of insertion stream

@@ -1,0 +1,134 @@
+/*
+ *  Anaive implementation of neighbourhood detection algorithm for insertion-only Streams
+ *  See algorithm 2 in paper
+ */
+
+#include <algorithm>
+#include <chrono>
+#include <fstream>
+#include <iostream>
+#include <map>
+#include <string>
+#include <set>
+
+using namespace std;
+
+int BYTES=0; // space used atm
+
+/*-----------------*
+ * DATA STRUCTURES *
+ *-----------------*/
+using time_point=chrono::high_resolution_clock::time_point;
+using vertex=string;
+
+struct edge { // undirected edge
+  vertex fst;
+  vertex snd;
+};
+
+/*------------*
+ * SIGNATURES *
+ *------------*/
+
+void update(vertex v1, vertex v2, map<vertex,int>& degrees, map<vertex,set<vertex> >& neighbourhoods);
+int naive(string edge_file_path, int c, int d, vertex& root, set<vertex>& neighbourhood);
+void parse_edge(string str, edge& e);
+
+/*------*
+ * BODY *
+ *------*/
+
+int main() {
+  int d=586; int c=20; // d=max_degree,c=accuracy
+  string edge_file_path="../../data/facebook.edges";
+
+  vertex root;
+  set<vertex> neighbourhood;
+
+  time_point before=chrono::high_resolution_clock::now();
+  int edge_count=naive(edge_file_path,c,d,root,neighbourhood);
+  time_point after=chrono::high_resolution_clock::now();
+
+  auto duration=chrono::duration_cast<chrono::microseconds>(after-before).count();
+
+  if(edge_count!=-1) {
+    cout<<"root="<<root<<endl<<"# edges checked="<<edge_count<<endl<<"Neighbourhood ("<<neighbourhood.size()<<")"<<endl; // output result
+    cout<<"time="<<duration<<"mus"<<endl<<"space used="<<BYTES/1024<<"mb"<<endl<<"{";
+    for (set<vertex>::iterator it=neighbourhood.begin(); it!=neighbourhood.end(); it++) cout<<*it<<",";
+    cout<<"}"<<endl;
+  }
+
+
+  return -1;
+}
+
+int naive(string edge_file_path, int c, int d, vertex& root, set<vertex>& neighbourhood) {
+  map<vertex,int> degrees;
+  map<vertex,set<vertex> > neighbourhoods;
[…]

### 2020-02-28 src/insertionStreams/naive.cpp
Insertion Naive Tets

@@ -30,8 +30,9 @@ struct edge { // undirected edge
  * SIGNATURES *
  *------------*/
 
-void update(vertex v1, vertex v2, map<vertex,int>& degrees, map<vertex,set<vertex> >& neighbourhoods);
+void execute_test(int c_min, int c_max, int c_step, int d, string in_file, string out_file);
 int naive(string edge_file_path, int c, int d, vertex& root, set<vertex>& neighbourhood);
+void update(vertex v1, vertex v2, map<vertex,int>& degrees, map<vertex,set<vertex> >& neighbourhoods);
 void parse_edge(string str, edge& e);
 
 /*------*
@@ -41,25 +42,40 @@ void parse_edge(string str, edge& e);
 int main() {
   int d=586; int c=20; // d=max_degree,c=accuracy
   string edge_file_path="../../data/facebook.edges";
+  string output_file_path="../../results/facebook_naive_results.csv";
 
-  vertex root;
-  set<vertex> neighbourhood;
-
-  time_point before=chrono::high_resolution_clock::now();
-  int edge_count=naive(edge_file_path,c,d,root,neighbourhood);
-  time_point after=chrono::high_resolution_clock::now();
-
-  auto duration=chrono::duration_cast<chrono::microseconds>(after-before).count();
-
-  if(edge_count!=-1) {
-    cout<<"root="<<root<<endl<<"# edges checked="<<edge_count<<endl<<"Neighbourhood ("<<neighbourhood.size()<<")"<<endl; // output result
-    cout<<"time="<<duration<<"mus"<<endl<<"space used="<<BYTES/1024<<"mb"<<endl<<"{";
-    for (set<vertex>::iterator it=neighbourhood.begin(); it!=neighbourhood.end(); it++) cout<<*it<<",";
-    cout<<"}"<<endl;
-  }
-
+  execute_test(2,100,1,d,edge_file_path,output_file_path);
 
   return -1;
+}
+
+// deterministic so dont need repetitions
+void execute_test(int c_min, int c_max, int c_step, int d, string in_file, string out_file) {
+  ofstream outfile(out_file);
+  outfile<<"name,"<<in_file<<endl<<"d,"<<d<<endl<<endl<<endl;
+  outfile<<"c,time (microseconds),space (bytes),mean egdes checked"<<endl;
+
+  vertex root; set<vertex> neighbourhood;
+  for (int c=c_min;c<=c_max;c+=c_step) {
+    BYTES=0;
[…]

### 2020-03-03 poster/src/algorithms.tex
Poster

\documentclass[11pt,twoside,a4paper]{article}

\usepackage[ruled, vlined, linesnumbered]{algorithm2e}
\usepackage{algcompatible}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{changepage}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{fancyhdr} % Header
\usepackage[margin=1.4in]{geometry}

\pagenumbering{arabic}

\algblockdefx{FORALLP}{ENDFAP}[1]%
  {\textbf{for all }#1 \textbf{do in parallel}}%
  {\textbf{end for}}

\begin{document}

\begin{algorithm}
\caption{Na\"ive Single-Pass Insertion-Streaming Algorithm for Neighbourhood Detection}
\SetKwInOut{Require}{require}
\Require{Stream $\{(s_0,t_0)\dots(s_n,t_n)\}$, degree bound $d$, precision bound $c$}
$N\leftarrow\{\{\}\}\ \{\text{neighbourhoods}\}$\\
\For{$i=0\dots n$} {
	append $t_i$ to $N[s_i]$\\
	\lIf{\text{size}($N[s_i])= \frac{d}{c}$} {
		\Return{($s_i$,$N[s_i]$)}
	}
}
\Return{FAIL}
\end{algorithm}

\begin{algorithm}
\caption{One-pass $c$-approximation Insertion-Only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
\SetKwInOut{Require}{require}
\Require{Space $s$, degree bound $d$.}
$s\leftarrow\lceil\ln(n)\cdot n^{\frac1c}\rceil$\\
\For{$i\in[0,c-1]$\textbf{ in parallel}}{
	$(a_i,S_i)\leftarrow\text{Deg-Res-Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
}
\Return{Uniform random neighbourhood $(a_i,S_i)$ from the successful runs}
\end{algorithm}

\begin{algorithm}
\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Vertex Sampling)}}
\SetKwInOut{Require}{require}
\Require{Space $s$, degree bound $d$.}
Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$\\
Sample a uniform random subset $A'\subseteq A$ of size $10\ x\ln n$ of vertices.\\
\For{$a\in A'$}{
Run $10\frac{d}{c}\ln n$ $l_0$-samplers on the set of edges incident to $a$.
}
\Return{Any neghbourhood of size $\frac{d}{c}$ among the stored edges, if there is one.}
\end{algorithm}

\begin{algorithm}
\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Edge Sampling)}}
\SetKwInOut{Require}{require}
\Require{Space $s$, degree bound $d$.}
Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$\\
Run $20\frac{nd}{c}\left(\frac1x+\frac1c\right)\ln(n)$ $l_0$ samplers on the stream, in parallel. Each producing a uniform random edge.\\
\Return{Any neghbourhood of size $\frac{d}{c}$ among the stored edges, if there is one.}
\end{algorithm}

\end{document}

### 2020-03-03 poster/src/poster.tex
Poster

% The document class supplies options to control rendering of some standard
% features in the result.  The goal is for uniform style, so some attention 
% to detail is *vital* with all fields.  Each field (i.e., text inside the
% curly braces below, so the MEng text inside {MEng} for instance) should 
% take into account the following:
%
% - author name       should be formatted as "FirstName LastName"
%   (not "Initial LastName" for example),
% - supervisor name   should be formatted as "Title FirstName LastName"
%   (where Title is "Dr." or "Prof." for example),
% - degree programme  should be "BSc", "MEng", "MSci", "MSc" or "PhD",
% - dissertation title should be correctly capitalised (plus you can have
%   an optional sub-title if appropriate, or leave this field blank),
% - dissertation type should be formatted as one of the following:
%   * for the MEng degree programme either "enterprise" or "research" to
%     reflect the stream,
%   * for the MSc  degree programme "$X/Y/Z$" for a project deemed to be
%     X%, Y% and Z% of type I, II and III.
% - year              should be formatted as a 4-digit year of submission
%   (so 2014 rather than the accademic year, say 2013/14 say).
%
% Note there is a *strict* requirement for the poster to be in portrait 
% format so that we display them on the poster boards available.

\documentclass[ % the name of the author
                    author={Dominic Hutchinson},
                % the name of the supervisor
                supervisor={Dr. Christian Konrad},
                % the degree programme
                    degree={MEng Maths  and Computer Science},
                % the dissertation    title (which cannot be blank)
                     title={Implementing and Evaluating Space Efficient Algorithms for Detecting Large Neighbourhoods in Graph Streams},
                % the dissertation subtitle (which can    be blank)
                  subtitle={},
                % the dissertation     type
                      type={Research},
                % the year of submission
                      year={2020} ]{poster}

\usepackage[ruled, vlined]{algorithm2e}
\SetAlFnt{\small}
\SetAlCapFnt{\small}
\SetAlCapNameFnt{\small}
\usepackage{algcompatible}
\usepackage{graphicx}

\begin{document}

% -----------------------------------------------------------------------------

\begin{frame}{}

\vfill

\begin{columns}[t]
  \begin{column}{0.900\linewidth}
  \begin{block}{\Large Graph Streams \& The $\mathtt{Neighbourhood\ Detection\ Problem}$}
 	\noindent
	Graph Streams are a sequence of instructions which describe how to construct a graph. Each instruction provides details about an edge in the graph. Graph Streams come in two forms
	\begin{itemize}
		\item Instructions in an \textit{Insertion-Only Streams} only ever insert new edges into the graph.
		\item Instructions in an \textit{Insertion-Deletion Streams} either insert a new edge, or remove an existing edge, from the graph.
	\end{itemize}
	In the $\mathtt{Neighbourhood\ Detection\ Problem}(n,d)$ we are given a graph with $n$ vertices \& at least one vertex of degree $d$. We are tasked to return a vertex with at least $\frac{d}c$ of its neighbours, from some approximation factor $c\geq1$.\\
	\noindent
	There are some good motivativing applications of the $\mathtt{Neighbourhood\ Detection\ Problem}$.
	\begin{itemize}
		\item Given a network of social media connections detect popular influencers and analyse the demographics they attract in order to plan targetted advertising campaigns.
		\item Given a log of traffic to a network detect whether a DDOS attack has occurred and, if so, by whom.
		\item Given a list of sales made on a website detect which items are most popular and which items they are commonly bought with.
	\end{itemize}
  \end{block}
  \end{column}
\end{columns}

\vfill

\begin{columns}[t]

  \begin{column}{0.422\linewidth}
  \begin{block}{\Large 1. Insertion-Stream Algorithm}
Below is a proposed algorithm for solving the $\mathtt{Neighbourhood}$ $\mathtt{Detection}$ $\mathtt{Problem}$ for \textit{Insertion-Only Streams}. This algorithm requires $O(n\log n+n^{\frac1c}d\log^2n)$ space in theory.
\vspace{.5cm}\\
\begin{algorithm}[H]
\caption{One-pass $c$-approximation Insertion-Only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
\SetKwInOut{Require}{require}
\Require{Space $s$, degree bound $d$.}
$s\leftarrow\lceil\ln(n)\cdot n^{\frac1c}\rceil$\\
\For{$i\in[0,c-1]$\textbf{ in parallel}}{
	$(a_i,S_i)\leftarrow\mathtt{Deg}\mbox{-}\mathtt{Res}\mbox{-}\mathtt{Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
}
\Return{\small Uniform random neighbourhood $(a_i,S_i)$ \text{from successful runs}}
\end{algorithm}
\vspace{.5cm}
	$\mathtt{Deg}\mbox{-}\mathtt{Res}\mbox{-}\mathtt{Sampling}(d_1,d_2,s)$ uniformly samples $s$ times from the set of nodes with degree at least $d_1$. For each of these sampled nodes a neighbourhood of size $\min\{d_2,d-d_1+1\}$ is stored.
  \end{block}
  \end{column}

  \begin{column}{0.422\linewidth}
  \begin{block}{\Large 2. Insertion-Deletion Algorithm}
Below is a proposed algorithm for solving the $\mathtt{Neighbourhood}$ $\mathtt{Detection}$ $\mathtt{Problem}$ for \textit{Insertion-Deletion Streams} by first sample a set of vertices. This algorithm requires $O(\frac{xd}c\log^kn)$ space in theory.
\vspace{.5cm}\\
\begin{algorithm}[H]
\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Vertex Sampling)}}
\SetKwInOut{Require}{require}
\Require{Space $s$, degree bound $d$.}
Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$\\
Sample a uniform random subset $A'\subseteq A$ of size $10\ x\ln n$ of vertices.\\
\For{$a\in A'$}{
Run $10\frac{d}{c}\ln n$ $l_0$-samplers on the set of edges incident to $a$.
}
\Return{\small Any neghbourhood of size $\frac{d}{c}$ among the stored edges, if there is one.}
\end{algorithm}
\vspace{.5cm}
\textit{$l_0$ Samplers} return an index which has been uniformly sampled from the indicies of non-zero elements of a vector.
  \end{block}
  \end{column}
\end{columns}

\vfill

\begin{columns}[t]
  \begin{column}{0.422\linewidth}
  \begin{block}{\Large 3. Preliminary Results}
	I have implemented the \textit{Insertion-Stream Algorithm} and a na\"ive algorithm for solving the problem in order to compare the results.
	\begin{itemize}
		\item[Fig1] Space requirements for each algorithm as the approximation factor is varied on the same graph.
		\item[Fig2] Space requirement for the proposed algorithm against it's theoretical space requirements.
\end{itemize}		
Both figures show promising results for the proposed algorithm.
	\begin{center}\begin{tabular}{c}
	\includegraphics[scale=1]{img/Figure1.png}
	\end{tabular}\end{center}	
  \end{block}
  \end{column}
  \begin{column}{0.422\linewidth}
  \begin{block}{\Large 4. Further Aims}
  \begin{itemize}
  \item Perform tests on different size graphs for the same approximation factor.
  \item Perform tests on different density graphs for the same approximation factor.
  \item Implement the proposed algorithm for \textit{Insertion-Deletion Streams}.
  \item Investigate an application of these algorithms.
  \end{itemize}
	\begin{center}\begin{tabular}{c}
	\includegraphics[scale=1]{img/Figure2.png}
	\end{tabular}\end{center}
  \end{block}
  \end{column}
\end{columns}

\vfill

\end{frame}

% -----------------------------------------------------------------------------

\end{document}

### 2020-04-03 src/graphs/starGraph.cpp
Create star graphs

@@ -1,0 +1,59 @@
+/*-----------------*
+ * Creates a graph stream for a star graph with a given number of nodes
+ *
+ * Star graphs have one node which is connected to every other node (the other nodes are only conneted to this node)
+ * In a Star Graph of order N, one node has degree N-1 and the remaining 1 have degree 1.
+ *
+ * Creates
+ *    _.edges = Insertion edge stream
+ *    _.vertices = list of vertices and their degrees
+ *-----------------*/
+
+#include <algorithm>
+#include <chrono>
+#include <fstream>
+#include <iostream>
+#include <random>
+#include <string>
+
+using namespace std;
+
+/*-------*
+ *  BODY *
+ *-------*/
+
+int main(int argc, char* argv[]) {
+
+  if (argc!=3) {
+    cout<<"ERROR: ./sg [file_name] [graph_order]"<<endl;
+  } else {
+    string file_name=argv[1];
+    int N=atoi(argv[2]); // Graph degree
+
+    // Prepare files for output
+    ofstream edge_file(file_name+".edges");
+    ofstream vertex_file(file_name+".vertices");
+
+    // Choose node to be centre (Uniformly)
+    default_random_engine generator;
+    generator.seed(chrono::system_clock::now().time_since_epoch().count()); // seed generator with current time
+    uniform_int_distribution<int> uniform_d(1,N);
+    int centre=uniform_d(generator);
+    cout<<"Centre:"<<centre<<endl;
+
+    // create edge between centre and every other node
+    for (int v=1; v<N+1; v++) {
+      if (v!=centre) {
+        edge_file<<centre<<" "<<v<<endl;
+        vertex_file<<v<<" 1"<<endl;
+      } else { // no edge
+        vertex_file<<v<<" "<<N-1<<endl;
+      }
+    }
+
+    edge_file.close();
+    vertex_file.close();
+
+  }
+  return -1;
+}

### 2020-04-03 poster/src/poster.tex
Poster

@@
 	\begin{itemize}
-		\item Instructions in an \textit{Insertion-Only Streams} only ever insert new edges into the graph.
-		\item Instructions in an \textit{Insertion-Deletion Streams} either insert a new edge, or remove an existing edge, from the graph.
+		\item \textit{Insertion-Only Streams} only ever insert new edges into the graph with each instruction.
+		\item \textit{Insertion-Deletion Streams} either insert a new edge, or remove an existing edge, from the graph with each instruction.
 	\end{itemize}
@@
 	\begin{itemize}
-		\item Given a network of social media connections detect popular influencers and analyse the demographics they attract in order to plan targetted advertising campaigns.
+		\item Given a network of social media accounts, detect popular influencers and analyse the demographics they attract in order to plan targetted advertising campaigns.
 		\item Given a log of traffic to a network detect whether a DDOS attack has occurred and, if so, by whom.

### 2020-04-03 src/insertionStreams/insertionStreams.cpp
Display results from insertion only

@@ -42,18 +42,43 @@ int single_pass_insertion_stream(int c, int d, int n,ifstream& stream, vector<ve
 void update_resevoir(vertex n, int d1, int d2, int count, int size, vector<vertex>& resevoir, vector<edge>& edges);
 void parse_edge(string str, edge& e);
 
+void display_results(int c, int d, int n, string file_name);
+
 /*-----*
 * BODY *
 *------*/
 
 int main() {
-  //int d=586, n=747, reps=100;
-  //execute_test(2,20,1,reps,d,n,"../../data/facebook.edges","results/facebook_results.csv");
-  //int d=5948, n=12417, reps=20;
-  //execute_test(2,2,1,reps,d,n,"../../data/gplus.edges","results/gplus_results.csv");
+  //int d=586, n=747, reps=100; int c=3;
+  //display_results(c,d,n,"../../data/facebook.edges");
+  //execute_test(2,100,1,reps,d,n,"../../data/facebook.edges","../../results/facebook_results.csv");
+  int d=5948, n=12417, reps=20; int c=3; // NOTE - # edges=1,179,613
+  display_results(c,d,n,"../../data/gplus.edges");
+  //execute_test(21,100,1,reps,d,n,"../../data/gplus.edges","../../results/gplus_results.csv");
   //int d=104947, n=102100, reps=10; // c=runs, d/c=d2, n=# vertices, NOTE - set d=max degree, n=number of vertices
-  //execute_test(9,17,8,reps,d,n,"../../data/gplus_large.edges","results/gplus_large_results.csv");
+  //execute_test(9,17,8,reps,d,n,"../../data/gplus_large.edges","../../results/gplus_large_results.csv");
   return 0;
+}
+
+void display_results(int c, int d, int n, string file_name) {
+  vector<vertex> neighbourhood; vertex root; // variables for returned values
+  ifstream stream(file_name);
+  BYTES=0; RESEVOIR_BYTES=0; DEGREE_BYTES=0; MAX_BYTES=0; MAX_RESEVOIR_BYTES=0;
+
+  time_point before=chrono::high_resolution_clock::now(); // time before execution
+  int edges_checked=single_pass_insertion_stream(c,d,n,stream,neighbourhood,root);
[…]

### 2020-04-03 src/insertionStreams/naive.cpp

@@ -40,9 +40,9 @@ void parse_edge(string str, edge& e);
  *------*/
 
 int main() {
-  int d=586; int c=20; // d=max_degree,c=accuracy
-  string edge_file_path="../../data/facebook.edges";
-  string output_file_path="../../results/facebook_naive_results.csv";
+  int d=5948; int c=20; // d=max_degree,c=accuracy
+  string edge_file_path="../../data/gplus.edges";
+  string output_file_path="../../results/gplus_naive_results.csv";
 
   execute_test(2,100,1,d,edge_file_path,output_file_path);
 
@@ -57,6 +57,7 @@ void execute_test(int c_min, int c_max, int c_step, int d, string in_file, strin
 
   vertex root; set<vertex> neighbourhood;
   for (int c=c_min;c<=c_max;c+=c_step) {
+    cout<<c<<endl;
     BYTES=0;
     vertex* p=&root; p=nullptr;
     neighbourhood.clear();

### 2020-04-03 src/utility.cpp
list_vertices now outputs as a CSV file

@@ -123,24 +123,38 @@ void generate_insertion_deletion(string file_name) {
   outfile.close();
 }
 
+// produces list of vertices & their degree
+// works for insetion & insertion-deletion streams
 void list_vertices(string file_name) {
-  ofstream outfile(file_name+".vertices");
   ifstream stream(file_name+".edges");
+  ofstream outfile(file_name+".vertices");
 
-  set<string> vertices; string line; edge e;
+  map<string,int> degrees; string line; edge e;
+  int i=0;
 
-  while(getline(stream,line)) {
+  while (getline(stream,line)) {
+    i+=1;
+    if (i%100000==0) cout<<i<<","; // update on how many edges have been checked
     parse_edge(line,e);
 
-    if (vertices.find(e.fst)==vertices.end()) {
-      vertices.insert(e.fst);
-      outfile<<e.fst<<endl;
+    if (degrees.count(e.fst)) { // vertex already in graph
+      if (e.insertion) degrees[e.fst]+=1; // insertion edge
+      else degrees[e.fst]-=1; // deletion edge
+    } else { // vertex not in graph
+      degrees[e.fst]=1; // cannot delete an edge which is not in the graph
     }
 
-    if (vertices.find(e.snd)==vertices.end()) {
-      vertices.insert(e.snd);
-      outfile<<e.snd<<endl;
+    if (degrees.count(e.snd)) { // vertex already in graph
+      if (e.insertion) degrees[e.snd]+=1; // insertion edge
+      else degrees[e.snd]-=1; // deletion edge
+    } else { // vertex not in graph
+      degrees[e.snd]=1; // cannot delete an edge which is not in the graph
     }
+  }
+
+  // write to vertices file
+  for (map<string,int>::iterator i=degrees.begin(); i!=degrees.end(); i++) {
+    outfile<<i->first<<","<<i->second<<endl; // name,degree
   }
 }
 

### 2020-04-03 src/graphs/completeGraph.cpp
Create complete graphs

@@ -1,0 +1,47 @@
+/*-----------------*
+ * Creates an insertion graph stream for a complete graph with a given number of nodes
+ *
+ * Complete graphs contain an edge between each vertex
+ * In a Star Graph of order N, every node has degree N-1
+ *
+ * Creates
+ *    _.edges = Insertion edge stream (formatted as "v1 v2")
+ *    _.vertices = list of vertices and their degrees (formatted as CSV file)
+ *-----------------*/
+
+#include <algorithm>
+#include <fstream>
+#include <iostream>
+#include <string>
+
+using namespace std;
+
+/*-------*
+ *  BODY *
+ *-------*/
+
+int main(int argc, char* argv[]) {
+
+  if (argc!=3) {
+    cout<<"ERROR: ./sg [file_name] [graph_order]"<<endl;
+  } else {
+    string file_name=argv[1];
+    int N=atoi(argv[2]); // Graph degree
+
+    // Prepare files for output
+    ofstream edge_file(file_name+".edges");
+    ofstream vertex_file(file_name+".vertices");
+
+
+    for (int v=1; v<N+1; v++) {
+      vertex_file<<v<<","<<N-1<<endl;
+      for (int u=v+1; u<N+1; u++) edge_file<<v<<" "<<u<<endl;
+    }
+
+    edge_file.close();
+    vertex_file.close();
+
+  }
+
+  return 0;
+}

### 2020-04-03 src/graphs/starGraph.cpp

@@ -1,12 +1,12 @@
 /*-----------------*
- * Creates a graph stream for a star graph with a given number of nodes
+ * Creates an insertion graph stream for a star graph with a given number of nodes
  *
  * Star graphs have one node which is connected to every other node (the other nodes are only conneted to this node)
  * In a Star Graph of order N, one node has degree N-1 and the remaining 1 have degree 1.
  *
  * Creates
- *    _.edges = Insertion edge stream
- *    _.vertices = list of vertices and their degrees
+ *    _.edges = Insertion edge stream (formatted as "v1 v2")
+ *    _.vertices = list of vertices and their degrees (formatted as CSV file)
  *-----------------*/
 
 #include <algorithm>
@@ -45,9 +45,9 @@ int main(int argc, char* argv[]) {
     for (int v=1; v<N+1; v++) {
       if (v!=centre) {
         edge_file<<centre<<" "<<v<<endl;
-        vertex_file<<v<<" 1"<<endl;
+        vertex_file<<v<<",1"<<endl;
       } else { // no edge
-        vertex_file<<v<<" "<<N-1<<endl;
+        vertex_file<<v<<","<<N-1<<endl;
       }
     }
 
@@ -55,5 +55,5 @@ int main(int argc, char* argv[]) {
     vertex_file.close();
 
   }
-  return -1;
+  return 0;
 }

### 2020-04-03 src/graphs/randomGraph.cpp
Create relatively random graphs

@@ -1,0 +1,68 @@
+/*
+ * Creates an insertion graph stream for a graph with a defined number of edges & vertices
+ *
+ * Models creating the upper right triangle of a matrix which represents the edges in the graph
+ *
+ * Creates
+ *    _.edges = Insertion edge stream (formatted as "v1 v2")
+ *    _.vertices = list of vertices and their degrees (formatted as CSV file)
+ */
+
+#include <algorithm>
+#include <chrono>
+#include <fstream>
+#include <iostream>
+#include <map>
+#include <random>
+#include <string>
+#include <vector>
+
+using namespace std;
+
+/*-------*
+ *  BODY *
+ *-------*/
+
+int main(int argc, char* argv[]) {
+
+  if (argc!=3 && argc!=4) {
+    cout<<"ERROR: ./sg [file_name] [graph_order] (prob_edge)"<<endl;
+  } else {
+    string file_name=argv[1];
+    int N=atoi(argv[2]); // Graph degree
+
+    // Prepare files for output
+    ofstream edge_file(file_name+".edges");
+    ofstream vertex_file(file_name+".vertices");
+
+    default_random_engine generator;
+    generator.seed(chrono::system_clock::now().time_since_epoch().count()); // seed with current time
+
+    map<int,int> degrees;
+    for (int i=1; i<N; i++) degrees[i]=0; // record degree of every node
+
+    // bernoulli_d(generator);
+    for (int i=1; i<N+1; i++) {
+      double P;
+      if (argc==4) { // specified probability
+        P=atof(argv[3]); // Probs of making an edge
+      } else { // randomly choose probability
+        uniform_int_distribution<int> uniform_d(1,100);
+        P=(double)uniform_d(generator)/100;
+      }
+      bernoulli_distribution bernoulli_d(P);
+
+      for (int j=i; j<N+1; j++) {
+        if (j!=i && bernoulli_d(generator)==1) {
+          degrees[i]+=1; degrees[j]+=1; // update degrees
+          edge_file<<i<<" "<<j<<endl;
+      }}
+      vertex_file<<i<<","<<degrees[i]<<endl;
+    }
+
+    edge_file.close();
+    vertex_file.close();
+  }
+
+  return 0;
+}

### 2020-04-04 src/graphs/randomGraph.cpp
Random Insertion-Deletion graphs

@@ -1,9 +1,23 @@
 /*
- * Creates an insertion graph stream for a graph with a defined number of edges & vertices
+ * Creates insertion-only or insertion-deletion graph streams
  *
- * Models creating the upper right triangle of a matrix which represents the edges in the graph
+ * USING - ./rg [file_name] [I/ID] [graph_order] (prob_edge)
+ *  file_name   = name of .edges & .vertices file to output (can be a path)
+ *  I/ID        = Insertion-only or Insertion-Deletion
+ *  graph_order = number of vertices in graph
+ *  prob_edge   = probability of an edge (insertion-only & optional)
  *
- * Creates
+ * INSERTION ONLY
+ *  If prob_edge is _not_ specified then for each vertex U[0,1] is sampled to define its probability of having an edge to another arbitrary vertex
+ *  Otherwise, all vertices have the same probability (prob_edge).
+ *  This probability is used in a Bernoulli distribution to decide whether to add an edge.
+ *
+ * INSERTION-DELETION
+ *  Each vertex samples from Poi(0.8), this value is then used as mean for another Poisson distribution.
+ *  The second distribution is sampled to decide the number of edges between this vertex and another one.
+ *  Edges added as Insertion then deletion until the required number of edges is produced
+ *
+ * Returns
  *    _.edges = Insertion edge stream (formatted as "v1 v2")
  *    _.vertices = list of vertices and their degrees (formatted as CSV file)
  */
@@ -19,50 +33,112 @@
 
 using namespace std;
 
+/*-------------*
+ *  SIGNATURES *
+ *-------------*/
+
+// p=NULL to randomly generate
+void insertion_only(ofstream& edge_file, ofstream& vertex_file, int num_vertices, double p, default_random_engine generator);
+void insertion_deletion(ofstream& edge_file, ofstream& vertex_file, int num_vertices, default_random_engine generator);
 /*-------*
  *  BODY *
  *-------*/
 
 int main(int argc, char* argv[]) {
 
-  if (argc!=3 && argc!=4) {
-    cout<<"ERROR: ./sg [file_name] [graph_order] (prob_edge)"<<endl;
+  if (argc!=4 && argc!=5) {
[…]

### 2020-04-04 src/insertionDeletionStreams/insertionDeletionStreams.cpp

@@ -1,368 +1,0 @@
-/*-----------------*
- * One pass c-approximation Streaming Algorithm for Neighbourhood Detection for Insertion-Only Graph Streams
- * NOTE - I have cheated by making bidirectional maps between vectors & indicies but only counting one of them
- * TODO
- *  c has to be quite large?? st (|A'|<=|A|)
- *  For all c where n/c < sqrt{n} surely space required is ~ equal??
- *-----------------*/
-
-#include <algorithm>
-#include <chrono>
-#include <cstring>
-#include <fstream>
-#include <iostream>
-#include <map>
-#include <math.h>
-#include <random>
-#include <string>
-#include <set>
-
-using namespace std;
-
-int P=1073741789; // >2^30
-int BYTES;
-int MAX_BYTES;
-
-/*-----------------*
- * DATA STRUCTURES *
- *-----------------*/
-
-using vertex = string; // typemap vertex
-using time_point=chrono::high_resolution_clock::time_point;
-
-struct edge { // undirected edge
-   vertex fst;
-   vertex snd;
-   bool insertion;
-};
-
-struct hash_params { // parameters for hash function
-  unsigned long a;
-  unsigned long b;
-  unsigned long m;
-};
-
-/*-----------*
- * SIGNATURES *
- *------------*/
-
-//  main algorithm
-int vertex_sampling(string edge_file_path, string vertex_file_path, int num_vertices, int c, int d, set<vertex> &neighbourhood, vertex& root);
-void sample_vertices(string vertex_file_path, int num_vertices, int sample_size, set<vertex>& vertex_sample, set<vertex>& vertex_not_in_sample);
-void index_vertices(set<vertex> vertex_sample, set<vertex> vertex_not_in_sample, map<vertex,int>& vertex_to_index, map<int,vertex>& index_to_vertex);
-map<vertex,bool*> initialise_edge_vectors(set<vertex> vertex_sample, int num_vertices);
-void construct_edge_vectors(string edge_file_path, set<vertex> vertex_sample, int num_vertices, map<vertex,int> vertex_to_index, map<vertex,bool*>& edge_vectors);
[…]

### 2020-04-04 src/insertionDeletionStreams/l0Samplers/l0Sampler.cpp
L0 Sampler Implementation - Unique random hash map works better.

@@ -1,0 +1,335 @@
+/*
+ *  An implementation of an L0-Sampler for a specified vertex in a graph stream.
+ *    This samples uniformly from the neighbourhood of the vertex.
+ *    Edges are provided as a stream of inputs
+ *  TODO
+      - Appears to fail randomly after some runs?? Investigate
+ */
+
+#include <algorithm>
+#include <chrono>
+#include <iostream>
+#include <fstream>
+#include <map>
+#include <random>
+#include <set>
+#include <string>
+#include <vector>
+
+using namespace std;
+
+int P=1073741789; // >2^30
+
+/*-----------------*
+ * DATA STRUCTURES *
+ *-----------------*/
+using vertex = int; // typemap vertex
+
+struct edge { // undirected edge
+    vertex fst;
+    vertex snd;
+    int value;
+  };
+
+struct hash_params { // parameters for hash function
+  unsigned long a;
+  unsigned long b;
+  unsigned long m;
+};
+
+/*------------*
+ * SIGNATURES *
+ *------------*/
+
+ // 1-sparse
+ bool verify_1_sparse(int phi,int iota,int tau);
+ void update_1_sparse_counters(int index,int delta,int row,int col,long** phi_s,long** iota_s,long** tau_s);
+
+
+ // Hashing
+ hash_params generate_hash(int m);
+ int hash_function(int key, hash_params ps);
+
+// Utility
+void parse_edge(string str, edge& e);
+int identify_endpoint(edge e,vertex target);
+long*** initalise_zero_3d_array(int depth, int num_cols, int num_rows);
+hash_params** initalise_2d_hash_params_array(int num_cols, int num_rows);
+void free_3d_long_array(long*** arr, int depth, int num_cols, int num_rows);
+void free_2d_hash_params_array(hash_params** arr, int num_cols, int num_rows);
+
+/*------*
+ * BODY *
+ *------*/
+
+int main() {
+  // details of graph to perform on
+  string file_path="../../../data/facebook_deletion.edges";
+  vertex target=2290;
+  int num_vertices=747;
+
+  // constraints on s-sparse recovery
+  double delta=.01; // acceptable failure for L0
+  double gamma=.01; // acceptable failure for s-sparse recovery
[…]

### 2020-04-04 src/insertionDeletionStreams/l0Samplers/l0SamplerRandomHash.cpp

@@ -1,0 +1,367 @@
+/*
+ *  TEST DESCRIPTIOn
+ *    In this implementation the hash function assigns values uniformly at random to a unique value.
+ *    This is inefficient as it has to store where every value goes but it does mean that it is k-wise independent.
+ *    This is to test whether using a pairwise independent hash function was the problem.
+ *    If this comes out as uniform, then the pairwise independent hash function was not sufficient.
+ *    NOTE this only concerns the hash function used to choose whether to pass update to an s-sparse recovery
+ *        not the one used during s-sparse recovery
+ *
+ *  An implementation of an L0-Sampler for a specified vertex in a graph stream.
+ *    This samples uniformly from the neighbourhood of the vertex.
+ *    Edges are provided as a stream of inputs
+ *  TODO
+      - Appears to fail randomly after some runs?? Investigate
+ */
+
+#include <algorithm>
+#include <chrono>
+#include <iostream>
+#include <fstream>
+#include <map>
+#include <random>
+#include <set>
+#include <string>
+#include <vector>
+
+using namespace std;
+
+int P=1073741789; // >2^30
+
+/*-----------------*
+ * DATA STRUCTURES *
+ *-----------------*/
+using vertex = int; // typemap vertex
+
+struct edge { // undirected edge
+    vertex fst;
+    vertex snd;
+    int value;
+  };
+
+struct hash_params { // parameters for hash function
+  unsigned long a;
+  unsigned long b;
+  unsigned long m;
+};
+
+/*------------*
+ * SIGNATURES *
+ *------------*/
+
+// 1-sparse
+bool verify_1_sparse(int phi,int iota,int tau);
+void update_1_sparse_counters(int index,int delta,int row,int col,long** phi_s,long** iota_s,long** tau_s);
+
+
+// Hashing
+hash_params generate_hash(int m);
+int hash_function(int key, hash_params ps);
+// n = num keys, m=possible hash values, hash=map from key to hash value
+void generate_random_hash(int n, int m, map<int,int>& hash);
+
+// Utility
+void parse_edge(string str, edge& e);
+int identify_endpoint(edge e,vertex target);
[…]

### 2020-04-04 src/insertionDeletionStreams/l0Samplers/sSparseRecovery.cpp

@@ -59,6 +59,10 @@ void parse_edge(string str, edge& e);
 int identify_endpoint(edge e,vertex target);
 long** initalise_zero_2d_array(int num_cols, int num_rows);
 
+/*------*
+ * BODY *
+ *------*/
+
 int main() {
   int s=100; // sparsity to recover
   double delta=.1; // acceptable failure
@@ -77,8 +81,8 @@ int main() {
  *-------------------*/
 
 set<vertex> s_sparse_recovery(string edge_file, vertex target, int num_vertices, int s, double delta) {
-  int num_rows=log(s/delta);
   int num_cols=2*s;
+  int num_rows=log(s/delta);
 
   // arrays for 1-sparse recovery
   long** phi_s =initalise_zero_2d_array(num_cols,num_rows); // sum of weights (sum ai)

### 2020-04-19 paper/src/paper.tex
Insertion-Only write up mostly done

--- 2020-02-06 paper/src/paper.tex
@@
 \documentclass[11pt,twoside,a4paper]{report}
+
+\usepackage[margin=1in, paperwidth=8.3in, paperheight=11.7in]{geometry}
 
@@
 \usepackage{changepage}
+\usepackage{enumerate}
+\usepackage{enumitem}
 \usepackage{fancyhdr} % Header
+\usepackage{graphicx}
 
@@
 \newcommand{\nats}{\mathbb{N} }
-\newcommand{\horizontalline}{\newline\vspace{.3cm}\hfill\makebox[.5\linewidth]{\rule{.5\textwidth}{0.4pt}}\hfill\vspace{.1cm}}
-% title
-\title{Implementing \& Evaluating Space Efficient Algoirthms for Detecting Large Neighbourhoods in Graph Streams}
-\author{Dom Hutchinson}
-\date{\today}
-\maketitle
+\newcommand{\horizontalline}{\newline\vspace{.3cm}\hfill\makebox[.5\linewidth]{\rule{.5\textwidth}{0.4pt}}\hfill\vspace{.05cm}}
+
+\renewcommand\thetable{\arabic{table}} % figure numbering strategy
+\renewcommand\thefigure{\arabic{figure}} % figure numbering strategy
+
+\pagenumbering{roman}
 
@@
 
-\chapter{Background}
+% Algorithm environment
+\SetArgSty{textnormal}
+\DontPrintSemicolon
 
-\section{Motivation}
-\par For most algorithms time efficiency is viewed with primary importance, rather than space efficiency.
-However if the memory used for an algorithm exceeds the size of the computer's RAM then virtual memory would have be used.
-Virtual memory has a much longer read/write than RAM meaning that the algorithm's run-time will be much greater in this scenario.
-Thus for algorithms which are designed to work with large data sets space efficiency is more important.
-\par In the early 2000s companies began to realise that data could be used as a commodity, and with this the amount of data being collected has surged in the years since.
-In this same time period we have nots seen the same surge in the amount of RAM typical computers have.
-Thus people wishing to capitalise on this increase in available data need to give greather thought to the space efficiency of the algorithms they are using.
-Otherwise run-times could become unusuable.
-\par Social networks are good sources of data about people.
-This data can leveraged in many ways, notably within advertising.
-Behaviourly targeted adverts have been shown to have more than twice the click through rate of standard online advertiesments \cite{targetedAds}.
-The last decade has seen developments in the space of targeted advertising, one of which is the rise of social media influencers.
-These are people who are deemed to have a large online following, which they are able to influence.
-Advertisers will often approach these influencers with deals by which the influencer will promote a product to their audience, hopefully driving sales.
-\par The question becomes, ``Who is an influencer \& how do we find them?''.
-This question can be generalised as the $\mathtt{Neighborhood\ Detection\ Problem}$.
-Which is the problem I shall discuss in this paper.
-%TODO
-Although in my description of the Neighbourhood Detection Problem I validate its relevance by referring to find social media influencers, it is applicable to many other scenarios.
+% Title
+\title{Implementing \& Evaluating Space Efficient Algorithms for Detecting Large Neighbourhoods in Graph Streams}
+\author{Dom Hutchinson}
+\date{\today}
+\maketitle
+\newpage
+
+\tableofcontents
+\newpage
+
+\renewcommand\thechapter{\Roman{chapter}}
+\renewcommand\thesection{\thechapter.\roman{section}}
+\setcounter{chapter}{1}
+\chapter*{Summary/Abstract}
+\addcontentsline{toc}{chapter}{Summary/Abstract}
+
+\section{Background}
+
+%\begin{itemize}
+%	\item High Degree Detection.
+%	\item Large Neighbourhood Detection.
+%\end{itemize}
+
+%TODO testing space usage is important as it suggests what will happen when larger graphs are used (the longer we can keep space requirements below RAM space the better).
+
+ % TODO is it ok to define streams & such later
+Graphs are a popular way of representing relationships between objects.
+They are particularly popular within the field of data science \& with the recent movement of \textit{Big Data} it is becoming common to want to deal with large graphs.
+Moreover, it is common to want to use graphs which cannot be loaded fully onto a computer's RAM and instead require the use of virtual memory.
+Having to use virtual memory drastically increases request time for data \& thus negatively affects the runtime of algorithms which manipulate these large graphs.
+Often, in practice, the fastest algorithms for analysing graphs are those which are more space efficient.
+%rather than those that, in theory, have lower run-time complexity
+\par Graph streams have been introduced to reduce the space required to store a graph since they require $O(|E|)$ space, rather than $O(|V|^2)$ space which a traditional matrix representation requires.
+% TODO define |E| & |V| as edge and vertex sets
+Since the number of edges in a graph is cannot be greater than, and is usually much lower than, the square of the number of vertices this is a big improvement.
+However many graphs still exist which would overflow RAM when loaded in full when in stream form.
+This is solved by reading the stream one entry at a time as this requires $O(1)$ space.
+Thus the most space-efficient algorithms for analysing graphs are streaming-algorithms which read instructions sequentially.
+\par Graph streams can be used to show the evolution of a graph by insertion \& deleting edges over some time period, in these implementations it is possible for a graph stream to require more space than a matrix representation but since only one instruction is every loaded into the RAM at one time this does not affect the total memory used by a streaming-algorithm.
+%TODO this might seem out of place (too much detail too early) but I want to make the point that O(|E|) can be greater than O(|V^2|) and this doesn't matter
+% TODO perhaps mention that graph streams can show the evolution of a graph (in this case they may have more entries than |V|^2 but doesn't matter since only one entry is every read at a time).
+% TODO why single pass??
+\horizontalline
+\par High degree detection is a common and trivial problem in graph theory.
+The \texttt{High Degree Detection Problem} entails finding a vertex in a graph with sufficiently high degree.
+This can be trivially solved for graph streams by keeping counters for the number of edges connected to each vertex and updating the counters after each instruction from the graph stream.
+% This requires $O(|V|)$ space
+\par Although it is useful to identify influential vertices in a graph, it is hard to fully utilise this knowledge without additional information about the neighbourhood of these vertices.
+Knowing the neighbourhood of a vertex, or at least a subset of it, allows you to make inferences about that node.
+This leads to the \texttt{Neighbourhood Detection Problem} which is formally stated as
+
+\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+\textbf{Problem 1} \texttt{Neighbourhood Detection}.\\
+Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A,B$, where $|A|=n$ and $|B|=\text{poly }n$, and edge-set $E$.\\
+In $\texttt{Neighbourhood Detection}(G,d,c)$ we are tasked with outputting a vertex from $A$ with at least $d/c$ of its neighbours in $B$.
+We can assume that $G$ contains at least one node of degree $d$.\\
+Here $d\in\nats$ is a threshold parameter \& $c\in(0,1]$ is an approximation parameter.
+}}\end{adjustwidth}\vspace{.3cm}
+
+The \texttt{Neighbourhood Detection Problem} is defined for bi-partite graphs in order to generalise the problem.
+Note that, if we were to define the vertex sets $A,B$ to be equivalent then then the problem holds for every type of graph.
+\par Again, the \texttt{Neighbourhood Detection Problem} can be solved trivially by storing the neighbourhoods of every vertex in the graph \& updating them after each instruction from the graph stream.
+This requires $O(|E|)$ space and can very quickly lead to overflowing the RAM which is undesirable since it means slower run-times for large graphs.
+
+\section{Motivating Applications} %TODO I don't really like this (partly because it is in list form), moreover it isn't really connected in
+% NOTE maybe it is best that it is fairly separate from the rest so that it can be skipped over.
+%\begin{itemize}
+%	\item Real world occurrences \& uses.
+%\end{itemize}
+
+There are several real-world applications of the \texttt{Neighbourhood Detection Problem} which motivate investigating it.
+These are situations where simply knowing highly connected elements of a network is not sufficient.
 \begin{itemize}
-	\item[-] TODO
+	\item[-] Given a list of connections within a social network, solving the problem would allow you to not only find \textit{`social influencers'} but those they are connect to too.
+This allows you to verify that they are indeed a \textit{`social influencers'}, and not someone connected to a lot of bot accounts, and to assess what demographics they have influence over.\\
+	For a marketing agency this would mean they don't waste money on people who don't actually have any influence to sell a product, nor on people who have the wrong demographics for a certain product.
+	\item[-] Given a list of passengers on public transport you could find people who have come into contact with a lot of people.
+This is particular relevant in the wake of the COVID-19 outbreak as it could be used to identify people who are likely to have come into contact with infected people.
+%TODO not a fan of these, might just need fleshing out
+	\item[-] Given a list of receipts in a shop you could find the items which are popular, along with which items they are commonly bought with.
+This information could be used to inform how shelves are stacked and which promotions are run by the store, in order to encourage sales.
+	\item[-] Given a log of traffic within a network you could identify which resources are being accessed most often, and by whom.
+This can be used to inform which resources should be upgraded \& which connections should be prioritised.
 \end{itemize}
-\section{Neighbourhood Detection Problem}
-%TODO Describe Graph streams, insertion-only & insertion-deletion
-\par Graph streams are a dynamic way of representing graphs.
-They are simply a list of instructions for how to construct a graph and in this paper I shall discuss two types: insertion-only streams \& insertion-deletion streams.
-There are other types which allow for updating vertex values, among other properties, but they are not relevant to the problem we are considering in this paper.
-\par Insertion-only streams are a list of pairs of vertices, representing endpoints of edges.
-Each time we read an entry from an insertion-only stream we are adding the edge it describes to the graph.
-We apply a limitation that no duplicate edegs appear in an insertion-only stream and thus the order of the stream has no affect on the validity of the algorithms which shall be discussed.
-\par Entries in an insertion-deletion list specify a pair of vertices, representing an edge, and a boolean which states whether this is an insertion or deletion entry.
-If it is an insertion entry then we are adding the described edge to the graph; if it is a deletion entry then we are removing it.
-Again we add a limitation that no edges are repeated \& that we can not recieve a request to delete an edge from the graph which does not currently exist in the graph.
-This does however mean that the order of an insertion-deletion is important as the deletion of an edge must come after it.
-\par Despite these limitations it is still possible to construct almost any type of graph from a graph stream.
-Suppose you wish to analyse which devices in a network are making the most request.
-Here vertices represent devices \& edges represent requests.
-In this scenario you would want to allow for multiple edges (say $n>2)$ between the same pair of devices (say $\{x,y\}$).
-One solution would be to add $n$ new vertices $\{v_{x,y,1},\dots,v_{x,y,n}\}$ each with a label stating they are just indetermiate nodes and not full devices.
-Creating edges from $x$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ \& from $y$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ allows for an interpretation of multiple edges.
-\par If a graph stream is ordered by time (which is common in real life applications) then it represents a dynamic-evolution of the graph over time.
-Allow the user to roll back the graph should they wish to make assessments during particular time preiods, rather than just assess the final graph.
-\par The algorithms which shall be discussed in this paper are designed to work with just a single pass of a graph stream.
-This is not only time efficient as it does not require multiple reads of the stream, but means that should more information be gained after the algorithm is run it is easy to incorporate it into algorithm if we store the final state of the data structures used.
-Single pass algorithms mean we don't necessarily have to store the graph stream, we could simply send entries from the system being modelled straight into the algorithm, without being stored in a file.
-This is useful when a graph stream potentially could take up multiple terra-bytes, but is unnecessary in most applications.
+
+\section{Structure} %TODO after most the write up
+In this paper I am going to implement the two algorithms proposed by Dr Christian Konrad in \cite{orig} for solving the \texttt{Neighbourhood Detection Problem} for graphs streams in a space-efficient manner.
+I will use this implementation to assess how practical these algorithms are and now to maximise their practicalities.
+\par I shall structure this investigation as follows
+
+\begin{itemize}
+	\item Insertion-Only Streams
+	\begin{itemize}
+		\item Layout theory
+		\item Discuss features that need to be/can be adjusted in the implementation.
+		\item Implement - mentioning any other problems.
+		\item Evaluate \& test different combinations of parameters to optimise the practical uses of the algorithm.
+		\item Discuss limitations of the implementation. (ie What graphs work)
+	\end{itemize}
+	\item Repeat for Insertion Deletion.
+\end{itemize}
+
+\section{Objectives}
+The objectives of this project are
+\begin{itemize}
+	\item Implement Christian's algorithms and discuss any considerations that need to be taken in doing so.
+	\item Tune parameters to make algorithms as practical as possible.
+	\item How practical are the algorithms.
+\end{itemize}
+
+\renewcommand\thechapter{\arabic{chapter}}
+\renewcommand\thesection{\thechapter.\arabic{section}}
+\setcounter{chapter}{0}
+\chapter{Preliminaries}
+
+\setcounter{page}{1} % restart page counter
+\pagenumbering{arabic}
+
+\section{Definitions}
+
+%\begin{itemize}
+%	\item Graph
+% \item Unweighted, undirected
+%	\item Bipartite Graph (Why this wasn't really applied in practice, A=B)
+%	\item Graph Stream (Insertion-Only, Insertion-Deletion)
+%	\item Neighbourhood, degree & sparsity of a vertex
+%	\item Streaming Algorithm
+%	\item $1-\varepsilon$ Approximation Algorithm
+%	\item TODO Space efficiency??
+%\end{itemize}
+
+A \textit{Graph} is a data structure used to represent pairwise relationships between objects.
+A Graph is defined to have a vertex set, which holds its objects, and an edge-set, which holds the relationship between these objects.
+Graphs are traditionally visualised with circles for each vertex and lines between vertices that share an edge.
+There are variations on graphs where the edges are weighted and/or directional.
+In this project we are only required to deal with undirected, unweighted graphs.
+\par A \textit{Bipartite Graph} is a graph whose vertex set can be partitioned into two distinct sets in such a way that all edges in its edge-set run between these partitions \& not within them.
+% Formally, a graph $G=(V,E)$ where $\exists\ A,B\subset V$ st $V=A\cup B,\ A\cap B=\emptyset$ and $\forall\ (u,v)\in E,\ u\in A\ \&\ v\in B$.
+\par The \textit{Neighbourhood} of a vertex in a graph is the set of vertices which it shares an edge with.
+The \textit{Degree} of a vertex is the size of its neighbourhood.
+A vertex is $s$ \textit{Sparse} if its degree is less than, or equal to, $s$.
+\par A \textit{Graph Stream} is a technique for storing graphs.
+A graph stream is made up of a series of instructions which are read sequentially and describe how to construct the graph.
+These instructions describe modifications to the edges of the graph.
+In this project we are only concerned with two types of graph stream.
+\begin{itemize}
+	\item \textit{Insertion-Only Streams} where each instruction adds a new edge to the graph.
+	\item \textit{Insertion-Deletion Streams} where each instruction either adds a new edge to the graph or removes an existing edge from the graph.
+\end{itemize}
+Neither of these allow for duplicates of the same edge in the graph, at the same time.
+Insertion-Deletion streams are used for graphs whose edge-set is dynamic over time, consider a social network where users are able to follow \& un-follow each other.
+There are other versions of graph streams which all for changes to the weights of edges but these are out of scope for this project.
+It is noteworthy that insertion-only streams equivalent to insertion-deletion streams which have no deletions.
+This means that they can be implemented in exactly the same way.
+% TODO maybe state that these are equivalent to the above if we treat 0 weight to be no edge & anything else to be an edge %TODO maybe mention when order is important (or can represent time)
+\par \textit{Streaming Algorithms} are algorithms which are designed to take a stream of sequential instructions as their input.
+Streaming algorithms can be very space-efficient since only one instruction is read at a time, requiring constant space.
+Typically streaming algorithms aim to only require a single pass of the input stream so that you can extend the input stream without having to restart the algorithm.
+Single-pass streaming algorithms are online algorithms.
+%TODO make the mention of online algorithms more succinct
+\par An \textit{$\alpha$-Approximation Algorithm} is an algorithm which returns a result within an $\alpha$ factor to the optimal solution to the problem.
+Algorithms which solve the \texttt{Neighbourhood Detection Problem} are $\frac1c$-approximation algorithms since they only return $d/c$ of the neighbours of a vertex which we know to have at least $d$ neighbours.
+
+\section{Technologies}
+%\begin{itemize}
+%	\item C++
+%	\begin{itemize}
+%		\item Low level
+%		\item Memory manipulation
+%		\item Faster
+%		\item Perhaps something about why not java or python etc.
+%	\end{itemize}
+%	\item Standard library online
+%\end{itemize}
+For the implementations in this project I chose to use C++.
+I initially considered using either python or C++, as they work in the object-orientated paradigm and I am proficient in them.
+I chose to go with C++ (specifically \texttt{C++11}) as it allows for more control with memory management and typically faster run-times.
+For this project the memory management is more important as I am looking to evaluate the space-efficiency of algorithms and being able to control their memory allocation means I can identify each time a change occurs to the amount of memory being used.
+The faster run-times is important when evaluating the practicalities of using these algorithms, as longer run-times make algorithms less practical.
+\par I limited myself to using the \textit{C++ Standard Library} (\texttt{std}) as the underlying implementations of this library are well document.
+This was particularly important for the abstract data types I used as I could check what the underlying implementations and adjust my evaluation accordingly.
+When implementing randomness I used the \texttt{<random>} module of \texttt{std} and seeded the generators with the current time so that each run would have a different generator.
+
+\section{Evaluation Techniques}
+
+% TODO expectations
+%\begin{itemize}
+%	\item What features are tested (time \& space)? How?
+%	\item Against naïve algorithm.
+%	\item Insertion-Only v Insertion-Deletion algorithm for insertion-only streams.
+%\end{itemize}
+
+%TODO should this be in the future tense (eg in this paper i will)
+%TODO maybe keep this more vague and move the details to their respective sections
+%TODO talk about verifying my results
+%TODO typical test = vary c between [3,20]. set d=max degree
+
+When evaluating these algorithms there were three main parameters which needed to be tested for: success-rate; time-taken; and, space-used.
+These are all interconnected and come with their own trade-offs.
+Typically, the changes to space-used \& time-taken when different strategies are implemented occur in the same direction but at different rates.
+Whereas, you expect success-rate to go in the opposite direction as the algorithm is using fewer resources.
+For these particular algorithms the success rate always remained high but for some setups it takes a lot longer to find a solution. %TODO do I means this??
+\par The evaluation process of each algorithm can be broken into two stages.
+First, I evaluated possible efficiencies to the underlying implementation.
+The changes tested at this stage did not affect the success-rate of the algorithms but did, in some cases, have a drastic effect on the amount of resources used by each algorithm.
+Second, I evaluated different set-ups for the parameters of the algorithm.
+This stage involved a lot more \textit{tuning} of values \& strategies as these affected the logic of the algorithm resulting in trade-offs between resource requirements \& success-rate.
+The changes in the first were a lot more general \& there was little variation in the results seen for different graph types, whereas in the later stage I had to consider different graph types.
+\par I used a global variable to count the space used by different components of the algorithms (as well as the algorithm as a whole) and a timer to evaluate the run-time.
+For each variation I ran the tests multiple times, recording the results of each run.
+Typically for testing I set $d$ to be the maximum degree of the graph as these algorithms are typically used to investigate the nodes of greatest degrees in a graph, and I would then vary the approximation factor $c$ as seem reasonable for the given graph.
+
+\subsection{Generating Graphs}
+
+%\begin{itemize}
+%	\item How graphs streams are represented.
+%	\item Different graph types.
+%	\item Generating graphs
+%	\item Hard to have wide variety of node degrees.
+%	\item Acquired from SNAP.
+% \item Graph pre-processing (.vertices file).
+%	\item Programs to verify success (ie returned neighbourhood is valid).
+%\end{itemize}
+For this project I required a number of graphs to evaluate my programs with.
+By the nature of the algorithms these graphs had to be undirected \& unweighted bi-partite graphs.
+The bi-partite requirement is unimportant as any graph can be treated as bi-partite by making the vertex set partitions copies of each other.
+\par I implemented programs to generate graphs with specific properties which I wanted to test, namely graphs with high disparity in the degree of each node (such as star graphs) and graphs with no disparity between degrees (such as complete graphs).
+% TODO more justification
+I implemented the following programs.
+\begin{itemize}
+	\item \texttt{completeGraph.cpp} - Generates a graph stream for a complete graph.
+	\item \texttt{randomGraph.cpp} - Uses a Bernoulli distribution to determine whether two vertices share an edge in an insertion-only stream, and a poisson distribution to determine how many insertions \& deletions occur between each pair of vertices.
+	\item \texttt{starGraph.cpp} - Generates a graph stream for a star graph.
+\end{itemize}
+All these programs take parameters for the number of vertices and \texttt{randomGraph.cpp} takes parameters for the proportion of edges.
+Additionally these, programs are able to generate insertion-only or insertion-deletion streams but the final graphs defined by the streams generated by \texttt{completeGraph.cpp} \& \texttt{starGraph.cpp} are always of the appropriate shape.
+%TODO reword this
+%TODO I found it hard to have a wide variety of node degrees
+\par I wanted some graphs generated from the real-world so that I could test the algorithms on the sort of data sets they are most likely to be used on.
+I found a collection of large graph streams, from different applications, in the \texttt{Stanford Network Analysis Project} (\texttt{SNAP}) datasets collection \cite{SNAP}.
+%TODO I chose social network ones??
+The vertices in the instructions of these graph streams were not ordered meaning that these graphs are not strictly bi-partite. % TODO better explanation?
+This meant I needed to change my implementations slightly such that both ends of the edge are treated as $A$-vertices.
+This is equivalent to having a bi-partite graph where the $A$ \& $B$ vertex sets are copies of each other, which is perfectly fine.
+\par The graphs from \texttt{SNAP} were all insertion-only streams so I wrote a program which would generate an insertion-deletion stream from them by adding a defined proportion of deletion instructions throughout the stream.
+This program maintained the property that no edge was deleted before it was inserted.
+\par I wrote \texttt{utility.cpp} to perform a number of tasks to these graphs.
+Most notably, it can produce a file which lists the final degree of every vertex in the graph stream and it can verify whether a given set is a subset of the neighbourhood of a given vertex.
+I used the later function to verify the outputs generated by implementation of Dr Konrad's algorithms.
+%TODO table of used graphs
+
+%TODO confirm values for gplus_large
+\begin{center}
+	\begin{table}[h]
+		\caption{Details of graphs used during evaluation}
+		\tiny
+		\begin{tabular}{|l|l|l|l|l|l|l|l|}
+			\hline
+			\textbf{Name}&\textbf{Source}&\textbf{Type}&\textbf{\# Vertices}&\textbf{\# Edges}&\textbf{\# Instructions}&\textbf{Max Degree}&\textbf{File Size}\\
+			\hline
+			$\mathtt{facebook\_small}$&SNAP&Insertion-Only&52&292&-&36&3 KB\\
+			$\mathtt{facebook\_small\_deletion}$&Generated&Insertion-Deletion&52&258&326&33&5 KB\\
+			$\mathtt{facebook}$&SNAP&Insertion-Only&747&60,050&-&586&587 KB\\
+			$\mathtt{facebook\_deletion}$&Generaetd&Insertion-Deletion&747&53,457&66,643&522&846 KB\\
+			$\mathtt{gplus}$&SNAP&Insertion-Only&12,417&1,179,613&-&5,948&53 MB\\
+			$\mathtt{gplus\_deletion}$&Generated&Insertion-Deletion&12,417&1,049,309&1,309,917&4,998&60 MB\\
+			$\mathtt{gplus\_large}$&SNAP&Insertion-Only&102,100&30,238,035&-&104,947&1.3 GB\\
+			$\mathtt{1000vertices}$&Generated&Insertion-Only&1,000&246,676&-&984&2 MB\\
+			$\mathtt{doublepoisson05ID}$&Generated&Insertion-Deletion&10,000&8,639,534&24,452,736&4,977&305 MB\\
+			$\mathtt{1000star}$&Generated&Insertion-Only&1,000&999&-&999&9 KB\\
+			$\mathtt{1000starID}$&Generated&Insertion-Deletion&1,000&999&1,023&999&11 KB\\
+			$\mathtt{1000complete}$&Generated&Insertion-Only&1,000&499,500&-&999&4 MB\\
+			$\mathtt{1000completeID}$&Generated&Insertion-Deletion&1,000&499,500&509,524&999&5 MB\\
+			\hline
+		\end{tabular}
+	\end{table}
+\end{center}
+
+\chapter{Insertion-Only Streams}
+
+\section{Degree-Based Reservoir Sampling}
+
+\texttt{Degree-Based Reservoir Sampling} is a technique for uniformly sampling a subset from the set of vertices whose degrees are greater than a specified minimum bound, while storing part of the neighbourhood for each of the sampled nodes.
+This sample is known as the \textit{reservoir}.
+
+\par \texttt{Degree-Based Reservoir Sampling} takes three parameters: the lower bound for the degree of vertices at which to start sampling, $d_1\in\nats$; the maximum neighbourhood size to store for each vertex, $d_2\in\nats$; and the maximum reservoir size, $s\in\nats$.
+\texttt{Degree-Based Reservoir Sampling} succeeds if it finds a neighbourhood of size $d_2$.
+
+%\par Degree-based reservoir sampling \underline{only} works for insertion-only graph streams.
+
+\par Three data stores are used in \texttt{Degree-Based Reservoir Sampling}: a map of the degree of every node in the graph; a set for the reservoir; and a set for edges incident to the vertices in the reservoir.
+
+\par The reservoir has an invariant that at any point in time it contains a uniform sample of the vertices whose degrees are known to be greater than $d_1$ at that time.
+This invariant is maintained by controlling how vertices are added to the reservoir.
+The first time the degree counter for a vertex $v$ surpasses $d_1$ we consider whether to add $v$ to the \textit{reservoir}.
+There are two cases
+\begin{itemize}
+	\item If the reservoir is \underline{not} full (ie its size is less than $s$) then we add $v$ to the reservoir.
+	\item Otherwise, if the reservoir is already full then we use a Bernoulli random variable with probability $p=\frac{x}s$, where $x$ is the size of the reservoir.
+If this random variable succeeds then we pick, uniformly at random, a vertex $u$ currently in the reservoir and replace it with $v$.
+We remove all the edges in the edge set which are incident to $u$ and no other vertices in the reservoir.
+\end{itemize}
+If $v$ is not added to the reservoir at this time, it never will be in the future.
+Note that in the case that the reservoir is full \& $v$ is inserted into it, $v$ is replacing an element that was in the reservoir and any progress which has been made towards finding a neighbourhood for that vertex is annulled in doing so.
+The amount of progress is not taken into account \& thus if there is a high turn-over of elements in a reservoir then less progress is made towards finding a neighbourhood for each of them.
+This is something I will look into during the evaluation.
+
+\par Whenever an edge is encounter which is in incident to a vertex in the reservoir then we add it to the set of edges.
+The sampler runs until the graph stream is exhausted at which point it returns, uniformly at random, one of the neighbourhoods of sufficient size.
+
+\par \texttt{Degree-Based Reservoir Sampling} only works on insertion-only streams as it does not account for when the degree of a vertex falls below the threshold, after being sampled.
+This means the reservoir is not necessarily a uniform random sample of the vertices with degree $\geq d_1$.
+%TODO is this sufficient
+
+\par For this project we are interested in how \texttt{Degree-Based Reservoir Sampling} can be implemented for bi-partite graphs, specifically we want to sample from the $A$-vertices of the graph.
+For a bi-partite graph with $n$ $A$-vertices \texttt{Degree-Based Reservoir Sampling} requires $O(n\log n+sd_2\log n)$ space, assuming $O(\log n)$ space is required to store an edge or a vertex, since it stores a degree counter for every $A$-vertex and at most $d_2$ edges for each of the $s$ vertices in the reservoir.
+\textbf{Algorithm 1} outlines pseudocode for \texttt{Degree-Based Reservoir Sampling} on a bi-partite graph.
+
+\begin{algorithm}
+	\caption{\texttt{Degree-Based Resevoir Sampling}$(d_1,d_2,s)$}
+	\SetKwInOut{Require}{require}
+	\Require{Lower degree bound $d_1\in\nats$, Upper neighbourhood bound $d_2\in\nats$, Reservoir size $s\in\nats$, Insertion-only stream $\{(a_0,b_0),\dots,(a_n,b_n)\}$}
+	$D\leftarrow\{\}$ \tcp*{Degree counter}
+	$R\leftarrow\{\}$ \tcp*{Reservoir}
+	$S\leftarrow\{\}$ \tcp*{Collected edges}
+	$x\leftarrow0$\tcp*{\# nodes of degree$\geq d_1$}
+	\For{$i\in[0,n]$} {
+		$D[a_i]\leftarrow D[a_i]+1$\tcp*{Increment degree counter}
+		\If{$D[a_i]\equiv d_1$}{
+			\tcp{Consider inserting $a_i$ to reservoir}
+			$x\leftarrow x+1$\\
+			\If{$|R|<s$} {
+			\tcp{Reservoir is not full}
+			$R\leftarrow R\cup\{a_i\}$
+			} \Else {
+				\tcp{Reservoir is full}
+				\If{\texttt{Bernoulli}$\left(\frac{s}x\right)$} {
+					Let $a'$ be a uniform random element of $R$\tcp*{Element to replace}
+					Delete edges in $S$ incident to $a'$\\
+					$R\leftarrow(R\backslash\{a'\})\cup\{a\}$\tcp*{Swap $a_i$ and $a'$}
+				}
+			}
+		}
+		\If{$a_i\in R$}{
+			$S\leftarrow S\cup(a_i,b_i)$\tcp*{Store edge}
+			\If{$D[a_i]\equiv d_1+d_2$} {
+				\Return{edges in $S$ incident to $a_i$\tcp*{Success}}
+			}
+		}
+	}
+	\Return{FAIL\tcp*{No sufficient neighbourhoods found}}
+\end{algorithm}
+
+\section{Proposed Algorithm}
+
+%\begin{itemize}
+%	\item Assumptions: no duplicate edges (can mitigate cases when this would be required by being smart)
+%	\item Insertion stream neighbourhoods cannot be diminished over time.
+%	\item Possibility of failure (missed edges) \& mitigation
+%	\begin{itemize}
+%		\item Theorem 3.2 (probs not proof)
+%	\end{itemize}
+%	\item Theoretical space requirements
+%\end{itemize}
+
+The \texttt{Neighbourhood-Detection Problem} can be solved using \texttt{degree-based reservoir sampling} sampling by setting $d_2=\frac{d}c$.
+The question remains as to what to set $d_1$ and $s$ to in order to get a high probability of success.
+In \cite{orig} it is proven that setting $s=\lceil\ln(n)\cdot n^{\frac1c}\rceil$ and using $c$ \texttt{degree-based reservoir samplers} each with a different lower bound, incremented from $\frac{d}c$ to $d$ stepping by $\frac{d}c$ each time, results in a high probability of success.
+The algorithm suggested by Dr.
+Konrad runs the samplers in this way, with $s$ set to $\lceil\ln(n)\cdot n^{\frac1c}\rceil$.
+The proposed algorithm returns, uniformly at random, any of the neighbourhoods return by a successful sampling.
+%TODO proof?
+Say it's insertion-only
+\par Each of the $c$ samplers requires $O(n\log n+sd_2\log n)$ space.
+Since the degree map can be shared between samplers the total space is $O(n\log n+c\cdot sd_2\log n)=O(n\log n+c\lceil\log(n)n^{\frac1c}\rceil\frac{d}c\log n)=O(n\log n+n^{\frac1c}d\log^2n)$.
+%TODO with algorithm parameter values
+The proposed algorithm does require you to know the number of vertices in the graph stream before running the algorithm, or at least the log of it.
+If this is unknown then it is quick to run through the whole stream \& build a set of vertices in the graph.
+This requires $O(n\log n)$ space, assuming $O(\log n)$ space is required to store a vertex.
+In many real-world scenarios this number (or a good approximation) will be known due to it being important for other tasks.
+\par \textbf{Algorithm 2} is the algorithm proposed in \cite{orig} for solving the \texttt{Neighbourhood-Detection Problem} for insertion-only streams.\\
+
+\begin{algorithm}
+	\caption{One-pass $c$-Approximation Insertion-Only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
+	\SetKwInOut{Require}{require}
+	\Require{Space $s$, degree bound $d$.}
+	$s\leftarrow\lceil\log(n)\cdot n^{\frac1c}\rceil$\\
+	\For{$i\in[0,c-1]$\text{\textbf{ in parallel}}} {
+		$(a_i,S_i)\leftarrow\mathtt{Deg}\mbox{-}\mathtt{Res}\mbox{-}\mathtt{Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
+	}
+	\Return{\small Uniform random neighbourhood $(a_i,S_i)$ \text{from successful runs}}
+\end{algorithm}
+
+\section{Implementation}
+
+I centred my implementation around working for the graph streams from \textit{SNAP} \cite{SNAP} as they came from real-world sources \& thus are very important for testing the real-world practicalities of the proposed algorithm.
+These graph streams are not formatted in a strictly bi-partite way.
+I adjusted \texttt{degree-based reservoir sampling} to account for this by by repeating \texttt{lines 6-19} of \textbf{Algorithm 1} after the end of the \texttt{for} loop, with all occurrences of $a_i$ replaced by $b_i$.
+%TODO something about implementing in parallel. Nah
+\vspace{.3cm}\newline
+After an initial implementation I considered two alterations which could be made:
+\begin{enumerate}
+	\item \textit{Early Termination} - Returning the first neighbourhood of sufficient size, rather than uniformly sample from all which succeed; and,
+	\item \textit{Shared Edge-Set} - Having all samplers share an edge-set, rather than each have one each.
+\end{enumerate}
+
+\noindent\textbf{Algorithm 2} states to run through the whole stream before sampling uniformly from the successfully returned neighbourhoods, but this is not necessary to fulfil \texttt{neighbourhood detection}.
+Returning the first neighbourhood of size $\frac{d}c$ is sufficient for solving the problem.
+Implementing this should reduce run-time as only part of the stream is being used and should reduce space used by the edge-sets in reservoir sampling as no data is stored after the first solution is found.
+As this change terminates at the first success there is no change to the success rate of the algorithm.
+One down side is that you will likely get less variety in the returned neighbourhoods as the algorithm will not encounter every vertex which meets the degree requirements, however this is not relevant to the problem.
+
+\par I implemented this change and evaluated the implementation at different values of $c$.
+Space-requirements more than halved in every instance.
+For graphs with the least variation in degree distribution space-requirements decreased by more than $97\%$ after early termination was introduced (as there are many vertices with valid neighbourhoods).
+Introducing early termination meant that run-times reduced as $c$ increased, whereas before run-times increased with $c$.
+These results are shown in \texttt{Figure 1} \& \texttt{Figure 2}.
+These are really good results, so I implemented this change going forward.
+%TODO explain reasons behind these results
+%TODO comment about effect on space used by reservoirs vs degree map
+%TODO this change means that time taken is now a measure of success as failure is (basically) the only time the whole stream is run through
+
+\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+
+\textbf{Lemma 2.1} If \textbf{Algorithm 2} terminates after the first sufficient neighbourhood it finds, no vertex occurs in multiple reservoirs at the same time.\\
+
+% TODO a proof by contradiction
+\textit{Proof (by contradiction)} Let $v$ be a vertex in the graph stream.
+Suppose, without loss of generality, that $v$ is a member of the reservoirs of both \texttt{Deg-Res-Sampling}$(\frac{d}{c},\frac{d}{c},s)$ \& \texttt{Deg-Res-Sampling}$(\frac{2d}{c},\frac{d}{c},s)$.
+This means that $v$ was sampled twice, after encountering the $\frac{d}c^\text{th}$ \underline{and} $\frac{2d}c^\text{th}$ edge incident to $v$ in the stream.
+After sampling $v$ for the first time, encountered edges which are incident to $v$ are stored.
+This means that by the time $v$ is sampled for the second time $\frac{2d}c-\frac{d}c=\frac{d}c$ edges incident to $v$ have been stored.
+This is a sufficient neighbourhood for \textbf{Algorithm 2} to terminate.
+This is a contradiction since termination occurs before the second sampling occurs.$\hfill\square$
+
+}}\end{adjustwidth}\vspace{.3cm} %TODO refine
+
+\par A consequence of \textbf{Lemma 2.1} is that an edge is only stored in multiple edge sets if its endpoints are sampled by different samplers. %TODO mention this is rare
+This means that introducing a shared edge-set for all samplers will not have a great effect on the space requirements of \texttt{Algorithm 2}.
+\par However, the alterations to the implementation which would have to made to allow for a shared edge-set could introduce time savings.
+Namely, I would need a data-structure which tells me \underline{precisely} which reservoir each vertex is in that so I know the appropriate value of $d_1$ in \texttt{Algorithm 1:Line 18}, otherwise improper termination could occur.
+An appropriate data-structure for this would be a \texttt{map} from each vertex to a bitstring whose indices indicate whether the vertex lies in an associated reservoir.
+With this implementation, inserting or deleting a vertex to/from a reservoir and querying whether a vertex is in a given reservoir becomes querying a map \& then a known bit in a bitstring which both take $O(1)$ time.
+This \texttt{map} replaces the \texttt{ordered set}s currently used to store the reservoirs of each sampler and thus reduce these action times from $O(\log n)$ to $O(1)$ time.
+As these actions occur a lot this change should result in a reasonable reduction in run-time.
+% TODO something about space used to store a vertex in the map vs in sets
+
+\par I implemented this map, along with a shared edge-set, %NOTE even though this shouldn't change much
+and ran tests for values of $c\in[3,100]$.
+These tests showed an ~8\% reduction in run-time when the map \& shared edge-set where introduced.
+It should be noted that in a few cases the algorithm was slower but <1\% which I think is a reasonable trade off.
+% TODO maybe include the variance
+There were no gains in the space requirements when a shared edge-set was used, nor were there losses. The results can be seen in \texttt{Figure 3} \& \texttt{Figure 4}.
+\par Implementing a shared-edge set does not change the logic of the algorithm so does not affect its success rate.
+Due to the reduction in mean run-time \& no noteworthy changes in space requirements, I recommend implementing this change.
+%TODO do I mean ``logic''??
+
+%TODO discuss performance against theory
+\section{Parameter Tuning}
+
+%\begin{itemize}
+%	\item Vary size of reservoir ($s$) \& number of samplers together to find an optimum
+%	\item Full optimised against v Naïve.
+%\end{itemize}
+
+\par The changes discussed above have been to the implementation of \texttt{degree-bound reservoir samplers}, rather than to the proposed algorithm.
+This means these changes have had no affect on the success rate of the algorithm.
+I will now discuss potential changes to the proposed algorithm.
+These changes will have an effect on the success rate of the algorithm.
+It is possible that no change in the success-rate will be seen \& rather the change will occur in execution time as more instructions have to be read before a solution is found.
+I am investigating the practicalities of the algorithm so keeping the run-times usable is a high priority.
+In \texttt{Algorithm 2} there are two parameters which can be played around with: the number of samplers; and, the size of the reservoirs, $s$.
+
+\par The total sample size, $\left\lceil\ln(n)n^{\frac1c}\right\rceil\times\text{\#\ Samplers}$, is the main factor effecting the space-used by the algorithm as this is the number of vertices it needs to store edges for.
+Noting that \texttt{Algorithm 2} sets the number of samplers to be $c$ and that $cp\left\lceil\ln(n)n^{\frac1c}\right\rceil\leq c\left\lceil p\ln(n)n^{\frac1c}\right\rceil\ \forall\ p\in[0,1]$ it is apparent that reducing the number of samplers has a greater effect on total sample size, and thus space-used, than reducing sample size by the same proportion.
+For this reason I investigated the number of samplers first.
+
 \horizontalline
-% Build to problem
-\par The $\mathtt{High\ Degree\ Detection\ Probem}$ is a problem in graph theory where we seek to find a vertex of sufficiently high degree, in a given graph.
-This node does not necessarily have to have the greatest degree in the graph, and often we cannot be sure whether it does as the greatest degree is unknown beforehand.
-This problem can be formally stated as
-\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
-\textbf{Problem 1} $\texttt{High Degree Detection}$.\\
-Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A,B$, where $|A|=n$ and $|B|=\text{poly }n$ , and edge set $E$.
-Set a restriction that at least one vertex in $A$ has degree, at least, $d$.\\
-In $\texttt{High Degree Detection Detection}(G,d)$ we are tasked with outputting a vertex from $A$ with degree, at least, $d$.
-}}\end{adjustwidth}\vspace{.3cm}
-\par This problem can be solved efficiently for both insertion-only \& insertion-deletion graph streams with a fairly simple algorithm which simply runs through all the edges, incrementing or decrementing (depending on whether it is an insertion or deletion edge)  a count for the degree of each vertex as it goes.
-Returning the first vertex it finds with degree equal to $d$.\\
-%TODO Do I have to justify the algorithm
-\begin{algorithm}[H]
-\caption{Single Pass High Degree Detection}
+
+\par When changing the number of samplers we need to consider their lower bounds, $d_1$, too.
+If the range of vertex degrees which individual samplers sample, $[d_1,d_1+d_2]$, at overlap then duplicate sampling of vertices can occur which offers no advantages \& leads to a reservoir space being wasted.
+And, if there are gaps between these ranges then there are vertices which may not be sampled.
+%TODO flesh this out more.
+Thus, the best set up is to have the sampling ranges be sequential with no gaps and so when varying the number of samplers we should only add or remove from the endpoints of these ranges.
+The samplers with lower lower-bounds, $d_1$, start sampling earlier \& from a larger pool.
+I expect these early samplers to be more important wrt performance gains, than later samplers.
+It is not necessarily true that the first sampler is the most important as it is sampling from the largest pool and so has the greatest turn-over rate in its reservoir.
+
+\par I ran the following tests to investigate what had a greater effect on performance, removing later samplers or earlier samplers.
+\begin{enumerate}
+	\item Removing the \underline{first} $x$ samplers from \texttt{Algorithm 2} for $x\in[1,c)$, for $c\in[1,20]$.\\
+	\textit{i.e.} Change the \texttt{for} loop in \texttt{Algorithm 2:Line 2} to run for $i\in[x,c-1]$.
+	\item Removing the \underline{last} $x$ samplers from \texttt{Algorithm 2} for $x\in[1,c)$, for $c\in[1,20]$.\\
+	\textit{i.e.} Change the \texttt{for} loop in \texttt{Algorithm 2:Line 2} to run for $i\in[0,c-x]$.
+\end{enumerate}
+
+\par These tests confirmed the hypothesis that the lower a sampler's lower-bound is, the greater the effect of its inclusion on performance.
+Time-taken increased as more early samplers were removed, and decreased as more late samplers were removed.
+Comparing the relative difference of these results shows that removing early samplers is more detrimental to performance than including later samplers.
+This is due to late samplers only starting to sample later \& they may not have begun sampling by the time one of the previous samplers has succeed, terminating the algorithm.
+When the same proportion of samplers were removed for different values of the approximation factor, $c$, the relative performance was worse for greater values of $c$.
+This is to be expected as the real number of samplers removed increases with $c$.
+As expected the space-used decreased whenever more samplers were removed.
+During the second test some failures did occur when only the first sampler was used on larger graphs.
+The results to these tests on \texttt{gplus} can be seen in \texttt{Figure 5} \& \texttt{Figure 6}, results for other graphs followed similar trends.
+
+\par From these tests it is apparent that there are advantages to be found by removing late samplers, but not from removing the earliest.
+These tests did show that the failure rate does increase when too few of the early samplers are used.
+Thus I repeated the tests, this time testing when $\leq25\%$ of the earliest samplers were used and ran more repetitions so I could evaluate the failure rate better.
+
+\par This test showed that the success rate was less than 1 only when just the first sampler was used, even on the largest graphs.
+Even in these cases it was still above $.9$.
+Since failure only occurs when the whole stream is run through, runs which failed take a lot longer to run than those that succeed (5 minutes compared to $~10$ seconds for \texttt{gplus\_large}).
+This meant that the implementations which only used the first two samplers were quicker on average.
+
+\par The success rate when only the first sampler was used was lower for larger graphs, but was higher for greater values of the approximation factor.
+In all my tests, running just the first two samplers provided the best results.
+However, the largest graph I tested on only contains $~100,000$ vertices so it is likely this may not hold for significantly larger graphs.
+I suggest using the first $\mathtt{max}\left(2,\left\lceil\frac15\ln n\right\rceil\right)$ samplers.
+As we don't want the sampling ranges to overlap we need to add a restriction that the number of samplers is never greater than the approximation factor, giving the number of samplers used as $\mathtt{min}\left(c,\mathtt{max}\left(2,\left\lceil\frac15\ln n\right\rceil\right)\right)$.
+
+\par When tested on \texttt{gplus\_large} introducing this change halved run times for tight approximation factors ($c\in[3,10]$), for looser approximation factors time-savings of over 90\% were made.
+This is because the number of samplers removed increases as the value of the approximation factor increases.
+This implementation had little effect on space-usage when compared to the implementation before the number of samplers was reduced, whilst maintaining a perfect success rate.
+The lack of change in space-usage shows that these later samplers were not being used and were just adding wasted computation time.
+These results can be seen in \texttt{Figure 7} \& \texttt{Figure 8}.
+%TODO something about how the fact that space hasn't changed much means later samplers weren't coming into play often & thus sample size is unlikely to be important.
+%TODO comment of special cases
+\horizontalline
+\par Removing later samplers had little effect on the space used by the algorithm.
+This is due to these samplers not getting the chance to sample very often as the algorithm often terminated before many vertices of sufficient degree were encountered.
+This means that reducing the size of those samplers would have had virtually no affect on the performance of the algorithm.
+This helps justify optimising the number of samplers before the sample sizes.
+The previous optimisation means that very few samplers are being run so it is unlikely that any great improvements will be found, if any, when reducing sample size.
+
+\par Decreasing the sample size $s$ means less edges are stored, reducing space requirements, but increases the turn-over rate of the elements in the reservoir as the probability a vertex is sampled is the same but the sample size is smaller.
+The greater turn-over rate is detrimental to the success of the algorithm as it more likely for a vertex to be removed when it is close to succeeding.
+For very low sample sizes it is likely that the success rate of the algorithm will remain high, but the run-times will be dramatically increased.
+
+\par I tested having the sample size set to between $10\%$ and $200\%$ of the proposed size, at $10\%$ increments, in order to get an overview of how sample size affects performance.
+This test showed a positive linear relationship between sample size \& space-used.
+The only inference that could be drawn from sample-size \& time-taken is that there is a point at which time-taken dramatically increases, this is the point where the algorithm fails to return a result.
+Overall, there is no clear strategy to improving the algorithm by varying the sample size.
+There may have been if we had not already reduced the number of samplers, but for reasons discussed earlier it is justified to optimise the number of samplers first.
+These results can be seen in \texttt{Figure 9} \& \texttt{Figure 10}.\\
+
+%TODO mention theoretical space efficieny of this (should be same as before)
+\par Here is \textbf{Algorithm 2} rewritten to include the proposed parameter optimisations
+\begin{algorithm}
+	\caption{One-pass $c$-Approximation Insertion-Only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
 	\SetKwInOut{Require}{require}
-\Require{Stream $\{(i_0,x_0,y_0)\dots(i_n,x_n,y_n)\}$, degree bond $d$}
-$D\leftarrow\{\{\}\}\ \{\text{degree map}\}$\\
-\For{$j\in[0\dots n]$} {
-	\uIf{$x_j\in D$}{
-		\lIf{$i_j$}{$D[x_j]-=1$}
-		\lElse{$D[x_j]+=1$}
+	\Require{Space $s$, degree bound $d$.}
+	$s\leftarrow\lceil\log(n)\cdot n^{\frac1c}\rceil$\\
+	$n_s\leftarrow\min\left(c,\max\left(2,\left\lceil\frac15\ln(n)\right\rceil\right)\right)$\\
+	\For{$i\in[0,n_s]$\text{\textbf{ in parallel}}} {
+		$(a_i,S_i)\leftarrow\mathtt{Deg}\mbox{-}\mathtt{Res}\mbox{-}\mathtt{Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
 	}
-	\lElse{$D[x_j]=1$}
-	\lIf{$D[x_j]=d$}{\Return{$x_j$}}
-	\textbf{repeat} with $y_j$
-}
-\Return{FAIL}
-\end{algorithm}\vspace{.3cm}
-In \textbf{Algorithm 1} $i$ is a boolean which is true if the edge is an insertion edge, and false if it is a deletion edge.
-$x\ \&\ y$ are the two endpoints of the edge.
-The algorithm makes the reasonable assumption that you will never recieve a deletion edge for an edge which is not in the graph, nor shall you recieve the same edge twice.
-\par The $\mathtt{High\ Degree\ Detection\ Probem}$ is limited in its applications as it only returns a vertex with high degree \& gives you no information as to its neighbourhood.
-Solving this problem would not be very helpful when it comes to choosing influencers as it would only return people who have large followings, but not who their followers are.
-Thus an advertisier would not be able to evaluate whether their followers represented the audience they wished to market to.
-\par The $\mathtt{Neighbourhood\ Detection\ Problem}$ is an extension of the $\mathtt{High\ Degree}$ $\mathtt{Detection\ Problem}$.
-The task is to return a vertex with sufficiently high degree \& a certain proportion that vertex's neighbourhood.
-This is formally stated as
-\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
-%TODO Shorten this, make it neater.
-\textbf{Problem 2} $\texttt{Neighbourhood Detection}$.\\
-Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A,B$, where $|A|=n$ and $|B|=\text{poly }n$, and edge set $E$.\\
-In $\texttt{Neighbourhood Detection}(G,d,c)$ we are tasked with outputting a vertex from $A$ and at least $d/c$ of its neighbours in $B$.\\
-Here $d$ is a threshold parameter \& $c$ is an approximation parameter.
-}}\end{adjustwidth}
-\par Solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ allows for a much more general assessment of the influence of the returned vertex.
-In the context of finding social media influences the returned neighbourhood can be assessed to determine whether the returned person is suitable for a given advertising campaign.
-When assessing the members of the neighbourhood you could look at the demographics of the members \& how active the members are (among other features) in order to determine whether they fit the target of the campaign.
-For cases where $c\neq 1$ you only get a fraction of a vertices whole neighbourhood but for sufficiently high $d$ you an implement many statistical tools to still be able to draw meaningful inferences.
-Later it shall be shown that there are meaningful advantages to increasing the value of $c$.
+	\Return{\small Uniform random neighbourhood $(a_i,S_i)$ \text{from successful runs}}
+\end{algorithm}
 
-\section{Why Space Efficiency?}
-Simply solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ is fairly trivial.
-\textbf{Algorithm 2} is an algorithm for solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ for insertion-only graph streams by recording the vertices in the neighbourhood of every $A$-vertex and then returning the first neighbourhood we enounter with $\frac{d}{c}$ members.\\
-\begin{algorithm}[H]
-\caption{Na\"ive Single-Pass Insertion-Streaming Algorithm for Neighbourhood Detection}
+%TODO rewrite algorithm 2 with these parameter optimisations
+\section{Evaluation}
+
+%Run the following instances on different graphs \& compare
+%\begin{itemize}
+%	\item Initial implementation (no tuning, not quit early, not sharing)
+%	\item Implementation with early quitting & shared edges
+%	\item Implementation with tuned algorithm parameters
+%	\item Naïve Implementation
+%\end{itemize}
+% Compare max space to size of edges file
+% Identify which change had what effect (to space, time, success)
+%TODO 1000star is useful as there is only one valid solution
+To evaluate my final implementation I compared its performance to that of different stages of the optimisation and to a naïve implementation of the algorithm. These different stages were
+\begin{enumerate}[label=\roman*)]
+	\item Initial Implementation - No early termination, shared edge-set or sampler removal. %insertionStreamsNotQuitEarly
+	\item After Technical Optimisation - Implements early termination \& a shared edge-set. %insertionStreamsSharedEdgeSet
+	\item After Algorithmic Optimisation - Implements early termination, shared edge-set and sampler removal. %insertionStreamsRemovedSamplers
+\end{enumerate}
+The naïve algorithm I used is given in \texttt{Algorithmn 4}.
+This aproach records the neighbourhood of every encountered vertex \& returns the first one which surpasses the $\frac{d}c$ threshold.
+This algorithm will always succeed as there is a requirement that at least one vertex in the graph is of degree, at least, $d$.
+As the naïve algorithm terminates at the first neighbourhood of sufficient size \& uses all edges in the stream up to that point, it is guaranteed to find the first vertex with a sufficient neighbourhood.
+This means the naïve algorithm is very quick, provided its space-usage is within the memory allocated to the algorithm.
+In the worst case, where the last element is the only element of degree $\frac{d}c$ and all other $n-1$ elements have degree $\frac{d}c-1$, $\frac{d}{c}+(n-1)(\frac{d}c-1)\in O(n\frac{d}c)$ space is required.
+The average case has the same complexity and the best case, where the first $\frac{d}c$ instructions are incident to the same vertex, has $O(\frac{d}c)$ space-complexity.
+\begin{algorithm}
+	\caption{Na\"ive Single-Pass Insertion-Only Streaming Algorithm for Neighbourhood Detection}
 	\SetKwInOut{Require}{require}
@@
 	\For{$i=0\dots n$} {
-	append $t_i$ to $N[s_i]$\\
-	\lIf{\text{size}($N[s_i])\geq \frac{d}{c}$} {
-		\Return{($s_i$,$N[s_i]$)}
+		append $t_i$ to $N[s_i]$\tcp*{Insert neighbour}
+		\If{\text{size}($N[s_i])\geq d$} {
+			\tcp{Sufficiently large neighbourhood found}
+			$S\leftarrow \{N[s_i][0],\dots,N[s_i][\frac{d}{c}]\}$\tcp*{First $\frac{d}c$ elements of neighbourhood}
+			\Return{($s_i,S$)}
 		}
 	}
+	\Return{FAIL}\tcp*{No neighbourhoods found}
+\end{algorithm}
+
+% TODO a table of results??
+
+\par The empirical results showed that the space usage for the naïve algorithm was significantly greater than that of any of the proposed implementations.
+This high usage of space could be seen in practice as the naïve algorithm did not return a solution when the approximation factor was 2 for \texttt{gplus\_large}, even after 3 hours of computation.
+For all the graphs, except \texttt{1000star}, the implementations of the proper algorithm used less than the stream file size for all approximation values.
+For the largest graph tested, \texttt{gplus\_large}, only $13.7\%$ of the stream file size was required to find a 2-approximation using the initial implementation.
+After my optimisations this was reduced to $1.39\%$.
+The space-usage results for \texttt{gplus}, the largest file which the naïve algorithm was usable on, are given in \texttt{Figure 11} with the results normalised by the stream's size (51,839 KB).
+
+%TODO comment on performance compared to theory
+
+\par As I supposed the naïve algorithm was generally the fasted algorithm, in the cases where it succeeded, as it always used the fewest number of instructions required.
+My final implementation, after algorithmic tuning, was between 50-100\% slower than the naïve implementation in these cases.
+My final implementation is useably quick: when run on \texttt{gplus\_large} it took ~150s to find a solution when $c=2$ and less than a minute for $c\geq5$.
+The naïve algorithm being quicker is irrelevant for this evaluation as it fails to terminate for larger files due to its space inefficiency.
+The time taken results for \texttt{gplus}, the largest file tested, are shown in \texttt{Figure 12}.
+
+\section{Conclusion}
+
+\par During the optimisation process I made three changes which helped improve the performance of the algorithm: I made the algorithm return the first neighbourhood it encountered of sufficient degree.
+This reduced both time \& space requirements as less of the graph stream was consumed; I set up the samplers to share an edge set.
+This was meant to reduce space requirements, but due to early termination duplication of edges was now very rare.
+This change did reduce run times as the time complexity for operations on reservoirs was reduced from $O(\log n)$ to $O(1)$; And, I reduced the number of samplers being run by removing most of the late samplers which were very rarely used (due to early termination).
+This change had little effect on space requirements, but did reduce run-time substantially, especially for high values of the approximation factor.
+After all these changes the algorithm maintained a perfect success rate during all final tests.
+
+\par I proposed changing the sample size of each reservoir.
+Preliminary tests presented no strategies which could consistently improve one performance metric without unreasonably affecting another.
+Thus this idea was not implemented.
+It is likely that if the number of samplers had not been reduced by as much as then varying the sample size could have had a positive effect on the algorithm, but the gains from reducing the number of samplers are so good they must be incorporated.
+
+\par The final evaluation shows that the proposed algorithm is much more space-efficient that a naïve approach \& even for close approximations of large streams my final implementation uses less than $6\%$ of the file size and for $c\geq5$ it uses $\leq1\%$.
+This space efficiency increases the maximum file size at which the algorithm is usable, with without requiring virtual memory.
+The run-times of my final implementation are very usable with solutions being found in a couple of minutes for close approximations ($c\in[2,4]$) and taking significantly less time for loser approximations.
+Whilst evaluating my final implementation no failures occurred.
+All these results show that the proposed algorithm for insertion-only streams \underline{is} indeed practical in real-world scenarios.
+
+\chapter{Insertion-Deletion Streams}
+
+\section{$L_0$ Sampling}
+
+\begin{itemize}
+	\item L0 sampling uniformly samples from the non zero elements of a vector.
+	\item In this setting we consider the edge vector of each vertex (indicates which vertices it shares an edge with).
+	\item The formally proposed algorithm.
+\end{itemize}
+
+\subsection{Implementation}
+%NOTE this was hard
+\begin{itemize}
+	\item How to L0 sample a graph stream
+	\item Hash functions are hard (trade off due to dodgy hash generation).
+\end{itemize}
+
+\section{The Algorithm}
+
+\begin{itemize}
+	\item There are two algorithms.
+	\item Scenarios when each algorithm is most suitable.
+	\item L0 samplers in parallel (lots of pre-allocated space).
+\end{itemize}
+
+\begin{algorithm}[H]
+	\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Vertex Sampling)}}
+	\SetKwInOut{Require}{require}
+	\Require{Space $s$, degree bound $d$.}
+	Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$\\
+	Sample a uniform random subset $A'\subseteq A$ of size $10\ x\ln n$ of vertices.\\
+	\For{$a\in A'$} {
+		Run $10\frac{d}{c}\ln n$ $l_0$-samplers on the set of edges incident to $a$.
+	}
+	\Return{\small Any neighbourhood of size $\frac{d}{c}$ among the stored edges, if there is one.}
+\end{algorithm}
+
+\begin{algorithm}[H]
+	\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Edge Sampling)}}
+	\SetKwInOut{Require}{require}
+	\Require{Space $s$, degree bound $d$.}
+	Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$.\\
+	Run $10\frac{nd}{c}\left(\frac1x+\frac1c\right)\ln (nm)$ $l_0$-samplers on the set of edges incident to $a$.\\
+	\Return{\small Any neighbourhood of size $\frac{d}{c}$ among the returned edges, if there is one.}
+\end{algorithm}
+
+\section{Implementation}
+%TODO will be shorter than for Insertion-only
+
+\begin{itemize}
+	\item L0 samplers in parallel
+	\begin{itemize}
+		\item require lots of preallocated space (know how much space is used at start of algorithm)
+	\end{itemize}
+	\item Require list of vertices before vertex sampling.
+\end{itemize}
+
+\section{Parameter Tuning}
+
+\begin{itemize}
+	\item gamma \& delta for L0 sampling
+	\item vertex sample size \& samplers per vertex
+	\item num edge samplers
+\end{itemize}
+
+\section{Evaluation}
+
+% L0 space v sample size space
+
+\begin{algorithm}[H]
+\caption{Na\"ive Single-Pass Insertion-Streaming Algorithm for Neighbourhood Detection}
+\SetKwInOut{Require}{require}
+\Require{Stream $\{(w_0,s_0,t_0)\dots(w_n,s_n,t_n)\}$, degree bound $d$, precision bound $c$}
+$N\leftarrow\{\{\}\}\ \{\text{neighbourhoods}\}$\\
+\For{$i=0\dots n$} {
+	\If{$w_i\equiv1$}{append $t_i$ to $N[s_i]$\tcp*{Insertion instruction}}
+	\ElseIf{$w_i\equiv-1$}{remove $t_i$ from $N[s_i]$\tcp*{Deletion Instruction}}
+}
+\tcp{Look for a sufficiently large neighbourhood}
+\For{$s_i\in N.keys$}{
+	\If{\text{size}($N[s_i])\geq d$} {
+		\tcp{Sufficiently large neighbourhood found}
+		$S\leftarrow \{N[s_i][0],\dots,N[s_i][\frac{d}{c}]\}$\tcp*{First $\frac{d}c$ elements of neighbourhood}
+		\Return{($s_i,S$)}
+	}
+}
+\tcp{No neighbourhoods found}
 \Return{FAIL}
 \end{algorithm}
-%TODO check space requirements
-\par\textbf{Algorithm 2} requires $O\left(n^2\right)$ space.
-Thus the space used to run the algorithm grows very quickly which could easily lead to the computer's RAM overflowing, increasing execution time.
-This is undesirable, especially given the run time complexity of \textbf{Algorithm 2} is alread bad, $O(n^2)$.
-% nc/d different vertices with neighbourhoods of d/c-1.
-Reading to position d/c-1, nc/d times
-For solutions which are space efficient run times become unmanagable for large graph streams.
-In 2014 in the UK, Facebook had 36.68 million users \cite{users16} each with an average of 155 friend connections \cite{friends16}.
-In order for this to be represented in a bi-partite graph for $\mathtt{neighbourhood\ detection}$ we would need a graph stream with ${\sim}5.6$ billion edges.
-\textbf{Algorithm 2} would not be suitable for analysing this graph.
-\par Instead we must look for much more space efficient algorithms for\\
-$\mathtt{neighbourhood\ detection}$.
-In the rest of this paper I shall discuss \& test an algorithm which can solve $\mathtt{neighbourhood\ detection}$ for insertion-only streams with space $O(n\log n+n^{\frac1c}d\log^2n)$; and, two approaches to solving for $\mathtt{neighbourhood\ detection}$ for insertion-deletion streams with space $\tilde O\left(\frac{xd}{c}\right)\ \&\ \tilde O\left(\frac{nd}{c}\left(\frac1x+\frac1x\right)\right)$ respectively.
 
-\section{Plan}
-In this paper I shall implement and test the two algorithms presented by Dr Chrisian Konrad in \cite{orig}.
-These algorithms solve the \textit{Neighbourhood Detection Problem} for insertion-only \& insertion-deletion graph streams in a space-efficient manner.
-Both algorithms require only a single pass of the stream.
-\par I shall use the na\"ive algorithm described in \textbf{Algorithm 2} as a benchmark for both time \& space efficiency of these two algorithms.
-I shall use a four different graph streams, described in \textbf{Table 1.1}, for my tests.
-% probs should have more graphs
-These graphs were acquired from the \textit{Stanford Network Analysis Project}, \cite{SNAP}.
-\par The theoretical space requirements for these two algorithms both depend on the degree variable, $d$, and precision variable, $c$.
-I shall perform tests where I vary each of these in order to see what the space requirements are in practice, and then compare these results to the theory.
+\section{Conclusion}
 
-\begin{table}[h]
-\caption{Test Graph Streams}
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-Name&\# Edges&\# Vertices&Max Degree&File Size, KB\\\hline
-$\mathtt{facebook\_small}$&292&52&36&3\\
-$\mathtt{facebook}$&60,050&747&586&587\\
-$\mathtt{gplus}$&1,179,613&12,417&5,948&52,839\\
-$\mathtt{gplus\_large}$&30,238,035&120,100&104,947&1,328,820\\\hline
-\end{tabular}
-\label{Tab:Tcr}
-\end{table}
+\chapter{Conclusions}
+
+\section{Future Work}
+
+\chapter{Related Works}
+
+\chapter{Appendix}
+\section*{THESE ARE ALL PLACEHOLDER}
+
+\begin{figure}
+	\label{Figure 1}
+  \includegraphics[width=\textwidth]{img/terminationSpace.png}
+	\caption{Normalised mean space used for different termination strategies}
+	%"Insertion-Only Tests"
+\end{figure}
+\begin{figure}
+	\label{Figure 2}
+  \includegraphics[width=\textwidth]{img/terminationTime.png}
+	\caption{Normalised mean time taken for different termination strategies}
+	%"Insertion-Only Tests"
+\end{figure}
+\begin{figure}
+	\label{Figure 3}
+  \includegraphics[width=\textwidth]{img/edgeSpace.png}
+	\caption{Normalised mean space used for different edge storage strategies}
+	%"Insertion-Only Tests"
+\end{figure}
+\begin{figure}
+	\label{Figure 4}
+  \includegraphics[width=\textwidth]{img/edgeTime.png}
+	\caption{Normalised mean time taken for different edge storage strategies}
+	%"Insertion-Only Tests"
+\end{figure}
+\begin{figure}
+	\label{Figure 5}
+  \includegraphics[width=\textwidth]{img/removedEarlySamplers.png}
+	\caption{Normalised mean time-taken \& space-used when removing different proportions of the early samplers on graph \texttt{gplus}}
+	%"num sampler tests"/results_removed_early_samplers_gplus
+\end{figure}
+\begin{figure}
+	\label{Figure 6}
+  \includegraphics[width=\textwidth]{img/removedLaterSamplers.png}
+	\caption{Normalised mean time-taken \& space-used when removing different proportions of the late samplers on graph \texttt{gplus}}
+\end{figure}
+\begin{figure}
+	\label{Figure 7}
+  \includegraphics[width=\textwidth]{img/reducedSamplersGPlusTime.png}
+	\caption{Time-taken before \& after removing late samplers from \texttt{gplus\_large}}
+	%"num sampler tests"/"Compare Improvement"
+\end{figure}
+\begin{figure}
+	\label{Figure 8}
+  \includegraphics[width=\textwidth]{img/reducedSamplersGPlusSpace.png}
+	\caption{Space-used before \& after removing late samplers from \texttt{gplus\_large}}
+	%"num sampler tests"/"Compare Improvement"
+\end{figure}
+\begin{figure}
+	\label{Figure 9}
+  \includegraphics[width=\textwidth]{img/variedSampleSizeTime.png}
+	\caption{Time-taken when varying the reservoir sample size.}
+	%"sample size tests"/results_proportional_sample_size_reduction_gplus
+\end{figure}
+\begin{figure}
+	\label{Figure 10}
+	\includegraphics[width=\textwidth]{img/variedSampleSizeSpace.png}
+	\caption{Space-used when varying the reservoir sample size.}
+	%"sample size tests"/results_proportional_sample_size_reduction_gplus
+\end{figure}
+\begin{figure}
+	\label{Figure 11}
+	\includegraphics[width=\textwidth]{img/insertionOnlyGplusSpace.png}
+	\caption{Space-used by different implementations of the algorithm for insertion-only streams when tested on \texttt{gplus}, normalised by stream size 51,839 KB}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+\begin{figure}
+	\label{Figure 12}
+	\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusLargeTime1.png}
+	\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusLargeTime2.png}
+	\caption{Time-taken by different implementations of the algorithm for insertion-only streams when tested on \texttt{gplus\_large}}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
 
 \begin{thebibliography}{9}
+	\addcontentsline{toc}{chapter}{Bibliography}
 	\bibitem{orig} Christian Konrad \textit{Streaming Frequent Items with Timestamps and Detecting Large Neighborhoods in Graph Streams.} November 2019

### 2020-05-03 paper/src/paper.tex
Write up almost done

--- 2020-02-10 paper/src/paper.tex
@@
 \usepackage{fancyhdr} % Header
+\usepackage{float} % [H] for figure
+\usepackage{graphicx}
+\usepackage{subfig}
+\usepackage{tikz}
 
@@
 \newcommand{\nats}{\mathbb{N} }
-\newcommand{\horizontalline}{\newline\vspace{.3cm}\hfill\makebox[.5\linewidth]{\rule{.5\textwidth}{0.4pt}}\hfill\vspace{.1cm}}
-% title
+\newcommand{\horizontalline}{\newline\vspace{.3cm}\hfill\makebox[.5\linewidth]{\rule{.5\textwidth}{0.4pt}}\hfill\vspace{.05cm}}
+
+\newcommand{\smallsim}{\smallsym{\mathrel}{\sim}}
+
+\makeatletter
+\newcommand{\smallsym}[2]{#1{\mathpalette\make@small@sym{#2}}}
+\newcommand{\make@small@sym}[2]{%
+  \vcenter{\hbox{$\m@th\downgrade@style#1#2$}}%
+}
+\newcommand{\downgrade@style}[1]{%
+  \ifx#1\displaystyle\scriptstyle\else
+    \ifx#1\textstyle\scriptstyle\else
+      \scriptscriptstyle
+  \fi\fi
+}
+\makeatother
+
+\renewcommand\thetable{\arabic{table}} % figure numbering strategy
+\renewcommand{\thefigure}{\arabic{chapter}.\arabic{section}.\arabic{figure}}
+
+\pagenumbering{roman}
+
+% Header
+\pagestyle{fancy}
+\fancyhead[L]{Dom Hutchinson (170 1111)}
+\fancyhead[C]{}
+\fancyhead[R]{\today}
+
+% Algorithm environment
+\SetArgSty{textnormal}
+\DontPrintSemicolon
+
+% Title
 \title{Implementing \& Evaluating Space Efficient Algorithms for Detecting Large Neighbourhoods in Graph Streams}
@@
 \maketitle
+\newpage
 
-% Header
-\pagestyle{fancy}
-\fancyhead[L]{Dom Hutchinson (1701111)}
-\fancyhead[C]{}
-\fancyhead[R]{\today}
+\chapter*{Declaration}
 
-\chapter{Background}
+\tableofcontents
+\newpage
+
+\chapter*{Abstract}
+
+\par The \texttt{Neighbourhood Detection Problem} is a problem in graph theory which tasks one with finding a vertex in the graph which has a certain number of neighbours and to then return a subset of these neighbours.
+This is a trivial problem to solve in theory, but in practice is hard to solve within a time frame which makes the implementation practical.
+A major reason for this is that many approaches to solving the problem require a lot of space and so quickly overflow a computer's RAM, slowing the execution significantly down.
+This common space inefficiency is becoming increasingly problematic since the move towards Big Data.
+
+\par In this paper I discuss and implement the two algorithms presented in \cite{orig} which offer space efficient solutions for \texttt{neighbourhood detection} for two types of graph stream.
+During the implementation I discuss and evaluate several methods for improving performance of these algorithms, focusing on space-efficiency in order to maximise the potential size of graphs the implementations are effective on.
+
+\par The final implementation of the proposed algorithm for insertion-only streams proves very promising with it able to find large neighbourhoods in a couple of minutes for a graph with 30 million edges.
+The algorithm for insertion-deletion streams proves much more difficult to evaluate due to restrictions incurred from the implementation of $L_0$ samplers.
+
+\chapter*{Acknowledgments}
+
+\renewcommand\thechapter{\Roman{chapter}}
+\renewcommand\thesection{\thechapter.\roman{section}}
+\setcounter{chapter}{1}
+\chapter*{Introduction}
+\addcontentsline{toc}{chapter}{Introduction}
 
 \section{Motivation}
-\par When designing algorithms time efficiency is generally viewed with primary importance, over space efficiency.
-However, if the memory used for an algorithm exceeds the RAM space allocated to it during execution then virtual memory would have to be used.
-Virtual memory has a much longer read/write time than RAM meaning that an algorithm's run-time will be much greater in this scenario.
-Thus, algorithms which are designed to work with large data sets space efficiency should be viewed as having primary importance.
-\par In the early 2000s companies began to realise that data could be used as a commodity, and with this, the amount of data being collected has surged in the years since.
-In this same time period, we have not seen the same surge in the amount of RAM typical computers have.
-Thus people wishing to capitalise on this increase in available data need to give greater thought to the space efficiency of the algorithms they are using.
-Otherwise, run-times could become unusable.
-\par Social networks are good sources of data about people.
-This data can be leveraged in many ways, notably within advertising.
-Advertisements which are targeted at people with certain observed behaviours have been shown to have more than twice the click-through rate of standard online advertisements \cite{targetedAds}.
-This has motivated the development of techniques for assessing the behaviours of a person just from their online presence.
-\par One technique is to look at who a person is connected to and then to use observations about these connections to infer the behaviours of said person.
-In turn, we have seen the rise of the ``Social Media Influencer".
-These are people who are deemed to have a large online following and, moreover, a following which they are able to influence.
-Advertisers will approach influencers with deals by which the influencer will promote a product to their audience, hopefully driving sales.
-\par The question for advertisers now becomes ``Who is an influencer?".
-To answer this we need to indentify people with large neigbourhoods in a network \& members of this network.
-We need to find the members of the neighbourhood (or at least a subset of them) in order to be able to assess what the demographics of the influencers audience are.
-This question can be generalised as the $\mathtt{Neighborhood\ Detection\ Problem}$.
-Which is the problem I shall discuss in this paper.
-\par Although in my description of the Neighbourhood Detection Problem I motivate its relevance by referring to finding social media influencers, it is applicable to many other scenarios.
+
+%\begin{itemize}
+%	\item High Degree Detection.
+%	\item Large Neighbourhood Detection.
+%\end{itemize}
+Graphs are used to represent relationships between objects and are popular in data science.
+A graph stream is a sequence of instructions which describe how to construct a graph.
+The sequential nature of a graph stream allows it to represent the evolution of a graph over time, which adjacency matrices cannot.
+This makes graph streams ideal for representing changing networks such as a social or computer network where objects can connect and then disconnect.
+
+\par A basic problem in graph theory is \texttt{High Degree Detection} where you want to find the vertices with the greatest degree.
+This problem can be trivially and efficiently solved by counting the number of edges incident to each vertex and then returning any vertices of sufficient degree.
+Solving \texttt{High Degree Detection} allows for the identification of the most influential vertices in a graph, however you cannot make any inferences about the nature of this influence.
+In a social network it is possible for someone to have lots of followers (\ie be of high degree), but for all these followers to be bot accounts meaning the account actually has zero influence.
+
+\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+\textbf{Problem 1} \texttt{Neighbourhood Detection}.\\
+Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A,B$, where $|A|=n$ and $|B|=\text{poly }n$, and edge-set $E$.\\
+In $\texttt{Neighbourhood Detection}(G,d,c)$ we are tasked with outputting a vertex from $A$ with at least $d/c$ of its neighbours in $B$.
+We can assume that $G$ contains at least one node of degree $d$.\\
+Here $d\in\nats$ is a threshold parameter \& $c\in(0,1]$ is an approximation parameter.
+}}\end{adjustwidth}\vspace{.3cm}
+
+\par\texttt{Neighbourhood Detection} is the natural set up from \texttt{High Degree Detection}, it tasks you with finding vertices of high degree \underline{and} a subset of their neighbours.
+Solving \texttt{Neighbourhood Detection} allows for inferences to be made about the influence a vertex has.
+Extending the social network example, \texttt{Neighbourhood Detection} allows for analysis of the followers of a popular account which can be used for targeted marketing campaigns.
+
+\par\texttt{Neighbourhood Detection} can be solved trivially by storing the neighbourhood of every vertex and then returning any subset of size $\frac{d}c$.
+This trivial solution is very space inefficient and, with the movement towards Big Data, would quickly overflow a computer's RAM.
+This results in the use of virtual data which significantly slows down data access time and thus the whole algorithm.
+This motivates the need for space efficient solutions to \texttt{Neighbourhood Detection}.
+
+\section{Motivating Applications} %TODO I don't really like this (partly because it is in list form), moreover it isn't really connected in
+% NOTE maybe it is best that it is fairly separate from the rest so that it can be skipped over.
+%\begin{itemize}
+%	\item Real world occurrences \& uses.
+%\end{itemize}
+
+There are several real-world applications of \texttt{Neighbourhood Detection} which motivate its investigation.
+These are situations where simply knowing highly connected elements of a network is not sufficient.
 \begin{itemize}
-	\item[-] Which items are most commonly bought \& which other items they are commonly bought with?
-	\item[-] Which resources in a network are being accessed most \& by whom?
+	\item[-] Given a list of connections in a social network, identify \textit{social influencers} and determine the demographics of their audience in order to decide who should promote a particular product.
+	\item[-] Given a list of receipts, identify which items are commonly sold together and use this information to determine how to stack shelves in order to increase sales.
+	\item[-] Given a log of traffic within a network identify which resources are being accessed most often, and by whom.
+This information can be used to determine what upgrades should be made to the network and to identify potential attacks on the network.
 \end{itemize}
-%TODO More
-Answering these questions can be used to inform resource allocation in order to maximise a certain goal. Be that profit; or, network response speeds.
-\section{Neighbourhood Detection Problem}
-% Describe Graph streams, insertion-only & insertion-deletion
-\par Graph Streams are a list of instructions for how to construct a graph and provide a dynamic way of representing graphs.
-In this paper, I shall discuss two types of graph streams: Insertion-Only Streams; and, Insertion-Deletion Streams.
-There are other types of graph streams that allow for additional operations such as updating vertex \& edge values, but these are not relevant to the problem being considering in this paper.
-\par Insertion-Only graph streams are a list of pairs of vertices, representing endpoints of unweighted-undirected edges.
-Each instruction describes a new edge to be inserted into the graph \& potential up to two new nodes which are in the graph.
-There is no way to remove or update an edge in an Insertion-Only graph stream.
-We apply a limitation that no duplicate edges appear in an insertion-only stream and thus the order of instructions in an Inserionly-Only stream has no effect on the validity of the algorithms which shall be discussed.
-\par Insertion-Deletion graph streams consist of a pair of vertices and a boolean.
-The vertices represent the endpoints of an unweighted-undirected edge \& the boolean describes whether this edge is being inserted or deleted from the graph.
-If it is an insertion instruction then we are adding the described edge to the graph; if it is a deletion instruction then we are removing it.
-Again we add a limitation that no edges are repeated \& that we will not receive a request to delete an edge from the graph which does not currently exist in the graph.
-This does, however, mean that the order of instructions in an Insertion-Deletion stream is important as the deletion of an edge must come after it's insertion.
-\par The limitations placed here often occur naturally in systems. On social networks you are not able to follow someone you are already following, nor unfollow someone you are not following.
-\par In spite of these limitations which are imposed it is still possible to construct almost any undirected-unweighted graph using the two graph streams described.
-Suppose you wish to analyse which devices in a network are making the most request.
-Here vertices represent devices \& edges represent requests.
-For this scenario, you would want to allow for multiple edges (say $n>2$) between the same pair of devices (say $\{x,y\}$).
-One solution would be to add $n$ new vertices $\{v_{x,y,1},\dots,v_{x,y,n}\}$ each with a label stating they are just intermediate nodes and not full devices.
-Creating edges from $x$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ \& from $y$ to each of $\{v_{x,y,1},\dots,v_{x,y,n}\}$ allows for an interpretation of multiple occurrences of the edge $(x,y)$.
-\par If a graph stream is ordered by time (which is common in real-life applications) then it represents a dynamic-evolution of the graph over time.
-Allowing the user to roll back the graph should they wish to make assessments during particular time periods, rather than just assess the final graph.
-\par The algorithms which shall be discussed in this paper are designed to work with just a single pass of a graph stream.
-This is not only time-efficient as it does not require multiple reads of the stream, but means that should more information be gained after the algorithm is run it is easy to incorporate it into the algorithm if we store the final state of the data structures used.
-Single-pass algorithms mean we don't necessarily have to store the graph stream in a local file, we could simply send instructions from the system being modeled straight into the algorithm.
-This is useful when a graph stream potentially could take up multiple terra-bytes, but is unnecessary in most applications and can mean it is harder to verify results later.
-\horizontalline
-% Build to problem
-\par The $\mathtt{High\ Degree\ Detection\ Probem}$ is a problem in graph theory where we seek to find a vertex of sufficiently high degree, in a given graph.
-This node does not necessarily have the greatest degree in the graph, and often we cannot be sure whether it does as the greatest degree is unknown beforehand.
-This problem can be formally stated as
+
+\section{Objectives}
+The objectives of this project are to:
+\begin{itemize}
+	\item Implement and understand the algorithms proposed in \cite{orig}.
+  \item Discuss and test alterations to the logic and implementation of the proposed algorithms, and the subroutines they rely upon, in order to improve the performance of the proposed algorithm.
+	\item Use large graphs from real-world scenarios to evaluate the practicalities of these implementations on real-world data sets.
+\end{itemize}
+
+\section{Structure}
+Having provide some background to the problem, the structure of this project is as follows.
+In \texttt{Chapter 1} I discuss and define general ideas, technologies and techniques which are used throughout this project.
+In \texttt{Chapter 2} I discuss, implement and evaluate the algorithm proposed in \cite{orig} for insertion-only graph streams.
+In \texttt{Chapter 3} I discuss, implement and evaluate the algorithm proposed in \cite{orig} for insertion-deletion graph streams.
+And, in \texttt{Chapter 4} review what was achieved in this project and provide some ideas for possible future work to build upon those achievements.
+
+\renewcommand\thechapter{\arabic{chapter}}
+\renewcommand\thesection{\thechapter.\arabic{section}}
+\setcounter{chapter}{0}
+\chapter{Preliminaries}
+
+\setcounter{page}{1} % restart page counter
+\pagenumbering{arabic}
+
+\section{Definitions}
+
+%\begin{itemize}
+%	\item Graph
+% \item Unweighted, undirected
+%	\item Bipartite Graph (Why this wasn't really applied in practice, A=B)
+%	\item Graph Stream (Insertion-Only, Insertion-Deletion)
+%	\item Neighbourhood, degree & sparsity of a vertex
+%	\item Streaming Algorithm
+%	\item $1-\varepsilon$ Approximation Algorithm
+%	\item TODO Space efficiency??
+%\end{itemize}
+
+A \textit{Graph} is a data structure used to represent pairwise relationships between objects.
+A Graph $G=(V,E)$ is defined to have a vertex set $V$, which holds the objects, and an edge-set $E$, which holds the relationship between the objects.
+Graphs are traditionally visualised with circles for each vertex and lines between vertices that share an edge.
+There are variations on graphs which allow for edges to have direction and weight.
+In this project we are using undirected, unweighted graphs meaning edges can be represented by an unordered pair of vertices $(u,v)$ for $u,v\in V$.
+
+\par A graph $G=(V,E)$ is said to be a \textit{Bipartite Graph} if its vertex set $V$ can be partitioned into two disjoint subsets $A,B$ and the every edges connects a vertex in $A$ to a vertex in $B$.
+Formally, $\exists\ A,B\subseteq V$ st $A\cup B=V,\ A\cap B=\emptyset$ and $\forall\ (a,b)\in E$ we have that $a\in A,b\in B$.
+\textit{Bipartite Graphs} are denoted as $G=(A\cup B,E)$.
+
+\begin{figure}[H]
+	\centering
+	\subfloat[Graph]{{\begin{tikzpicture}
+		\node[shape=circle] (v) at (0,0) {$v$};
+		\node[shape=circle] (w) at (0,-1) {$w$};
+		\node[shape=circle] (x) at (1,0) {$x$};
+		\node[shape=circle] (y) at (1,-1) {$y$};
+		\node[shape=circle] (z) at (2,-1) {$z$};
+
+		\path [-] (v) edge (w);
+		\path [-] (x) edge (y);
+		\path [-] (x) edge (z);
+	\end{tikzpicture}}}
+	\qquad
+	\subfloat[Bipartite Representation]{{\begin{tikzpicture}
+		\node[shape=circle] (A) at (-1,0) {$A$};
+		\node[shape=circle] (v) at (0,0) {$v$};
+		\node[shape=circle] (w) at (1,0) {$w$};
+		\node[shape=circle] (x) at (2,0) {$x$};
+		\node[shape=circle] (y) at (3,0) {$y$};
+		\node[shape=circle] (z) at (4,0) {$z$};
+		\node[shape=circle] (B) at (-1,-1) {$B$};
+		\node[shape=circle] (v') at (0,-1) {$v'$};
+		\node[shape=circle] (w') at (1,-1) {$w'$};
+		\node[shape=circle] (x') at (2,-1) {$x'$};
+		\node[shape=circle] (y') at (3,-1) {$y'$};
+		\node[shape=circle] (z') at (4,-1) {$z'$};
+
+		\path [-] (v) edge (w');
+		\path [-] (x) edge (y');
+		\path [-] (x) edge (z');
+		\path [-] (v') edge (w);
+		\path [-] (x') edge (y);
+		\path [-] (x') edge (z);
+	\end{tikzpicture}}}
+	\caption{Example of how all graphs have a bipartite representation}
+	\label{Figure 0}
+\end{figure}
+
+\par All graphs $G=(V,E)$ have a bipartite representation $G'=(A\cup B,E')$.
+Considering the example in \texttt{Figure 1}, defining $A=V$, $B$ to a be a copy of $V$, and the edge set \newline${E'=\{(u,v'):(u,v)\in E,u\in A,v'\in B\}}$ generates a bipartite presentation of $G$.
+The bipartite representation has twice the number of edges since each edge $(u,v)$ is now replaced by two edges $(u,v')$ and $(v,u')$.
+%TODO in practice we don't need these duplicates
+
+\par The \textit{Neighbourhood} of a vertex $v$ is the set of vertices which share an edge with, $N_v=\{u:(u,v)\in E\}$.
+The \textit{Degree} of a vertex $v$ is the size of its neighbourhood, $\delta_v=|N_v|$.
+A vertex $v$ is \textit{$s$-sparse} if its degree is less than, or equal to, $s$.
+\par The \textit{Edge Vector} of a vertex $v$ of length $|V|$ where each index gives the weight of the edge from $v$ to the vertex associated with that index.
+Typically a weight of 0 is stored if no edge exists between $v$ and a specific vertex.
+The graphs used in this project are unweighted so the values in the edge vector are booleans describing whether, or not, an edge exists.
+These edge vectors can be stored as bitstrings.
+\par \textit{Graph Streams} are an unbounded sequence of instructions $\{e_1,e_2,\dots\}$ which describe how to construct the graph.
+These instructions $e_i$ describe modifications to the edges of a graph.
+The vertex set of the described graph is inferred from the instructions.
+In this project we are only concerned with two types of graph stream.
+%TODO something about the unbounded part??
+\begin{itemize}
+	\item \textit{Insertion-Only Streams} where each instruction inserts a new edge to the graph.
+It is assumed that no edges are inserted multiple times.
+For the unweighted, undirected graphs used in this project insertion-only streams are a list of the edges in the graph and can be read in any order.
+Here the instructions have the form $e_i=(u,v)$ where $u,v\in V$ are the endpoints of the edge being inserted.
+	\item \textit{Insertion-Deletion Streams} where each instruction either adds a new edge to the graph or removes an existing edge from the graph.
+It is assumed that no edges are inserted multiple times, nor is an edge removed if it is not currently in the graph.
+For the unweighted, undirected graphs used in this project instructions of insertion-deletion streams take the form $e_i=(\Delta,u,v)$ where $u,v\in V$ are the endpoints of the edge and $\Delta$ is a boolean defining whether $e_i$ is an insertion or deletion instruction.
+The order of instructions is clearly important for insertion-deletion streams.
+\end{itemize}
+Other versions of graph streams exist which allow for changes to the weights of edges and directed edges, but these are out of scope for this project.
+
+\par \textit{Streaming Algorithms} are algorithms which are designed to take a stream of sequential instructions as their input.
+Streaming algorithms are space-efficient since only one instruction is read at a time, requiring constant space.
+Streaming algorithms which only require a single pass of the instruction stream are called \textit{Online Algorithms}.
+Online algorithms have the advantage of being able to take new instructions without having to recompute anything.
+
+\par An \textit{$\alpha$-Approximation Algorithm} is an algorithm which returns a result within an $\alpha$ factor to the optimal solution to the problem.
+Algorithms which solve the \texttt{Neighbourhood Detection Problem} are $\frac1c$-approximation algorithms since they only return $d/c$ of the neighbours of a vertex which we know to have at least $d$ neighbours.
+
+\par A family of hash functions $\mathcal{H}=\{h:X\to Y\}$ is described as being \textit{Pairwise Independent} if $\forall\ u,v\in X$ with $u\neq v$ and $\forall\ a,b\in Y$ if $h$ is chosen uniformly at random from $\mathcal{H}$ then $\mathbb{P}(h(u)=a\wedge h(v)=b)=\frac1{|Y|^2}$.
+This means that the values $u,v$ are hashed to are assigned uniformly at random and pairwise independently.
+
+\par A family of hash functions $\mathcal{H}=\{h:X\to Y\}$ is described as begin \textit{$k$-Wise Independent} if for any $k$ distinct elements of $X$ $\{v_1,\dots,v_k\}$ and any, not necessarily distinct, elements of $Y$ $\{a_1,\dots,a_k\}$ if $h$ is chosen uniformly at random from $\mathcal{H}$ then $$\mathbb{P}(h(v_1)=a_1\wedge\dots\wedge h(v_k)=a_k)=\frac1{|Y|^2}$$
+\section{Technologies}
+%\begin{itemize}
+%	\item C++
+%	\begin{itemize}
+%		\item Low level
+%		\item Memory manipulation
+%		\item Faster
+%		\item Perhaps something about why not java or python etc.
+%	\end{itemize}
+%	\item Standard library online
+%\end{itemize}
+For the implementations in this project I chose to use C++, specifically \texttt{C++11}.
+I initially considered using either python or C++, as they both work in the object-orientated paradigm and I am proficient in both.
+I chose C++ as it allows for more control with memory management and typically has faster run-times.
+For this project the memory management is more important as I am looking to evaluate the space-efficiency of algorithms.
+The faster run-times is important when evaluating the practicalities of the algorithms.
+\par I limited myself to using the \textit{C++ Standard Library} \texttt{std} as the underlying implementations of this library are well document.
+This was particularly important for the abstract data types I used as I could check what the underlying implementations and adjust my evaluation accordingly.
+When implementing randomness I used the \texttt{<random>} module of \texttt{std} and seeded the generators with the current time so that each run would have a different generator.
+
+\section{Evaluation Approach}
+
+% TODO expectations
+%\begin{itemize}
+%	\item What features are tested (time \& space)? How?
+%	\item Against naïve algorithm.
+%	\item Insertion-Only v Insertion-Deletion algorithm for insertion-only streams.
+%\end{itemize}
+
+%TODO talk about verifying my results
+%TODO typical test = vary c between [3,20]. set d=max degree
+
+During evaluation three performance metrics were measured: success-rate, time-taken, and, space-used.
+These are all interconnected and come with their own trade-offs.
+Typically, the changes to space-used \& time-taken when different strategies are implemented occur in the same direction, but at different rates.
+During most tests the success-rate was near perfect, but the other metrics were significantly higher as it took longer for a solution to be found.
+
+\par In real world scenarios we are most interested analysing the neighbourhoods of the most influential vertices.
+For this reason $d$ is set to the maximum degree in the graph during all tests.
+Further, the greater the proportion of a neighbourhood returned the better the inferences made are so I focussed testing on low values of $c$, namely $c\in[1,20]$.
+
+\par For this project I wanted some graphs generated from the real-world.
+I found a collection of large graph streams, from different applications, in the \texttt{Stanford Network Analysis Project} (\texttt{SNAP}) datasets collection \cite{SNAP}.
+I chose to use the graphs generated from social networks as I found that application most motivating, but this choice is essentially arbitrary.
+The graph streams from \cite{SNAP} were all insertion-only so I wrote a utility program which created insertion-deletion streams from them by using a Bernoulli random variable to decide whether to delete an edge immediately after inserting it.
+\textbf{Table 1} provides details about the graphs used during evaluation.
+
+\begin{center}
+	\begin{table}[h]
+		\tiny
+		\begin{tabular}{|l|l|l|l|l|l|l|l|}
+			\hline
+			\textbf{Name}&\textbf{Type}&\textbf{\# Vertices}&\textbf{\# Edges}&\textbf{\# Instructions}&\textbf{Max Degree}&\textbf{File Size}\\
+			\hline
+			$\mathtt{facebook\_small}$&Insertion-Only&52&146&-&18&3 KB\\
+			$\mathtt{facebook\_small\_deletion}$&Insertion-Deletion&52&131&161&16&5 KB\\
+			$\mathtt{facebook}$&Insertion-Only&747&30,025&-&293&587 KB\\
+			$\mathtt{facebook\_deletion}$&Insertion-Deletion&747&26,718&33,332&267&846 KB\\
+			$\mathtt{gplus}$&Insertion-Only&12,417&1,179,613&-&5,948&12 MB\\
+			$\mathtt{gplus\_deletion}$&Insertion-Deletion&12,417&1,049,309&1,309,917&4,998&16 MB\\
+			$\mathtt{gplus\_large}$&Insertion-Only&102,100&30,238,035&-&104,947&1.3 GB\\
+			\hline
+		\end{tabular}
+    \caption{Details of graphs used during evaluation}
+	\end{table}
+\end{center}
+
+\chapter{Insertion-Only Streams}
+
+\section{Degree-Based Reservoir Sampling}
+
+\texttt{Degree-Based Reservoir Sampling} is a technique for uniformly sampling a vertex from the set of vertices whose degrees are greater than a specified minimum bound $d_1$ along with $d_2$ of its neighbours.
+This is achieved by maintaining a sample of this set of vertices, known as the \textit{reservoir}, and storing the first $d_2$ of the edges incident to a vertex after it is sampled.
+The size of the reservoir is a parameter of the algorithm.
+\textbf{Algorithm 1} outlines pseudocode for performing \texttt{Degree-Based Reservoir Sampling} on a bipartite graph using an insertion-only graph stream.
+
+\begin{algorithm}
+	\caption{\texttt{Degree-Based Reservoir Sampling}$(d_1,d_2,s)$}
+	\SetKwInOut{Require}{require}
+	\Require{Degree bound $d_1\in\nats$, Neighbourhood bound $d_2\in\nats$, Reservoir size $s\in\nats$, Insertion-only stream $\{(a_0,b_0),\dots,(a_n,b_n)\}$}
+	$D\leftarrow\{\}$ \tcp*{Degree counter}
+	$R\leftarrow\{\}$ \tcp*{Reservoir}
+	$E\leftarrow\{\}$ \tcp*{Collected edges}
+	$x\leftarrow0$\tcp*{\# nodes of degree$\geq d_1$}
+	\For{$i\in[0,n]$} {
+		$D[a_i]\leftarrow D[a_i]+1$\tcp*{Increment degree counter}
+		\If{$D[a_i]\equiv d_1$}{
+			\tcp{Consider inserting $a_i$ to reservoir}
+			$x\leftarrow x+1$\\
+			\If{$|R|<s$} {
+				\tcp{Reservoir is not full}
+				$R\leftarrow R\cup\{a_i\}$
+			} \Else {
+				\tcp{Reservoir is full}
+				\If{\texttt{Bernoulli}$\left(\frac{s}x\right)$} {
+					Let $a'$ be a uniform random element of $R$\tcp*{Element to replace}
+					Delete edges in $E$ incident to $a'$\\
+					$R\leftarrow(R\backslash\{a'\})\cup\{a\}$\tcp*{Swap $a_i$ and $a'$}
+				}
+			}
+		}
+		\If{$a_i\in R$ \textbf{and} $D[a_i]< d_1+d_2$}{
+			$E\leftarrow E\cup(a_i,b_i)$\tcp*{Store edge}
+		}
+	}
+	\If{$\exists\ a\in R$ with $D[a]\geq d_1+d_2-1$} {
+		\Return{Uniform random vertex \& neighbourhood from those of size $d_2$}
+	} \lElse { \Return{\texttt{FAIL}}}
+\end{algorithm}
+
+\par Three data stores are used in \texttt{Degree-Based Reservoir Sampling}: a map of the degree of every node in the graph $D$; a set for the reservoir $R$; and a set for edges incident to the vertices in the reservoir $E$.
+
+\par The reservoir has an invariant that at any point in time it contains a uniform sample of the vertices whose degrees are known to be greater than $d_1$.
+This invariant is proved as \textbf{Lemma 2.1}, first I shall describe how the reservoir is maintained by controlling how vertices are inserted.
+The first time the degree counter for a vertex $v$ surpasses $d_1$, $v$ is considered for insertion into the \textit{reservoir}.
+There are two cases
+\begin{itemize}
+	\item If the reservoir is \underline{not} full (\ie $|R|<s$) then $v$ is inserted into the reservoir.
+	\item Otherwise, if the reservoir is already full, a Bernoulli random variable with probability $p=\frac{x}s$ is used, where $x$ is the number of vertices known to have degree at least $d_1$.
+If this random variable succeeds then pick, uniformly at random, a vertex $u$ currently in the reservoir and replace it with $v$.
+All the edges in $E$ which are incident to $u$, and no other vertex in the reservoir, are removed from $E$.
+\end{itemize}
+
+Whenever an edge $(u,v)$ is encountered and $u$ is in the reservoir and the number of edges stored in $E$ incident to $u$ is less than $d_2$ then $(u,v)$ is inserted into $(u,v)$. %TODO rewrite this
+\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+
+\textbf{Lemma 2.1} At any time $t$, $R$ contains a uniform sample of the vertices whose degrees are known to be at least $d_1$.\\
+
+\textit{Proof (by induction)}\\
+Let $X_t$ be the set of vertices whose degrees are know to be greater than $d_1$ after $t$ instructions have been consumed and $x_t:=|X_t|$.\\
+\underline{Base Case} - $x_t\leq s$.\\
+Trivially true since $R$ is not yet full so $R=X_t$.\\
+\underline{Inductive Case} - $x_t>s$.\\
+Assume the property holds for $x_t=z>s$, then the probability of any given element of $X_t$ is in $R$ is $\frac1z$.
+Without loss of generality, let $a$ be the next element considered for sampling and $b$ be any element of $R$.
+The probability $a$ is sampled is $\frac{s}{z+1}$ and, given $a$ is to be sampled, the probability $b$ is chosen for removal is $\frac1s$.
+Thus, the probability that $b$ is replaced by $a$ is $\frac{s}{z+1}\cdot\frac{1}{s}=\frac1{z+1}$.
+Similarly, the probability $b$ is \underline{not} replaced by $a$ is $1-\frac1{z+1}=\frac{z}{z+1}$.
+By the inductive hypothesis, the probability that $b$ was in $R$ was $\frac1z$.
+Thus, the probability that $b$ is still in $R$ is $\frac1z\times\frac{z}{z+1}=\frac1{z+1}$.\\
+
+Hence by mathematical induction, $R$ contains a uniform sample of the vertices of sufficient degree at all points in time.$\hfill\square$
+}}\end{adjustwidth}\vspace{.3cm}
+%TODO adding termination condition? Maybe add during implementation
+
+\par If $v$ is not added to the reservoir at this time, it never will be in the future.
+Note that in the case that the reservoir is full and $v$ is inserted into it, $v$ is replacing an element that was in the reservoir meaning any progress made towards finding a neighbourhood for that vertex is annulled.
+The amount of progress is not taken into account \& thus if there is a high turn-over of elements in a reservoir then less progress is made towards finding a neighbourhood for each of them.
+This is discussed further during evaluation.
+
+\par \texttt{Degree-Based Reservoir Sampling} only works on insertion-only streams as it does not account for when the degree of a vertex falls below the threshold, after being sampled.
+This means the reservoir is not necessarily a uniform random sample of the vertices with degree $\geq d_1$.
+%TODO is this sufficient
+
+\par For this project we are interested in how \texttt{Degree-Based Reservoir Sampling} can be implemented for bipartite graphs, specifically we want to sample from the $A$-vertices of the graph.
+For a bi-partite graph with $n$ $A$-vertices \texttt{Degree-Based Reservoir Sampling} requires $\mathcal{O}(n\log n+sd_2\log n)$ space, assuming $O(\log n)$ space is required to store an edge or a vertex, since it stores a degree counter for every $A$-vertex and at most $d_2$ edges for each of the $s$ vertices in the reservoir.
+
+\section{Proposed Algorithm}
+
+%\begin{itemize}
+%	\item Assumptions: no duplicate edges (can mitigate cases when this would be required by being smart)
+%	\item Insertion stream neighbourhoods cannot be diminished over time.
+%	\item Possibility of failure (missed edges) \& mitigation
+%	\begin{itemize}
+%		\item Theorem 3.2 (probs not proof)
+%	\end{itemize}
+%	\item Theoretical space requirements
+%\end{itemize}
+
+The \texttt{Neighbourhood-Detection Problem} can be solved using \texttt{degree-based reservoir sampling} sampling by setting $d_2=\frac{d}c$.
+The question remains as to what to set $d_1$ and $s$ to in order to achieve a high probability of success.
+%in little time and space
+
+\par \textbf{Algorithm 2} is the algorithm proposed in \cite{orig} for solving the \texttt{Neighbourhood-Detection Problem} for insertion-only streams.
+
+\begin{algorithm}
+	\caption{One-pass $c$-Approximation Insertion-Only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
+	\SetKwInOut{Require}{require}
+	\Require{Space $s$, degree bound $d$.}
+	$s\leftarrow\lceil\log(n)\cdot n^{\frac1c}\rceil$\\
+	\For{$i\in[0,c-1]$\text{\textbf{ in parallel}}} {
+		$(a_i,S_i)\leftarrow\mathtt{Deg}\mbox{-}\mathtt{Res}\mbox{-}\mathtt{Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
+	}
+	\Return{\small Uniform random neighbourhood $(a_i,S_i)$ \text{from successful runs}}
+\end{algorithm}
+
+\par In \cite{orig} it is proven that setting $s=\lceil\ln(n)\cdot n^{\frac1c}\rceil$ and using $c$ \texttt{degree-based reservoir samplers} each with a different lower bound, incremented from $\frac{d}c$ to $d$ stepping by $\frac{d}c$ each time, results in a high probability of success.
+The samplers are run in parallel so that the algorithm only requires a single pass of the stream.
+Implementing parallel running of the samplers just requires passing the same instruction to each sample before fetching the next instruction.
+
+\par Each of the $c$ samplers requires $\mathcal{O}(n\log n+sd_2\log n)$ space.
+Since the degree map can be shared between samplers the total space is $\mathcal{O}(n\log n+c\cdot sd_2\log n)=\mathcal{O}(n\log n+c\lceil\log(n)n^{\frac1c}\rceil\frac{d}c\log n)=\mathcal{O}(n\log n+n^{\frac1c}d\log^2n)$.
+The proposed algorithm does require you to know the number of vertices in the graph stream before running the algorithm, or at least the $\log$ of it.
+If this is unknown then it is quick to run through the whole stream \& build a set of vertices in the graph.
+This requires $\mathcal{O}(n\log n)$ space, assuming $O(\log n)$ space is required to store a vertex.
+However, in many real-world scenarios this number (or a good approximation) will be known due to it being important for other tasks.
+
+\section{Implementation}
+%TODO justify changes wrt gplus graph only, include plots as I go
+\par I centred my implementation around working for the graph streams from \textit{SNAP} \cite{SNAP} as they came from real-world sources \& thus are very important for testing the real-world practicalities of the proposed algorithm.
+These graph streams are not formatted in a strictly bipartite way.
+I adjusted \texttt{degree-based reservoir sampling} to account for this by by repeating \texttt{lines 6-17} (and the early termination clause) of \textbf{Algorithm 1} after the end of the \texttt{for} loop, with all occurrences of $a_i$ replaced by $b_i$.
+%TODO something about implementing in parallel. Nah
+\par I tested changes to the implementation on multiple graphs, for brevity and succinctness I shall only show the results for \texttt{gplus}.
+\texttt{Figure 2.3.1} shows the results of testing the initial implementation of \textbf{Algorithm 2}.
+These results shows that the space requirements of \textbf{Algorithm 2} are significantly less than the size of file stream and that the space requirements decrease as the approximation factor increases but there is an undesirable result that the time taken increases as the approximation factor increases.
+\begin{figure}[H]
+	\label{Figure 2}
+	\includegraphics[width=.5\textwidth]{img/gplusInitialTime.png}
+	\includegraphics[width=.5\textwidth]{img/gplusInitialSpace.png}
+	\caption{Results from testing initial implementation of \textbf{Algorithm 2} on \texttt{gplus} graph for different approximation factors}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\noindent After this initial implementation I considered two alterations which could be made:
+\begin{enumerate}
+	\item \textit{Early Termination} - Returning the first neighbourhood of sufficient size, rather than uniformly sample from all which succeed; and,
+	\item \textit{Shared Edge-Set} - Having all samplers share an edge-set, rather than each have one each.
+\end{enumerate}
+
+\textbf{Algorithm 1} states that the whole stream should be run through before the sampler returns a result uniformly from those which succeeded.
+This is unnecessary to fulfil \texttt{neighbourhood detection} as returning the first successful result would be sufficient.
+Similarly \textbf{Algorithm 2} states that all samplers should be run to completion and then a result is returned from those which succeeded.
+Again, this is unnecessary as return the first encountered neighbourhood of sufficient size would suffice for solving \texttt{neighbourhood detection}.
+These early terminations can be implemented by terminating the algorithm after it finds a vertex of degree $d_1+d_2-1$ which is in a reservoir (remembering that $d_1$ is different for each reservoir).
+
+\par Implementing early termination should reduce run-time as only part of the stream is being consumed and should reduce space used by the edge-sets in reservoir sampling as no data is stored after the first solution is found.
+As this change terminates at the first success there is no change to the success rate of the algorithm.
+One down side is that you will likely get less variety in the returned neighbourhoods as the algorithm will not encounter every vertex which meets the degree requirements, however this is not relevant to the problem.
+
+\begin{figure}[H]
+	\label{Figure 3}
+	\includegraphics[width=.5\textwidth]{img/gplusEarlyTerminationTime.png}
+	\includegraphics[width=.5\textwidth]{img/gplusEarlyTerminationSpace.png}
+	\caption{Results when using early termination, tested on \texttt{gplus} graph for different approximation factors. Against the results for the initial implementation.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par \texttt{Figure 2.3.2} shows the results implementing early termination, against not implementing it, for graph \texttt{gplus} at different approximation factors.
+The time taken now decreases as $c$ increases and is more than halved in all cases.
+The space-requirements are down in every case and are down more than $90\%$ for reasonable values of $c$.
+Both metrics now converge quickly, to $\smallsim\!\!1\frac1.5$ seconds and $3.3$MBs respectively, meaning we can make predictions on the performance of the algorithm for greater values of $c$.
+These are really good results, so I implemented this change going forward.
+%TODO explain reasons behind these results
+%NOTE space converged to is that used by degree map
+%TODO comment about effect on space used by reservoirs vs degree map
+%TODO this change means that time taken is now a measure of success as failure is (basically) the only time the whole stream is run through
+
 \vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
-\textbf{Problem 1} $\texttt{High Degree Detection}$.\\
-Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A\ \&\ B$, where $|A|=n$ and $|B|=\text{poly }n$ , and edge set $E$.
-Set a restriction that at least one vertex in $A$ has degree, at least, $d$.\\
-In $\texttt{High Degree Detection}(G,d)$ we are tasked with outputting a vertex from $A$ with degree, at least, $d$.
-}}\end{adjustwidth}\vspace{.3cm}
-\par This problem can be solved efficiently for both insertion-only \& insertion-deletion graph streams with a fairly simple algorithm which simply runs through all the edges, incrementing or decrementing (depending on whether it is an insertion or deletion instruction)  a count for the degree of each vertex as it goes.
-Returning the first vertex it finds with degree equal to $d$.\\
-%TODO Do I have to justify the algorithm
-\begin{algorithm}[H]
-\caption{Single Pass High Degree Detection}
+
+\textbf{Lemma 2.2} If \textbf{Algorithm 2} terminates after the first sufficient neighbourhood it finds, no vertex occurs in multiple reservoirs at the same time.\\
+
+% TODO a proof by contradiction
+\textit{Proof (by contradiction)} Let $v$ be a vertex in the graph stream.
+Suppose, without loss of generality, that $v$ is a member of the reservoirs of both \texttt{Deg-Res-Sampling}$(\frac{d}{c},\frac{d}{c},s)$ \& \texttt{Deg-Res-Sampling}$(\frac{2d}{c},\frac{d}{c},s)$.
+This means that $v$ was sampled twice, after encountering the $\frac{d}c^\text{th}$ \underline{and} $\frac{2d}c^\text{th}$ edge incident to $v$ in the stream.
+After sampling $v$ for the first time, encountered edges which are incident to $v$ are stored.
+This means that by the time $v$ is sampled for the second time $\frac{2d}c-\frac{d}c=\frac{d}c$ edges incident to $v$ have been stored.
+This is a sufficient neighbourhood for \textbf{Algorithm 2} to terminate.
+This is a contradiction since termination occurs before the second sampling occurs.$\hfill\square$
+
+}}\end{adjustwidth}\vspace{.3cm} %TODO refine
+
+\par A consequence of \textbf{Lemma 2.2} is that an edge is only stored in multiple edge sets if its endpoints are sampled by different samplers. %TODO mention this is rare
+This means that introducing a shared edge-set for all samplers will have little effect on the space requirements of \texttt{Algorithm 2}.
+\par However, the alterations to the implementation which would have to made to allow for a shared edge-set could introduce time savings.
+Namely, I would need a data-structure which tells me \underline{precisely} which reservoir each vertex is in that so I know the appropriate value of $d_1$ in \texttt{Algorithm 1:Line 18}, otherwise improper termination could occur.
+An appropriate data-structure for this would be a \texttt{map} from each vertex to a bitstring whose indices indicate whether the vertex lies in an associated reservoir.
+With this implementation, inserting or deleting a vertex to/from a reservoir and querying whether a vertex is in a given reservoir becomes querying a map \& then a known bit in a bitstring which both take $\mathcal{O}(1)$ time.
+This \texttt{map} replaces the \texttt{ordered set}s currently used to store the reservoirs of each sampler and thus reduce these action times from $O(\log n)$ to $\mathcal{O}(1)$ time.
+As these actions occur a lot this change should result in a reasonable reduction in run-time.
+% TODO something about space used to store a vertex in the map vs in sets
+
+\begin{figure}[H]
+	\label{Figure 4}
+	\includegraphics[width=.5\textwidth]{img/gplusSharedEdgeSetTime.png}
+	\includegraphics[width=.5\textwidth]{img/gplusSharedEdgeSetSpace.png}
+	\caption{Results when using early termination and a shared edge set, tested on \texttt{gplus} graph for different approximation factors. Against the results just using early termination.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par \texttt{Figure 2.3.3} shows the results when implementing both early termination and a shared edge set, against only using early terminiation.
+These tests showed a consistent reduction in run-time when both strategies are used.
+there is a slight increase in space usage for low values of $c$ but this difference disappears very quickly so is of little concern.
+Implementing a shared-edge set does not change the logic of the algorithm so does not affect its success rate.
+Due to the reduction in mean run-time \& no noteworthy changes in space requirements, I recommend implementing this change.
+
+%TODO discuss performance against theory
+\section{Parameter Tuning}
+
+%\begin{itemize}
+%	\item Vary size of reservoir ($s$) \& number of samplers together to find an optimum
+%	\item Full optimised against v Naïve.
+%\end{itemize}
+
+\par The changes discussed above have been to the technical implementation of the proposed algorithm, rather than to the its logic.
+This means these changes have had no affect on the success rate of the algorithm.
+I will now discuss potential changes to the logic of the proposed algorithm.
+These changes will affect on the success rate of the algorithm.
+It is possible that no change in the success-rate will be seen \& rather the change will occur in execution time as more instructions have to be read before a solution is found.
+I am investigating the practicalities of the algorithm so keeping the run-times usable is a high priority.
+In \texttt{Algorithm 2} there are two parameters which can be played around with: the number of samplers; and, the size of the reservoirs, $s$.
+
+\par The total reservoir size, $\left\lceil\ln(n)n^{\frac1c}\right\rceil\times\text{\#\ Samplers}$, is the main factor effecting the space-used by the algorithm as this is the number of vertices it needs to store edges for.
+Noting that \texttt{Algorithm 2} sets the number of samplers to be $c$ and that $cp\left\lceil\ln(n)n^{\frac1c}\right\rceil\leq c\left\lceil p\ln(n)n^{\frac1c}\right\rceil\ \forall\ p\in[0,1]$ it is apparent that reducing the number of samplers has a greater effect on total reservoir size, and thus space-used, than reducing reservoir size by the same proportion.
+For this reason I investigated the number of samplers first.
+
+\subsection{Number of Samplers}
+
+\par When changing the number of samplers we need to consider their lower bounds, $d_1$, too.
+If the range of vertex degrees which individual samplers sample, $[d_1,d_1+d_2]$, at overlap then duplicate sampling of vertices can occur which offers no advantages \& leads to a reservoir space being wasted.
+And, if there are gaps between these ranges then there are vertices which may not be sampled.
+%TODO flesh this out more.
+Thus, the best set up is to have the sampling ranges be sequential with no gaps and so when varying the number of samplers we should only add or remove from the endpoints of these ranges.
+The samplers with lower lower-bounds, $d_1$, start sampling earlier \& from a larger pool.
+I expect these early samplers to be more important wrt performance gains, than later samplers.
+It is not necessarily true that the first sampler is the most important as it is sampling from the largest pool and so has the greatest turn-over rate in its reservoir.
+
+\par I ran the following tests to investigate what had a greater effect on performance, removing later samplers or earlier samplers.
+\begin{enumerate}
+	\item Removing the \underline{first} $x$ samplers from \texttt{Algorithm 2} for $x\in[1,c)$, for $c\in[1,20]$.\\
+	\textit{i.e.} Change the \texttt{for} loop in \texttt{Algorithm 2:Line 2} to run for $i\in[x,c-1]$.
+	\item Removing the \underline{last} $x$ samplers from \texttt{Algorithm 2} for $x\in[1,c)$, for $c\in[1,20]$.\\
+	\textit{i.e.} Change the \texttt{for} loop in \texttt{Algorithm 2:Line 2} to run for $i\in[0,c-x]$.
+\end{enumerate}
+
+\begin{figure}[H]
+	\label{Figure 5}
+	\subfloat[Removing early samplers]{{\includegraphics[width=.5\textwidth]{img/gplusRemoveEarly.png}}}
+	\subfloat[Removing late samplers]{{\includegraphics[width=.5\textwidth]{img/gplusRemoveLate.png}}}
+	\caption{Results when removing different proportions of samplers, either from the start or end of the sampler range, for different values of $c$.
+These results are normalised with the values when no samplers are removed being 1 for that value of $c$.
+Results from graph \texttt{gplus}.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par \texttt{Figure 2.4.4} shows the results of tests, with the results normalised to make them comparable.
+These results confirm the hypothesis that the lower a sampler's lower-bound is, the greater the effect of its inclusion on performance.
+Time-taken increased as more early samplers were removed, and decreased as more late samplers were removed.
+The space used was not affected much by the removal of samplers, although there is a slight increase in space usage when more early samplers were removed.
+Comparing the normalised results shows that removing early samplers is more detrimental to performance than including late samplers.
+This is due to late samplers only starting to sample later \& they may not have begun sampling by the time one of the previous samplers has succeed, terminating the algorithm.
+When the same proportion of samplers were removed for different values of the  $c$, the relative performance was worse for greater values of $c$.
+This is to be expected as the real number of samplers removed increases with $c$.
+
+\begin{figure}[H]
+	\label{Figure 6}
+	\centering
+	\includegraphics[width=.5\textwidth]{img/gplusSuccessRate.png}
+	\caption{Success rate when only the first $n$ samplers are run on \texttt{gplus}.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par From these tests it is apparent that there are advantages to be found by removing late samplers, but not from removing the earliest.
+These tests did show that the failure rate does increase when too few of the early samplers are used.
+Thus I repeated the tests, this time testing using the first $n$ samplers for different values of $c$.
+\texttt{Figure 2.4.5} shows the results of this test and shows that the success rate dropped below 1 in the case that only the first sampler was used, in all other cases it was 1.
+
+\par Running the first two samplers provided the best results.
+However, the largest graph I tested on only contains $~100,000$ vertices so this may not hold for significantly larger graphs.
+I suggest using the first $\mathtt{max}\left(2,\left\lceil\frac15\ln n\right\rceil\right)$ samplers.
+As we don't want the sampling ranges to overlap we need to add a restriction that the number of samplers is never greater than the approximation factor, giving the number of samplers used as $\mathtt{min}\left(c,\mathtt{max}\left(2,\left\lceil\frac15\ln n\right\rceil\right)\right)$.
+
+\begin{figure}[H]
+	\label{Figure 7}
+	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/gplusRemovedSamplersTime.png}}}
+	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/gplusRemovedSamplersSpace.png}}}
+	\caption{Results when using a reduced number of samplers and the technical improvements, tested on \texttt{gplus} graph for different approximation factors, against just using the technical improvements.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par \texttt{Figure 2.4.6} shows the results when using the first $\mathtt{min}\left(c,\mathtt{max}\left(2,\left\lceil\frac15\ln n\right\rceil\right)\right)$ samplers against just using the technical improvements discussed in \texttt{Section 2.3}.
+As seen in \texttt{Figure 2.4.4}, the space usage is unchanged while the time taken is down in all cases.
+Run time is reduced by over $50\%$ for $c\geq 12$ and is below 1 second for larger values of $c$.
+A perfect success rate was maintained throughout these tests.
+Due to these good results, I implemented this change.
+%TODO something about how the fact that space hasn't changed much means later samplers weren't coming into play often & thus sample size is unlikely to be important.
+%TODO comment of special cases
+\subsection{Reservoir Size}
+
+\par \texttt{Figure 2.4.4 (b)} showed that removing late samplers had little effect on the space used by the algorithm.
+This is due to these samplers not getting the chance to sample very often as the algorithm often terminated before many vertices of sufficient degree were encountered.
+This means that reducing the size of those samplers would have had virtually no affect on the performance of the algorithm.
+This helps justify optimising the number of samplers before the sample sizes.
+The previous optimisation means that very few samplers are being run so it is unlikely that any great improvements will be found, if any, when reducing reservoir size.
+
+\par Decreasing the reservoir size $s$ means less edges are stored, reducing space requirements, but increases the turn-over rate of the elements in the reservoir as the probability a vertex is sampled is the same but the reservoir size is smaller.
+The greater turn-over rate is detrimental to the success of the algorithm as it more likely for a vertex to be removed when it is close to succeeding.
+For very low sample sizes it is likely that the success rate of the algorithm will remain high, but the run-times will be dramatically increased.
+
+\begin{figure}[H]
+	\label{Figure 8}
+	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/gplusReservoirSizeTime.png}}}
+	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/gplusReservoirSizeSpace.png}}}
+	\caption{Normalised results when varying the reservoir size used on \texttt{gplus} for different approximation values.
+Results normalised by the result when the proposed reservoir size was used for each approximation valuee.
+(Using changes from \texttt{Section 2.3} and a reduced number of samplers)}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par \texttt{Figure 2.4.7} shows the results when varying the reservoir size between $10\%$ and $200\%$ of the proposed size, for different values of $c$.
+The results are normalised to make them comparable across values of $c$.
+There is a positive linear relationship between space usage and reservoir size.
+The only inference that could be drawn from sample-size and time-taken is that there is a point at which time-taken dramatically increases, this is the point where the algorithm fails to find a sufficient neighbourhood.
+Overall, there is no clear strategy to improving the algorithm by varying the reservoir size.
+
+%TODO mention theoretical space efficieny of this (should be same as before)
+\par \textbf{Algorithm 3} gives \textbf{Algorithm 2} rewritten to include the proposed parameter optimisations.
+\begin{algorithm}
+	\caption{One-pass $c$-Approximation Insertion-Only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
 	\SetKwInOut{Require}{require}
-\Require{Stream $\{(i_0,x_0,y_0)\dots(i_n,x_n,y_n)\}$, degree bond $d$}
-$D\leftarrow\{\{\}\}\ \{\text{degree map}\}$\\
-\For{$j\in[0\dots n]$} {
-	\uIf{$x_j\in D$}{
-		\lIf{$i_j$}{$D[x_j]+=1$}
-		\lElse{$D[x_j]-=1$}
+	\Require{Space $s$, degree bound $d$.}
+	$s\leftarrow\lceil\log(n)\cdot n^{\frac1c}\rceil$\\
+	$n_s\leftarrow\min\left(c,\max\left(2,\left\lceil\frac15\ln(n)\right\rceil\right)\right)$\\
+	\For{$i\in[0,n_s]$\text{\textbf{ in parallel}}} {
+		$(a_i,S_i)\leftarrow\mathtt{Deg}\mbox{-}\mathtt{Res}\mbox{-}\mathtt{Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
 	}
-	\lElse{$D[x_j]=1$}
-	\lIf{$D[x_j]=d$}{\Return{$x_j$}}
-	\textbf{repeat} with $y_j$
-}
-\Return{FAIL}
-\end{algorithm}\vspace{.3cm}
-In \textbf{Algorithm 1} $i$ is a boolean which is true if the edge is an insertion edge, and false if it is a deletion edge.
-$x\ \&\ y$ are the two endpoints of the edge.
-The algorithm makes the reasonable assumption that you will never recieve a deletion-instruction for an edge which is not in the graph, nor an insertion-instruction for an edge which is already in the graph.
-\par The $\mathtt{High\ Degree\ Detection\ Probem}$ is limited in its applications as it only returns a vertex with high degree \& gives you no information as to its neighbourhood.
-Solving this problem would not be very helpful when it comes to identifying influencers as it would only return people who have large followings, but not who their followers are.
-Thus an advertisier would not be able to evaluate whether the returned person's followers represented the audience they wished to target.
-\par The $\mathtt{Neighbourhood\ Detection\ Problem}$ is an extension of the $\mathtt{High\ Degree}$ $\mathtt{Detection\ Problem}$.
-The task is to return a vertex with sufficiently high degree \& a certain proportion that vertex's neighbourhood.
-This is formally stated as
-\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
-%TODO Shorten this, make it neater.
-\textbf{Problem 2} $\texttt{Neighbourhood Detection}$.\\
-Let $G=(A\cup B,E)$ be a bi-partite graph with vertex sets $A,B$, where $|A|=n$ and $|B|=\text{poly }n$, and edge set $E$.\\
-In $\texttt{Neighbourhood Detection}(G,d,c)$ we are tasked with outputting a vertex from $A$ and at least $d/c$ of its neighbours in $B$.\\
-Here $d$ is a threshold parameter \& $c$ is an approximation parameter.
-}}\end{adjustwidth}\vspace{.3cm}
-\par Solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ allows for a more general assessment of the influence of the returned vertex.
-In the context of finding social media influencers the returned neighbourhood can be assessed to determine whether the returned person is suitable for a given advertising campaign.
-When assessing the members of the neighbourhood you could look at the demographics of the members \& how active the members are, among other features, in order to determine whether they fit the target audience of the campaign.
-For cases where $c\neq 1$ you only get a subset of a person's whole neighbourhood but for sufficiently high $d$ you an implement statistical tools to still be able to draw meaningful inferences.
-Later it shall be shown that there are meaningful advantages to increasing the value of $c$.
+	\Return{\small Uniform random neighbourhood $(a_i,S_i)$ \text{from successful runs}}
+\end{algorithm}
 
-\section{Why Space Efficiency?}
-Simply solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ is fairly trivial.
-\textbf{Algorithm 2} is an algorithm for solving the $\mathtt{Neighbourhood\ Detection\ Problem}$ for insertion-only graph streams by recording the vertices in the neighbourhood of every $A$-vertex and then returning the first encountered neighbourhood with $\frac{d}{c}$ members.\\
-\begin{algorithm}[H]
-\caption{Na\"ive Single-Pass Insertion-Streaming Algorithm for Neighbourhood Detection}
+%TODO rewrite algorithm 2 with these parameter optimisations
+\section{Evaluation}
+
+%Run the following instances on different graphs \& compare
+%\begin{itemize}
+%	\item Initial implementation (no tuning, not quit early, not sharing)
+%	\item Implementation with early quitting & shared edges
+%	\item Implementation with tuned algorithm parameters
+%	\item Naïve Implementation
+%\end{itemize}
+% Compare max space to size of edges file
+% Identify which change had what effect (to space, time, success)
+%TODO 1000star is useful as there is only one valid solution
+
+\begin{algorithm}
+	\caption{Na\"ive Single-Pass Insertion-Only Streaming Algorithm for Neighbourhood Detection}
 	\SetKwInOut{Require}{require}
@@
 	\For{$i=0\dots n$} {
-	append $t_i$ to $N[s_i]$\\
-	\lIf{\text{size}($N[s_i])= \frac{d}{c}$} {
-		\Return{($s_i$,$N[s_i]$)}
+		append $t_i$ to $N[s_i]$\tcp*{Insert neighbour}
+		\If{\text{size}($N[s_i])\geq d$} {
+			\tcp{Sufficiently large neighbourhood found}
+			$S\leftarrow \{N[s_i][0],\dots,N[s_i][\frac{d}{c}]\}$\tcp*{First $\frac{d}c$ elements of neighbourhood}
+			\Return{($s_i,S$)}
 		}
+	}
+	\Return{FAIL}\tcp*{No neighbourhoods found}
+\end{algorithm}
+
+\textbf{Algorithm 4} is a naïve algorithm for solving \texttt{neighbourhood detection} for insertion-only graph streams.
+This approach records the neighbourhood of every encountered vertex and returns the first one which surpasses the $\frac{d}c$ threshold.
+This algorithm will always succeed as there is a requirement that at least one vertex in the graph is of degree, at least, $d$.
+As the naïve algorithm terminates at the first neighbourhood of sufficient size and uses all edges in the stream up to that point, it is guaranteed to find the first vertex with a sufficient neighbourhood.
+This means the naïve algorithm is very quick, provided its space-usage is within the memory allocated to the algorithm.
+In the worst case, where all elements have degree $\frac{d}c-1$ except for the last which has degree $\frac{d}c$, $\frac{d}{c}+(n-1)(\frac{d}c-1)\in \mathcal{O}(n\frac{d}c)$ space is required.
+The average case has the same complexity and the best case, where the first $\frac{d}c$ instructions are incident to the same vertex, has $O(\frac{d}c)$ space-complexity.
+
+To evaluate my final implementation I compared its performance to that of different stages of the optimisation and to the naïve implementation given in \textbf{Algorithm 4}.
+These different stages were
+\begin{enumerate}[label=\roman*)]
+	\item Initial Implementation - No early termination, shared edge-set or sampler removal. %insertionStreamsNotQuitEarly
+	\item After Technical Optimisation - Implements early termination \& a shared edge-set. %insertionStreamsSharedEdgeSet
+	\item After Algorithmic Optimisation - Implements early termination, shared edge-set and sampler removal. %insertionStreamsRemovedSamplers
+\end{enumerate}
+
+% TODO a table of results??
+\begin{figure}[H]
+	\label{Figure 12}
+	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusTime.png}}}
+	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusSpace.png}}}
+	\caption{Time taken and space used by different implementations of the proposed algorithm for insertion-only streams when tested on \texttt{gplus}}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par \texttt{Figure 2.5.8} shows that the space usage was greatest for the naïve algorithm and lowest for the final implementation.
+The shortest run times came from the naive algorithm, followed by the final implementation.
+However, after three hours of computation the naïve algorithm had failed to return a result for the largest graph \texttt{gplus\_large} due to its high space usage.
+For the largest graph tested, \texttt{gplus\_large}, $13.7\%$ of the stream file size was required to find a 2-approximation using the initial implementation and only $1.39\%$ using the final implementation.
+This is a notable decrease in space requirements, meaning the algorithm can run efficiently on even large graphs.
+
+\par The run time of the final implementation is useably quick: when run on \texttt{gplus\_large} it took $\sim\!\!150$ seconds to find a solution when $c=2$ and less than a minute for $c\geq5$.
+The naïve algorithm being quicker is irrelevant for this evaluation as it fails to terminate for larger files due to its space inefficiency.
+
+\chapter{Insertion-Deletion Streams}
+
+\section{$L_0$ Sampling}
+
+%\begin{itemize}
+%	\item L0 sampling uniformly samples from the non zero elements of a vector.
+%	\item Break down into 1-sparse \& $s$-sparse recovery. (COMMENT ON SPACE USAGE)
+%	\begin{enumerate}
+%		\item Perfect $1$-sparse recovery uses a series of counters to determine whether a vector is exactly 1-sparse (ie exactly one non-zero element). If it is then it returns value of that index.
+%		\item Give proofs that it works?
+%		\item NOTE - if vector could have negative elements then other checks are require (in this usage they do not so don't worry)
+%	\end{enumerate}
+%	\begin{enumerate}
+%		\item $s$-sparse uses pairwise hash function (explain these (maybe in preliminaries)).
+%		\item $s$-sparse recovers an $s$-sparse (ie upto $s$ non-zero indices) vector of the passed vector.
+%		\item Each index is considered by $r$ 1-sparse samplers.
+%		\item NOTE $s$-sparse recovery space grows with $s$ so we want to use a small value (idea used during $L_0$ sampling)
+%	\end{enumerate}
+%	\item The formally proposed algorithm.
+%	\item In this setting we consider the edge vector of each vertex (indicates which vertices it shares an edge with).
+%\end{itemize}
+
+%TODO citations
+\par\texttt{$L_0$ Sampling} is a technique for sampling, approximately uniformly, from the non-zero indices of a vector when given a stream of updates to the weights of the indices.
+Here I shall describe one space-efficient technique for \texttt{$L_0$ Sampling} which succeeds with probability $1-\delta$ and requires $\mathcal{O}\left(\frac1\delta\log(1/\delta\gamma)\log^2(n)\right)$ space, this is significantly less than the $\mathcal{O}(n)$ space required to store the whole vector.
+This technique relies on two subroutines: \texttt{Perfect 1-Sparse Recovery}; and \texttt{Exact $s$-Sparse Recovery}.
+%TODO check space requirements (do I need to mention the two different success rates)
+
+\subsection{Perfect 1-Sparse Recovery}
+
+\par \texttt{Perfect 1-Sparse Recovery} takes a stream of updates and if the described vector is exactly $1$-sparse then it returns the index of the non-zero element, otherwise the routine fails.
+\textbf{Algorithm 5} provides an implementation \texttt{Perfect 1-Sparse Recovery} which works in cases where the weight of the non-zero element is non-negative.
+The non-negativity constraint is acceptable in this project as the graphs used as undirected and unweighted.
+
+\begin{algorithm}
+	\caption{\texttt{Perfect $1$-Sparse Recovery}}
+	\SetKwInOut{Require}{require}
+	\Require{Update Stream $\{(i,\Delta)\}$}
+	$\phi,\iota,\tau\leftarrow0$\\
+	\For{\textbf{all} updates $(i,\Delta)$} {
+		$\phi\leftarrow\phi+\Delta$.\\
+		$\iota\leftarrow\iota+i\cdot\Delta$.\\
+		$\tau\leftarrow\tau+i^2\cdot\Delta$.\\
+	}
+	\If{$\iota^2\equiv\phi\cdot\tau$}{
+		\tcp{Vector is $1$-Sparse}
+		\Return{$\frac\iota\phi$}
+		\tcp*{Index of non-zero entry}
+	}\Else{
+		\tcp{Vector is \underline{not} $1$-Sparse}
+		\Return{\texttt{FAIL}}
+	}
+\end{algorithm}
+
+\textbf{Algorithm 5} uses three counters: sum of the weights $\phi=\sum\Delta$; sum of the weights weighted by their index $\iota=\sum i\cdot\Delta$; and, the sum of the weights weighted by the their index squared $\tau=\sum i^2\cdot\Delta$.
+\textbf{Lemma 3.1} proves that $\iota^2\equiv\phi\cdot\tau$ is a sufficient test for whether the described vector is exactly $1$-sparse.
+If the vector is exactly $1$-sparse then $\iota/\phi$ is the index of the non-zero element \& $\phi$ is its value.
+The only data stores are three counters each requiring $\mathcal{O}(\log n)$ space, thus \texttt{Perfect 1-Sparse Recovery} requires $\mathcal{O}(\log n)$ total space.
+
+\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+
+\textbf{Lemma 3.1} $\iota^2\equiv\phi\cdot\tau$ \underline{iff} a vector is exactly $1$-sparse.\\
+
+% TODO a proof by contradiction
+\textit{Proof} Consider a vector where $i$ is the only non-zero index and that index stores value $u$.\\
+Then $\phi=u,\ \iota=i\cdot u$ and $\tau=i^2\cdot u$. This means
+$$\iota^2=(iu)^2=i^2u^2=(i^2u)\cdot u=\phi\cdot\tau$$
+Thus, if a vector is exactly 1-sparse then $\iota^2\equiv\phi\cdot\tau$.\\
+Consider an exactly 2-sparse vector with non-zero indices $i,j$ each storing values $u,v$ respectively. Then $\phi=u+v,\ \iota=i\cdot u+j\cdot v$ and $\tau=i^2\cdot u+j^2\cdot v$. This means
+\[\begin{array}{rcl}
+\iota^2&=&(iu+jv)^2\\
+&=&i^2u^2+2ijuv+j^2v^2\\
+&\neq&i^2u^2+i^2ju+ij^2v+j^2v^2\\
+&=&(u+v)(i^2u+j^2v)=\phi\cdot\tau
+\end{array}\]
+Similar arguments show $\iota^2\neq\phi\cdot\tau$ for greater levels of sparsity.\\
+Thus, $\iota^2\equiv\phi\cdot\tau$ \underline{iff} a vector is exactly $1$-sparse.
+$\hfill\square$
+
+}}\end{adjustwidth}\vspace{.3cm} %TODO refine
+
+\subsection{Exact $s$-Sparse Recovery}
+
+\par \texttt{Exact $s$-Sparse Recovery} recovers a uniform sample of upto $s$ of the non-zero indicies in a vector described by a stream with probability $1-\gamma$.
+\textbf{Algorithm 6} defines a space efficient implementation of \texttt{Exact $s$-Sparse Recovery}.
+
+\begin{algorithm}
+	\caption{\texttt{Exact $s$-Sparse Recovery}}
+	\SetKwInOut{Require}{require}
+	\Require{Update Stream $\{(i,\Delta)\}$, Pairwise Independent Hash Functions $\mathcal{G}_2:[n]\to[2s]$, Failure Rate $\gamma$} % n=num vertices (== length of vector)
+	$r\leftarrow\log(s/\gamma)$\tcp*{Number of rows}
+	$c\leftarrow2s$\tcp*{Number of columns}
+	$M\leftarrow$$r\times c$ matrix of \texttt{Perfect $1$-Sparse Recoveries}\\
+	$\{g_1,\dots,g_r\}\subseteq\mathcal{G}_2$\tcp*{Hash function for each row}
+	\For{\textbf{all} updates $(i,\Delta)$} {
+		\For{$y\in[1,r]$} {
+			$v\leftarrow g_y(i)$\\
+			Pass update $(i,\Delta)$ to $M[y][v]$
+		}
+	}
+	$A\leftarrow\{\}$\tcp*{Set of recovered items}
+	\For{$x\in[1,c]$} {
+		\For{$y\in[1,r]$} {
+			$i\leftarrow$ recovered index of $M[x][y]$\\
+			\lIf{$i\neq\mathtt{FAIL}$} {$A\leftarrow A\cup\{i\}$}
+		}
+	}
+	\Return{$A$}
+\end{algorithm}
+
+%TODO probs reformat this whole thing (say it is a space-efficient technique with given failure rate then expand)
+%TODO is it a subvector or is it that the described vector is exactly s-sparse
+\par \textbf{Algorithm 6} uses a two-dimensional matrix $M$ of \texttt{Perfect $1$-Sparse Recoveries} and assigns a hash function $g_r$ to each row.
+When an update $(i,\Delta)$ is received, it is passed to the $g_r(i)^\text{th}$ \texttt{$1$-Sparse Recovery} of each row.
+Once the updates have been exhausted the returned elements of the $1$-sparse recoveries are collected into a set $A$ and returned.
+\par $g_r$ maps, on average, $\frac{n}c$ indices to each $1$-sparse recovery.
+These \texttt{$1$-Sparse Recoveries} fail iff more than one of these assigned indices is non-zero.
+It is proved in \cite{L0Framework} that setting the number of rows and columns to $\log(s/\gamma)$ and $2s$ respectively means \textbf{Algorithm 6} returns at most $s$ indices with probability $1-\gamma$.
+
+\par \textbf{Algorithm 6} uses $2s\cdot\log(s/\gamma)$ \texttt{$1$-Sparse Recoveries} thus the total space requirement for this implementation of \texttt{Exact $s$-Sparse Recovery} is $\mathcal{O}(s\log(s/\gamma)\log(n))$.
+
+\subsection{$L_0$ Sampling}
+
+\begin{algorithm}
+	\caption{\texttt{$L_0$ Sampler}}
+	\SetKwInOut{Require}{require}
+	\Require{Update Stream $\{(i,\Delta)\}$, $k$-wise independent hash function $h:[n]\to[n^3]$, $L_0$ Failure Rate $\delta$, $s$-Sparse Failure Rate $\gamma$} % n=num vertices (== length of vector), k in \mathcal{O}(s) (ie set it to s)
+	$j\leftarrow\log_2(n)$\tcp*{Number of \texttt{Exact $s$-Sparse Recoveries}}
+	$s\leftarrow\frac1\delta$\tcp*{Sparsity of \texttt{Exact $s$-Sparse Recoveries}}
+	Initialise $S_1,\dots,S_j$ as \texttt{Exact $s$-Sparse Recoveries}\\
+	$r\leftarrow0$\tcp*{Estimate of sparsity}
+	\For{\textbf{all} updates $(i,\Delta)$} {
+		$r\leftarrow r+\Delta$\\
+		\For{$k\in[1,j]$} {
+			$v\leftarrow h(y)$\\
+			\lIf{$v\leq\frac{n^3}{2^k}$} {Pass update $(i,\Delta_i)$ to $S_y$}
+		}
+	}
+	$x\leftarrow\log_2(r)$\tcp*{$s$-Sparse recovery to recovery \& sample from}
+	$a\leftarrow$returned from $S_x$\tcp*{Set of returned non-zero indices}
+	\If{$|a|\in[1,s]$}{
+		\Return{Index in $a$ with lowest hash value}
 	}
@@
 \end{algorithm}
-%TODO check space requirements
-\par\textbf{Algorithm 2} requires $O\left(n^2\right)$ space.
-Thus the space used to run the algorithm grows very quickly which could easily lead to the computer's RAM overflowing, increasing execution time.
-For solutions which are space inefficient run times become unmanagable for large graph streams.
-In 2014 in the UK, Facebook had 36.68 million users \cite{users16} each with an average of 155 friend connections \cite{friends16}.
-In order for this to be represented in a bi-partite graph for $\mathtt{Neighbourhood\ Detection}$ we would need a graph stream with ${\sim}5.6$ billion edges.
-\textbf{Algorithm 2} would not be suitable for analysing this graph.
-\par Instead we must look for much more space efficient algorithms for\\
-$\mathtt{Neighbourhood\ Detection}$.
-In the rest of this paper I shall discuss \& test an algorithm which can solve $\mathtt{Neighbourhood\ Detection}$ for insertion-only streams with space $O(n\log n+n^{\frac1c}d\log^2n)$; and, two approaches to solving for $\mathtt{Neighbourhood\ Detection}$ for insertion-deletion streams with space $\tilde O\left(\frac{xd}{c}\right)\ \&\ \tilde O\left(\frac{nd}{c}\left(\frac1x+\frac1x\right)\right)$ respectively.
 
-\section{Plan}
-In this paper I shall implement and test the two algorithms presented by Dr Chrisian Konrad in \cite{orig}.
-These algorithms solve the $\mathtt{Neighbourhood\ Detection\ Problem}$ for insertion-only \& insertion-deletion graph streams in a space-efficient manner.
-Both algorithms require only a single pass of the stream.
-\par I shall use the na\"ive algorithm described in \textbf{Algorithm 2} as a benchmark for both time \& space efficiency of these two algorithms.
-I shall use a four different graph streams, described in \textbf{Table 1.1}, for my tests.
-% probs should have more graphs
-These graphs were acquired from the \textit{Stanford Network Analysis Project}, \cite{SNAP}.
-\par The theoretical space requirements for these two algorithms both depend on the degree variable, $d$, and precision variable, $c$.
-I shall perform tests where I vary each of these in order to see what the space requirements are in practice, and then compare these results to the theory.
+The process for \texttt{$L_0$ Sampling} given in \textbf{Algorithm 7} is similar to the process for \texttt{Exact $s$-Sparse Recovery} in \textbf{Algorithm 6}.
+$j$ \texttt{Exact $s$-Sparse Recoveries} are instantiated and a $k$-wise independent hash function $h(\cdot)$ (with $k\in\mathcal{O}(s)$) is used to determine which recovery to pass an update to.
+When an update $(i,\Delta)$ is received we update an estimate of the sparsity $r$ of the vector being sampled and pass the update to all $S_k$ where $h(i)\leq\frac{n^3}{2^k}$.
+Once the update stream is consumed we consider the set returned from the $\log(r)^\text{th}$ recovery and return the index which produces the lowest hash-value.
+Since the hash-values are assigned uniformly at random, this is equivalent to uniformly sampling from the returned set.
 
-\begin{table}[h]
-\caption{Test Graph Streams}
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-Name&\# Edges&\# Vertices&Max Degree&File Size, KB\\\hline
-$\mathtt{facebook\_small}$&292&52&36&3\\
-$\mathtt{facebook}$&60,050&747&586&587\\
-$\mathtt{gplus}$&1,179,613&12,417&5,948&52,839\\
-$\mathtt{gplus\_large}$&30,238,035&120,100&104,947&1,328,820\\\hline
-\end{tabular}
-\label{Tab:Tcr}
-\end{table}
+\par \textbf{Algorithm 7} uses $\log(n)$ \texttt{Exact $s$-Sparse Recoveries} thus the total space requirement for this implementation of $L_0$ sampling is $\mathcal{O}(s\log(s/\gamma)\log^2(n))$.
+During implementation it is shown that $\gamma,\delta\approx0.1$ produce good results, in this case the algorithm requires $O(\log^2 n)$ space for reasonably $n$.
 
-\begin{thebibliography}{9}
+\section{Proposed Algorithms}
+
+%\begin{itemize}
+%	\item There are two algorithms.
+%	\item Adapted L0 pseudocode for each.
+%	\item Scenarios when each algorithm is most suitable.
+%	\item L0 samplers in parallel (lots of pre-allocated space).
+%\end{itemize}
+
+In \cite{orig} two single-pass algorithms are proposed for \texttt{neighbourhood detection} in an insertion-deletion stream, both of which use $L_0$ sampling to acquire edges from the described graph.
+
+\subsection{Vertex Sampling}
+
+\begin{algorithm}
+	\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Vertex Sampling)}}
+	\SetKwInOut{Require}{require}
+	\Require{Degree bound $d$, Approximation factor $c$, Insertion-Deletion Stream $S$, List of Vertices $A$}
+	Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$\\
+	Sample a uniform random subset $A'\subseteq A$ of size $10\ x\ln n$ of vertices.\\
+	\For{$a\in A'$} {
+		Run $10\frac{d}{c}\ln n$ $l_0$-samplers on the set of edges incident to $a$.
+	}
+	\Return{\small Any neighbourhood of size $\frac{d}{c}$ among the stored edges, if there is one.}
+\end{algorithm}
+%TODO requires list of vertices
+\par\textbf{Algorithm 8} takes a uniform sample from the vertex set and then runs multiple $L_0$ samplers on the edge vector of each sampled vertex.
+This is equivalent to sampling uniformly from the neighbourhood of each vertex multiple times.
+From these edges a neighbourhood can be found and if it is of sufficient degree then it may be returned by the algorithm.
+The vertex sample is taken uniformly as we have no prior information about the degree of the vertices in the graph and thus assume each is equally likely to be of sufficient degree.
+
+\par Assuming $\mathcal{O}(\log n)$ space is required to store a vertex, the vertex sample requires $\mathcal{O}(x\log^2(n))$ and a total of $100\frac{xd}{c}\ln^2(n)$ $L_0$ samplers are run each requiring $\mathcal{O}(\log^2(n))$ space.
+This means the total space requirement for \textbf{Algorithm 8} are $\mathcal{O}(\frac{xd}{c}\log^4(n))$.
+
+\par As incident edges are sampled uniformly, using $10\frac{d}c\ln n\gg\frac{d}c$ samplers results in a high probability that a neighbourhood size $\frac{d}c$ is found for vertices of degree greater $\frac{d}c$.
+Thus the success of \textbf{Algorithm 8} relies on it sampling a a vertex of sufficient degree.
+This means \textbf{Algorithm 8} is not effective when there are very few vertices of sufficient degree.
+This scenario occurs for small values of $c$ and for graphs with degree distributions similar to a star graph.
+In these cases sampling the edges directly has a greater success rate.
+
+\subsection{Edge Sampling}
+
+\begin{algorithm}
+	\caption{One-pass $c$-approximation Insertion-Deletion Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$. \text{(Edge Sampling)}}
+	\SetKwInOut{Require}{require}
+	\Require{Degree bound $d$, Approximation factor $c$, Insertion-Deletion Stream $S$}
+	Let $x=\max\left\{\frac{n}{c},\sqrt{n}\right\}$.\\
+	Run $10\frac{nd}{c}\left(\frac1x+\frac1c\right)\ln (nm)$ $l_0$-samplers on the set of edges incident to $a$.\\
+	\Return{\small Any neighbourhood of size $\frac{d}{c}$ among the returned edges, if there is one.}
+\end{algorithm}
+
+\par\textbf{Algorithm 9} runs $L_0$ samplers directly on the set of edges in order to acquire a uniform sample of the edges in the graph.
+This sample is then checked for any neighbourhoods of degree $\frac{d}c$.
+The sample of edges can be considered as an insertion-only graph stream, thus \textbf{Algorithm 2} can be used to find sufficient neighbourhoods or in cases where the number of sampled edges is very low the naïve \textbf{Algorithm 4} could be more appropriate due to its speed.
+\par The edge sampling approach is most likely to succeed when there is a small set of vertices which the majority of edges are incident to.
+The total space complexity of \textbf{Algorithm 9} is $\mathcal{O}\left(\frac{nd}c\left(\frac1x+\frac1c\right)\log^3(n)\right)$ which is less than the total space complexity of \textbf{Algorithm 8}.
+%TODO something about sparsity of the graph (max edges is O(n^2))
+
+\section{Implementation}
+
+\par During these implementations I made the assumption that vertices were labelled with unique values in $[1,n]$.
+This simplifies many parts of the implementation as each vertex aligns to an index in the edge vector.
+This assumption is reasonable as it places no restrictions on the use of these implementations as a \texttt{map} can be used to assign index locations to each vertex.
+
+\subsection{Perfect 1-Sparse Recovery}
+
+\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+
+\textbf{Lemma 3.2} For an unweighted insertion-deletion stream $\phi\equiv1$ and $\iota$ is the non-zero index \underline{iff} a vector is exactly $1$-sparse.\\
+
+% TODO a proof by contradiction
+\textit{Proof} In an unweighted insertion-deletion streams $\Delta=-1$ for a deletion instruction and $\Delta=1$ for an insertion instruction.
+In the case where a vector is exactly 0-sparse (\ie is the zero vector) the result trivially fails since there is no non-zero index for $\iota$ to hold.
+Consider a vector $v$ with non-zero index $i$.
+If $v$ is exactly 1-sparse then $i$ must have been insert exactly one more time than it was deleted, and all other indexes must have had the same number of insertions and deletions.
+Since each deletion cancels out an insertion there is only one insertion left.
+This means $\phi=1$ and $\iota=1\cdoti=i$, the result holds.\\
+Suppose $v$ is exactly $s$-sparse with $s>1$.
+Then there exists $s-1$ other indices $j_1,\dots,j_{s-1}\neq i$ which are inserted exactly one more time than it is deleted.
+After letting the deletions cancel the insertions there are $s$ insertion instructions left, this means $\phi=s$ and the result does not hold.
+Thus the result holds \underline{iff} $v$ is exactly $1$-sparse.$\hfill\square$
+
+}}\end{adjustwidth}\vspace{.3cm}
+
+The insertion-deletion streams used in this project are unweighted.
+Thus, by \textbf{Lemma 3.2}, the exact 1-sparse recovery process describded in \textbf{Algorithm 5} can be simplified to the implementation described in \textbf{Algorithm 10}.
+This implementation removes the $\tau$ counter meaning the total space used by the algorithm is reduced by a third which is significant as lots of these processes are run during \textbf{Algorithm 8} and \textbf{Algorithm 9}.
+Implementing \textbf{Algorithm 10} is trivial.
+
+\begin{algorithm}
+	\caption{\texttt{Case Specific Perfect $1$-Sparse Recovery}}
+	\SetKwInOut{Require}{require}
+	\Require{Update Stream $\{(i,\Delta)\}$}
+	$\phi,\iota\leftarrow0$\\
+	\For{\textbf{all} updates $(i,\Delta)$} {
+		$\phi\leftarrow\phi+\Delta$.\\
+		$\iota\leftarrow\iota+i\cdot\Delta$.
+	}
+	\lIf{$\phi\equiv1$}{\Return{$\iota$}
+	}\lElse{\Return{\texttt{FAIL}}}
+\end{algorithm}
+
+\subsection{Exact $s$-Sparse Recovery}
+
+\par When implementing \texttt{Exact $s$-Sparse recovery} the main decision is how to implement the family of pairwise hash functions efficiently. I used the set of functions
+$$\mathcal{H}_2:=\{h_{a,b}(x)=((ax+b)\text{mod }p)\text{mod }2s|a,b\in[1,P-1]\}$$
+where $P$ is a prime greater than $2s$.
+These hash functions map to the desired range $[0,2s)$.
+This family of hash functions are ideal for this project as can be stored in $O(\log P)$ bits as only the values of $a,b$ need to stored for each function.
+Similarly it is easy to generate a function in this family by sampling twice from $[1,P-1]$.
+
+\par Finding primes is hard so I hard coded $P=1,073,741,789>2^{29}$.
+This is an upper bound on number of columns the \texttt{$s$-Sparse Recovery} algorithm can handle, which is equivalent to placing a lower bound on the acceptable failure rate $\delta$ of the \texttt{$L_0$ Sampling} algorithm.
+$2\cdot\frac1\delta<P\implies\delta>\frac2P\approx2\times10^{-9}$.
+It is shown during evaluation that this bound is inconsequential to the success and performance of \texttt{$L_0$ Sampling}.
+
+\subsection{$L_0$ Sampling}
+%NOTE this was hard
+%\begin{itemize}
+%	\item L0 sampling a graph stream. (ie what is the vector in this case).
+%	\item Adapted L0 pseudocode for graph streams.
+%	\item Hash functions are hard (trade off due to dodgy hash generation).
+%	\item Tests for uniformity
+%\end{itemize}
+
+%NOTE maybe move to implementation
+For this project $L_0$ sampling is used on two different vectors: the edge vector of a specified vertex; and the edge set of the whole graph.
+The edge set of a graph is often defined as an $n\times n$ adjacency matrix, to apply $L_0$ sampling to this matrix we flatten it into a $n^2$ dimensional vector by reading the rows sequentially.
+The graphs used in this project are unweighted so the values of the edge vector are booleans defining whether an edge exists or not.
+This is implement by setting $\Delta=1$ for insertion instructions and $\Delta=-1$ for deletion instructions.
+This means the values in the vectors are 0 or 1 only.
+
+%TODO testing??
+\begin{figure}[H]
+	\label{Figure 13}
+	\subfloat[Pairwise Hash Function]{{\includegraphics[width=.5\textwidth]{img/pairwiseL0Test.png}}}
+	\subfloat[Unique Hash Map]{{\includegraphics[width=.5\textwidth]{img/uniqueL0Test.png}}}
+	\caption{Occurrences of each value in a sample when different hashing methods are used for $L_0$ sampling, order by increasing number of occurrences.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+%TODO map is now an array since vertices are zero-indexed and sequential (less space & quicker)
+When implementing \texttt{$L_0$ Sampling} the main problem is how to implement the $k$-wise independent hash function efficiently.
+I looked into libraries which provided such functions but none allowed for the range to be controlled and taking the modulus with respect to $n^3$ breaks the independence property.
+Using a pairwise independent hash function would be ideal as they are easy to implement.
+\texttt{Figure 3.3.1 a)} shows the results when testing using a pairwise hash function.
+These results fail a $\chi^2$ test for a uniform distribution, indicating a pairwise independent hash function is not sufficient.
+
+\par A \texttt{map} which assigns each value in $[n]$ to a distinct value in $[n^3]$, with values chosen uniformly at random, is equivalent to an $n$-wise independent hash function.
+The probability of a collision when assigning each value is less than $\frac1{n^2}$ meaning no collisions are expected when assigning values.
+This means that such a \texttt{map} is generated in $\mathcal{O}(n)$ time on average and requires $\mathcal{O}(n\log n)$ space.
+\texttt{Figure 3.3.1 b)} shows the results when such a \texttt{map} is implemented as the hash function $h$.
+These results pass a $\chi^2$ test for a uniform distribution so I implemented this \texttt{map} as the $k$-wise independent hash function.
+As the vertices are assumed to be labelled in $[n]$ this \texttt{map} can be implemented as an array with the hash values stored in each index.
+Using an array is more space-efficient than a tradition \texttt{map}.
+
+\par The maximum hash value is a constraint on the size of the domain and thus the number of vertices which can be handled.
+In theory this is not a problem, but in practice the limit of this value depends on the data type used.
+I implemented the hash array as an array of \texttt{uint64\_t} (8 bytes) which has the greatest maximum value of the standard c++ numeric data type.
+The maximum domain is the cube root of this value, thus \textbf{Algorithm 8} will not work for graphs with more than $(2^{64}-1)^{1/3}\approx2.6\times10^6$ vertices and \textbf{Algorithm 9} will not work for graphs with more than $(2^{64}-1)^{1/6}\approx1,625$ vertices.
+This constraints can be reduced by reducing the relative range of the hash function.
+
+\begin{figure}[H]
+	\label{Figure 14}
+	\subfloat[Varying $\delta$ ($\gamma=.1$)]{{\includegraphics[width=.5\textwidth]{img/varyDelta.png}}}
+	\subfloat[Varying $\gamma$ ($\delta=.2$)]{{\includegraphics[width=.5\textwidth]{img/varyGamma.png}}}
+	\caption{Space Usage and time taken when varying $\gamma$ and $\delta$ in $L_0$ Sampling implementation. For graph \texttt{facebook\_deletion}}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par When setting $\gamma$ and $\delta$ there is a trade off between space-usage and the success rate.
+The value of $\delta$ affects the number of columns and rows in each $s$-sparse recovery, whereas $\gamma$ only affects the number of rows.
+Thus optimising $\delta$ first has the greatest effect.
+\texttt{Figure 3.3.2 a)} plots success-rate and time taken against space used when $\gamma=.1$ and $\delta$ is varied.
+The time taken is constant throughout, but the success rate plateaus at $\smallsim\!90\%$ when $d\geq.2$.
+\texttt{Figure 3.3.2 b)} shows that fixing $\delta=.2$ and varying $\gamma$ has no effect on performance.
+Thus $\gamma$ is set to .3 in order to minimise space usage.
+
+\subsection{Vertex Sampling Algorithm} %TODO HERE
+%\begin{itemize}
+%	\item L0 samplers in parallel
+%	\begin{itemize}
+%		\item require lots of preallocated space (know how much space is used at start of algorithm)
+%	\end{itemize}
+%	\item Require list of vertices before vertex sampling.
+%\end{itemize}
+
+\par The vertex sample $A$ can be acquired by uniformly sampling from $[n]$ until a sufficient number of unique values is sampled.
+For a sample of size $m$, each time a vertex is sampled the probability of a collision is at most $\frac{m}n$ meaning the expected number of total collisions is at most $m$.
+Thus, in the worst case a sample is generated in $\mathcal{O}(mn)$ time.
+
+\par The sampled edges form an insertion-only graph stream so a sufficient neighbourhood can be found among them using the algorithms discussed in \textbf{Chapter 2}.
+Due to the limitations from the $L_0$ sampler implementation I could not run this implementation on particular large graphs, this meant that the number of sampled edges is very small and thus the naive \textbf{Algorithm 4} provided the best results as it always succeeds.
+If the implementation was improved to run on larger graphs then \textbf{Algorithm 3} should be used for its overall performance improvements.
+
+%\par Running $L_0$ samplers in parallel requires space being allocated for them before the stream is passed.
+This means all the space used by the algorithm (except that used to check for a sufficient neighbourhood) is known before the algorithm is run.
+This allows for a lot of evaluation of the space efficiency of the algorithm before implementation.
+%TODO consider removing
+
+\begin{figure}[H]
+	\label{Figure 14}
+	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/facebookDeletionInitialTime.png}}}
+	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/facebookDeletionInitialSpace.png}}}
+	\caption{Space Usage and time taken for initial implementation of vertex sampling, for graph \texttt{facebook\_deletion}, when varying approximation factor.}
+	%"insertion_deletion_facebook
+\end{figure}
+
+\texttt{Figure 3.3.3} shows the results of an initial implementation of vertex sampling.
+The proposed sample size would have been greater than $n$ for \texttt{facebook\_deletion} so $\sqrt{n}$ was used as a placeholder value until a more suitable sample size was tested for.
+This sample size proved too small for $c\leq4$ and resulted in some failures (Success rate remained above $83\%$ in all cases).
+These failures do not have a significant effect the performance of the algorithm as checking for sufficient neighbourhoods accounts for a small amount of the time and space usage.
+As expected, the hashing functions used in $L_0$ sampling account for the majority of space and time usage.
+Reducing the number of samplers will reduce the time and space usage of the implementation but will increase the failure rate if the number of unique edges sampled is reduced.
+The suggested number of samplers allows for a lot of redundancy as many edges are sampled multiple times.
+
+\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+
+\textbf{Lemma 3.3} $\frac{\ln(1-p)}{\ln\left(1-\frac1m\right)}$ $L_0$ samplers are expected to find $p\delta$ unique neighbours of a vertex of degree $\delta$.\\
+
+\textit{Proof} Consider a vertex $v$ of degree $\delta$ and let $i$ be a neighbour of $v$.
+Each $L_0$ sampler run on the edge vector of $v$ does not sample $i$ with probability $\frac{\delta-1}\delta$.
+As each sampler is independent the probability of $i$ not being sampled after $t$ $L_0$ samplers have been run is $\left(\frac{\delta-1}\delta\right)^t$.
+For the probability that $i$ has been sampled to be at least $p$ requires $\left(\frac{\delta-1}{\delta}\right)^t\leq 1-p$.
+This means $t\geq\frac{\ln(1-p)}{\ln\left(1-\frac1m\right)}$ samplers need to have been run.
+Since each neighbour is independent, $p$ is also the expected proportion of the neighbours which are sampled.
+
+}}\end{adjustwidth}\vspace{.3cm}
+
+\par \textbf{Lemma 3.3} tells us the minimum number of $L_0$ samplers required for us to expect a sufficient neighbourhood to a vertex to be found and by tuning the value of $p$ the number of samplers and achieved success rate can be optimised.
+When a vertex is sampled it is appropriate to assume it has at least degree $\frac{d}c$.
+The greater the actual degree of the vertex, the lower $p$ needs to be for $\frac{d}c$ neighbours to be returned as collisions are less likely.
+So tuning $p$ for vertices of degree close to $\frac{d}c$ is a good indicator of how vertices of greater degree will perform.
+Testing showed that setting $p=.9$ resulted in a sufficient neighbourhood being found $23\%$ of the time for a vertex of degree $\frac{11d}{10c}$ and $97\%$ of the time for a vertex of degree $\frac{12d}{10c}$.
+The implementation of $L_0$ samplers being used only successfully samples $90\%$ of the time, to account for this the expected number of samplers is multiplied by $\frac{1}{.9}$.
+
+\begin{figure}[H]
+	\label{Figure 15}
+	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/facebookDeletionSamplersPerVertexTime.png}}}
+	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/facebookDeletionSamplersPerVertexSpace.png}}}
+	\caption{Time taken and space used when the number of $L_0$ samplers is optimised against the initial implementation, for graph \texttt{facebook\_deletion}, when varying approximation factor.}
+	%"insertion_deletion_facebook"
+\end{figure}
+\texttt{Figure 3.3.4} shows the results when $\frac1{.9}\cdot\frac{\ln(1-.9)}{\ln(1-\frac{c}d)}$ samplers are run for each sampled vertex, rather than the suggested $10\frac{d}{c}\ln(n)$.
+The total space usage more than halved in all cases, and the space not used by the hash maps reduced by over $80\%$.
+The time taken by the algorithm (excluding that spent producing the hash maps) decreased by over $40\%$ in all cases.
+Both of these reductions are due to the total number of $L_0$ samplers being reduced.
+The success rate was unchanged, showing that there was a lot of redundancy in the number of samplers being run.
+
+\par The optimal vertex sample size depends on the degree distribution of the graph as this describes how many vertices are of sufficient degree.
+Thus any prior knowledge of the degree distribution of the graph should be used.
+It is trivial to establish the degree distribution of a graph from a graph stream by counting the degree of each vertex although an exact distribution is not required as research has been done into typically degree distributions for common graphs from real world applications.
+In this project I have been using graphs from social networks which are theorised \cite{socialNetworkDistribution} to follow a Power-law distribution $f(x)=cx^{-\alpha}$.
+By Zipf's Law of Power-law distributions, the expected number of vertices of degree at least $\frac{d}{c}$ is $c$.
+Thus a uniform sample of $\frac{n}c$ vertices in $V$ is expected to contain a vertex of sufficient degree.
+A larger sample should be taken as it is not guaranteed that the maximum neighbourhood will be found for all sufficient vertices.
+Using this as a basis for the vertex sample size should make the algorithm very successful but as $\frac{n}c\gg\sqrt{n}$ for the most interesting values of $c$ its performance will be significantly worse.
+
+\begin{figure}[H]
+	\label{Figure 17}
+	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/facebookDeletionVertexSampleSizeTime.png}}}
+	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/facebookDeletionVertexSampleSizeSpace.png}}}
+	\caption{Space Usage and time taken when $\frac{12n}{10c}$ vertices are sampled against implementation of reduced number of sampelrs, for graph \texttt{facebook\_deletion}, when varying approximation factor.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\texttt{Figure 3.3.5} shows the performance results when a vertex sample size of $\frac{12n}{10c}$ was taken compared to the initial $\sqrt{n}$.
+As predicted, the algorithm succeed in all cases where the new sample size was used however it's performance is significantly worse in all cases where $\frac{12n}{10c}\geq n$.
+The small size \texttt{facebook\_deletion} means that Zipf's Law is unlikely to fit well and it is theorised that the degree distribution of social networks have a fatter tail than a standard Power-law distribution.
+It is clear that further research is required into a performance optimal sample size for different scenarios.
+
+\subsection{Edge Sampling Algorithm}
+
+Running an $L_0$ sampler on the set of edges means the domain of the $k$-wise independent hash function is the edge set and thus we require a way of indexing the edges such that each has a unique value which can then be hashed.
+This indexing needs to be bijective so that the recovered index can be converted back into an edge.
+My initial implementation assigned the edge $(u,v)$ to $i=(\mathtt{min}(u,v)-1)\cdot n+\mathtt{max}(u,v)-1$ which maps edges to values in $[0,n^2)$.
+The use of \texttt{max} and min means that $(u,v)$ and $(v,u)$ map to the same value which is desirable since the edges are undirected.
+An edge can be recovered from this mapping as $\mathtt{min}(u,v)=(i\text{ mod }n)+1$ and $\mathtt{max}(u,v)=\mathtt{floor}(1/n)+1$.
+This mapping means the hash maps used in $L_0$ sampling now have $\mathcal{O}(n^2)$ space and time complexity.
+%TODO extracting edge from the key
+\par As with the vertex sampling approach finding a sufficient neighbourhood in the sampled vertices should be done using the algorithms discussed in \textbf{Chapter 2}.
+
+\begin{figure}[H]
+	\label{Figure 18}
+	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/facebookSmallDeletionEdgeSampleSizeTime.png}}}
+	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/facebookSmallDeletionEdgeSampleSizeSpace.png}}}
+	\caption{Space Usage and time taken for initial implementation of edge sampling, for graph \texttt{facebook\_small\_deletion}, when varying approximation factor.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\texttt{Figure 3.3.6} shows the results from the initial implementation of \textbf{Algorithm 9}.
+As with the implementation of the vertex sampling algorithm, the majority of the space and time requirements are spent in generating the hash arrays for the $L_0$ samplers.
+The requirement for generating these is so high that I could not run the algorithm on \texttt{facebook\_deletion} as the space requirements were over $1$ GB and it took several minutes to generate a single hash array.
+
+\begin{figure}[H]
+	\label{Figure 19}
+	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/facebookSmallDeletionReducedHashMapTime.png}}}
+	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/facebookSmallDeletionReducedHashMapSpace.png}}}
+	\caption{Space Usage and time taken when a reduced hash map is implemented against the initial implementation of edge sampling, for graph \texttt{facebook\_small\_deletion}, when varying approximation factor.}
+	%"Final Evaluation - insertion-only"/Master
+\end{figure}
+
+\par To reduce the size of the hash maps I changed how the edges were indexed.
+The graphs used in this project are undirected so have at most $\frac12n(n-1)$ unique edges.
+These can be visualised as the upper triangle of an adjacency matrix.
+The following formula map between an edge $(u,v)$ to key $k\in[\frac{1}{2}n(n-1)]$.
+In the implementation I ensured that $v>u$ so that the edges $(u,v)$ and $(v,u)$ mapped to the same $k$ and that the equations would succeed.
+% min & max
+\[\begin{array}{rcl}
+k&=&\frac12n(n-1)-\frac12(n-v)(n-v-1)+u+v-1\\
+v&=&n-2-\text{floor}\left(\frac{1}2\sqrt{-8k+4n(n-1)-7}-\frac{1}2\right)\\
+u&=&k+v+1-\frac12n(n-1)+\frac12(n-v)(n-v-1)
+\end{array}\]
+Implementing this method of indexing edges optimises the hash arrays but they still impact performance too much for the implementation to be usable on anything except the smallest of graphs.\texttt{Figure 3.3.7} shows the performance improvements from this implementation.
+
+\par This implementation succeeded in every case due to large number of samplers successfully sampling every edge, this is inefficient as most of these edges went unused while checking for a sufficient neighbourhood.
+For larger graphs not all edges would be sampled as the relative difference between the number of edges and number of samplers decreases.
+However, this implementation is not usable on reasonably sized graphs so I cannot effectively test alterations to the number of samplers run.
+The edge sampling method should be designed to work in cases where the vertex sampling method fails.
+Primarily this is when it is unlikely that a sufficient vertex is sampled, this occurs in graphs where only a handful of vertices are of high degree and the rest are of very low degree.
+In these cases there are very few edges which do not involve a vertex of high degree and by estimating how many high degree vertices there are you can calculate the expected number of edges which need to be sampled before a sufficient neighbourhood is found.
+
+\section{Evaluation}
+
+The performance of the implementations of \textbf{Algorithm 8} and \textbf{Algorithm 9} are dominated by the hash functions used by the $L_0$ samplers.
+This meant that the implementations could not be run on graphs of sufficient size to evaluate certain aspects of the implementations, namely the vertex sample size in \textbf{Algorithm 8} and number of samples in \textbf{Algorithm 9}.
+As well, it would be inappropriate to compare the performance to that of the naïve algorithm described in \textbf{Algorithm 11} as the proposed algorithms account for very little of the time and space of the whole implementation.
+
+\begin{algorithm}[H]
+\caption{Na\"ive Single-Pass Insertion-Streaming Algorithm for Neighbourhood Detection}
+\SetKwInOut{Require}{require}
+\Require{Stream $\{(w_0,s_0,t_0)\dots(w_n,s_n,t_n)\}$, degree bound $d$, precision bound $c$}
+$N\leftarrow\{\{\}\}\ \{\text{neighbourhoods}\}$\\
+\For{$i=0\dots n$} {
+	\lIf{$w_i\equiv1$}{append $t_i$ to $N[s_i]$\hfill \texttt{// Insertion instruction}}
+	\lElseIf{$w_i\equiv-1$}{remove $t_i$ from $N[s_i]$\hfill \texttt{// Deletion instruction}}
+}
+\tcp{Look for a sufficiently large neighbourhood}
+\For{$s_i\in N.keys$}{
+	\If{\text{size}($N[s_i])\geq d$} {
+		\tcp{Sufficiently large neighbourhood found}
+		$S\leftarrow \{N[s_i][0],\dots,N[s_i][\frac{d}{c}]\}$\tcp*{First $\frac{d}c$ elements of neighbourhood}
+		\Return{($s_i,S$)}
+	}
+}
+\Return{FAIL\hfill\texttt{No sufficient neighbourhoods}}
+\end{algorithm}
+
+\chapter{Conclusions}
+
+\section{Achievements}
+
+The algorithm proposed in \cite{orig} for insertion-only graphs proves very practical for solving \texttt{neighbourhood detection} on real world graphs.
+The changes to the implementations made reduced the time and space used by the algorithm significantly, while not compromising on the success rate.
+The final evaluation shows that the proposed algorithm is much more space-efficient than a naïve approach.
+Even for close approximations of large streams the final implementation uses less than $6\%$ of the file size and for $c\geq5$ it uses $\leq1\%$.
+This space efficiency increases the maximum file size at which the algorithm is usable, without requiring virtual memory.
+The run-times of my final implementation are very usable with solutions being found for a graph of over 30 million edges in a couple of minutes for close approximations $c\in[2,4]$.
+\par Due to difficulties in generating functions from a $k$-wise independent hashing family the proposed algorithm for insertion-deletion graphs could not be evaluated on graphs which were sufficiently large as to be interesting.
+However, I still managed to discuss several improvements to the implementation.
+The simplification of \texttt{1-sparse recovery} to only require two counters is a great improvement as, when the hash function is discounted, the \texttt{1-sparse recoveries} account for the majority of the space used by the $L_0$ samplers and thus both proposed algorithms.
+The probabilistic approach taken to reduce the number of $L_0$ samplers run for each vertex significantly improved performance by reducing redundancy, without compromising success.
+The indexing method which mapped edges to unique values in $[0,\frac12n(n-1))$ was a really good improvement as it is the optimal indexing method for undirected graphs.
+Although, the introduction of a mathematical hashing function rather than a look up table would negate this improvement.
+
+\section{Future Work}
+
+To conclude this project I will mention a few areas that I think any further evaluation of the algorithms proposed in \cite{orig} should consider.
+
+\subsection*{Complete Real World Evaluation}
+In this project I soley used graphs from social networks as they were readily available and an area where \texttt{neighbourhood detection} is an interesting problem.
+It would be interesting to test the implementations produced in this project on data sets from other fields.
+Section \texttt{I.ii Motivating Applications} provides a few examples.
+Further, it would be interesting to investigate how the returned neighbourhood is used to draw inferences about the graph as this would provide greater direction to how the implementations should be optimised.
+A possible problem could be detecting DDOS attacks from a network log.
+
+\subsection*{$k$-Wise Independent Hashing Family}
+An alternative family of $k$-wise independent hash function is ${\displaystyle h(x)=\left(\left(\sum_{i=0}^{k-1}a_ix^i\right)\text{mod }P\right)\text{mod }m}$ where $a_0>0$ and $a_i<P\ \forall\ i$.
+A function from this family is quick to generate as only the $k$ $a_i$ values need to be random generated and can be stored as $k+2$ integers (the values of $a_i$, $k$ and $m$).
+Both of these are dramatic improvements from the implementation of hash arrays used in the implementation of an $L_0$ sampler.
+The difficulty with implementation functions from this family is that the values of $x^i$ quickly overflow causing inaccurate calculations.
+I am sure an implementation of this family of hash functions could be worked out, but I ran out of time to do so in this project.
+
+\subsection*{Vertex Sample Size}
+I discussed how the size of the vertex sample used in \textbf{Algorithm 8} should depend on the degree distribution of the graph, and investigated one line of thought for how this could be done for graphs of social network connections.
+Investigating this further was outside the scope of this project but would be an interesting feature to research further.
+\par The edge sampling approach proposed in \textbf{Algorithm 9} should be considered whilst evaluating the size of the vertex sample, as there will be cases where this approach is more efficient.
+The most obvious example of this would be a star graph where the vertex sampling approach would need to sample every vertex in order to guarantee find a sufficient neighbourhood, whereas the edge sampling approach only needs to sample $\frac{d}c$ edges.
+In more general scenarios it is likely that the edge sampling approach is better for close approximations and the vertex sampling approach better for loser approximations.
+
+\subsection*{Reservoir Sample Size}
+I tested changing the reservoir sample size used in \textbf{Algorithm 2} but no strategies where found which consistently improved performance.
+These tests changed the sample size of every sampler used, but it may be that reducing the sample size for later samplers produces good results as those samplers are sampling for a smaller pool.
+This would be the main feature to look at when further trying to optimised the implementation from \textbf{Chapter 2}.
+
+\begin{thebibliography}{4}
+	\addcontentsline{toc}{chapter}{Bibliography}
 	\bibitem{orig} Christian Konrad \textit{Streaming Frequent Items with Timestamps and Detecting Large Neighborhoods in Graph Streams.} November 2019
-\bibitem{targetedAds} Howard Beales \textit{The Value of Behavioral Targeting}.
-\bibitem{users16} https://www.statista.com/statistics/553538/predicted-number-of-facebook-users-in-the-united-kingdom-uk/
-\bibitem{friends16} https://www.telegraph.co.uk/news/science/science-news/12108412/Facebook-users-have-155-friends-but-would-trust-just-four-in-a-crisis.html
 	\bibitem{SNAP}http://snap.stanford.edu/data/
+	\bibitem{L0Framework} Graham Cormode, Donatella Firmani \textit{A Unifying Frameworkd for $\ell_0$-Sampling Algorithms.} September 2014
+  \bibitem{socialNetworkDistribution} Tushti Rastogi \textit{A power law approach to estimating fake social network accounts.} May 2016
 \end{thebibliography}

### 2020-05-03 src/insertionDeletionStreams/l0Samplers/l0Sampler.cpp

@@ -30,6 +30,7 @@ int P=1073741789; // >2^30
  * DATA STRUCTURES *
  *-----------------*/
 using vertex = int; // typemap vertex
+using time_point=chrono::high_resolution_clock::time_point;
 
 struct edge { // undirected edge
     vertex fst;
@@ -54,7 +55,7 @@ vertex l0sampling(string file_path, vertex target, int num_vertices, double delt
 hash_params* choose_hash_functions(int num_cols, int num_rows);
 void update_s_sparse(vertex endpoint, int edge_value, int num_rows, hash_params* ps_s, long** phi_s, long** iota_s, long** tau_s);
 set<vertex> recover_neighbourhood(int num_cols, int num_rows, long** phi_s, long** iota_s, long** tau_s);
-vertex recover_vertex(set<vertex> neighbourhood, int sparsity, map<int,int> hash_map);
+vertex recover_vertex(set<vertex> neighbourhood, int sparsity, int* hash_map);
 
 // 1-sparse
 bool verify_1_sparse(int phi,int iota,int tau);
@@ -64,7 +65,7 @@ void update_1_sparse_counters(int index,int delta,int row,int col,long** phi_s,l
 hash_params generate_hash(int m);
 int hash_function(int key, hash_params ps);
 // n = num keys, m=possible hash values, hash=map from key to hash value
-void generate_random_hash(int n, int m, map<int,int>& hash);
+int* generate_random_hash(int n, int m);
 
 // Utility
 void parse_edge(string str, edge& e);
@@ -81,34 +82,39 @@ void write_to_file(string outfile_path, map<vertex,int>& edge_count);
 
 int main() {
   // details of graph to perform on
-  string file_path="../../../data/artifical/test_deletion_double.edges";
-  vertex target=779;
-  int num_vertices=1000;
+  string file_path="../../../data/facebook_deletion.edges";
+  vertex target=1;
+  int num_vertices=747;
 
   // constraints on s-sparse recovery
-  double delta=.01; // acceptable failure for L0
-  double gamma=.01; // acceptable failure for s-sparse recovery
+  double delta=0.25; // acceptable failure for L0
+  double gamma=.1; // acceptable failure for s-sparse recovery
 
   // check if edges sampled ~ uniformly
   map<vertex,int> edge_count;
 
   int succ_count=0;
   int num_runs=10000;
+  time_point before=chrono::high_resolution_clock::now(); // time before execution
[…]

### 2020-05-03 src/insertionDeletionStreams/l0Samplers/l0SamplerHashTo64bit.cpp

@@ -30,6 +30,7 @@ int P=1073741789; // >2^30
  * DATA STRUCTURES *
  *-----------------*/
 using vertex = int; // typemap vertex
+using time_point=chrono::high_resolution_clock::time_point;
 
 struct edge { // undirected edge
     vertex fst;
@@ -54,7 +55,7 @@ vertex l0sampling(string file_path, vertex target, int num_vertices, double delt
 hash_params* choose_hash_functions(int num_cols, int num_rows);
 void update_s_sparse(vertex endpoint, int edge_value, int num_rows, hash_params* ps_s, long** phi_s, long** iota_s, long** tau_s);
 set<vertex> recover_neighbourhood(int num_cols, int num_rows, long** phi_s, long** iota_s, long** tau_s);
-vertex recover_vertex(set<vertex> neighbourhood, int sparsity, map<int,int64_t> hash_map);
+vertex recover_vertex(set<vertex> neighbourhood, int sparsity, uint64_t* hash_map);
 
 // 1-sparse
 bool verify_1_sparse(int phi,int iota,int tau);
@@ -64,7 +65,7 @@ void update_1_sparse_counters(int index,int delta,int row,int col,long** phi_s,l
 hash_params generate_hash(int m);
 int hash_function(int key, hash_params ps);
 // n = num keys, m=possible hash values, hash=map from key to hash value
-void generate_random_hash(int n, int64_t m, map<int,int64_t>& hash);
+uint64_t* generate_random_hash(int n, uint64_t m);
 
 // Utility
 void parse_edge(string str, edge& e);
@@ -81,34 +82,43 @@ void write_to_file(string outfile_path, map<vertex,int>& edge_count);
 
 int main() {
   // details of graph to perform on
-  string file_path="../../../data/artifical/test_deletion_double.edges";
-  vertex target=779;
-  int num_vertices=1000;
+  //string file_path="../../../data/gplus_deletion.edges";
+  //vertex target=10;
+  //int num_vertices=12417;
+
+  string file_path="../../../data/facebook_deletion.edges";
+  vertex target=1;
+  int num_vertices=747;
 
   // constraints on s-sparse recovery
-  double delta=.01; // acceptable failure for L0
-  double gamma=.01; // acceptable failure for s-sparse recovery
+  double delta=.2; // acceptable failure for L0
[…]

### 2020-05-07 paper/src/paper.tex
Content done

@@
 % Title
-\title{Implementing \& Evaluating Space Efficient Algorithms for Detecting Large Neighbourhoods in Graph Streams}
+\title{Implementing and Evaluating Space Efficient Algorithms for Detecting Large Neighbourhoods in Graph Streams}
 \author{Dom Hutchinson}
@@
 \renewcommand\thesection{\thechapter.\roman{section}}
+\renewcommand\thesubsection{\thesection.\roman{subsection}}
 \setcounter{chapter}{1}
@@
 %\end{itemize}
-Graphs are used to represent relationships between objects and are popular in data science.
+
+%TODO why big data is important
+% The commercial movement towards data science and the desire to find relationships has lead to big data....
+\par The movement towards Big Data has lead to ever larger data sets which traditional offline algorithms cannot work with efficiently due to their high memory-requirements.
+As the size of data-sets is increasing at a much greater rate than standard RAM capacities means that virtual memory has to be used for these algorithms, significantly increasing their run-times.
+\par The \textit{Data Stream Model} was introduced as a method for coping with these large data sets.
+In the \textit{Data Stream Model} instructions are received sequentially, as a `stream', and generally random access is not allowed.
+\textit{Graph Streams} are a common implementation of this model, where instructions describe updates to the edge set of a graph and are typically ordered as a time-series.
+This allows for a dynamic representation of a graph, which typical \texttt{adjacency list} implementations do not.
+%TODO citation=https://en.wikipedia.org/wiki/Streaming_algorithm#Data_stream_model
+\par Streaming algorithms are a class of algorithms, first formalised in 1996 \cite{TODO}, to work within the \textit{Data Stream Model}.
+Streaming algorithms are typically designed to work in $polylog(n)$, where $n$ is the number of vertices in the case of graph streams.
+Streaming algorithms are designed to work in as few passes of the stream as possible, preferable only a single-pass is used as this means the stream can easily be extended without the algorithm having to start over.
+Due to the limited space streaming algorithms are allocated, they are often produce an approximation of a solution in order to improve space efficiency.
+When implementing a streaming algorithm the trade-off between accuracy and space-usage needs to be considered.
+
+\par Graphs are used to represent relationships between objects and are popular in data science.
 A graph stream is a sequence of instructions which describe how to construct a graph.
@@
 We can assume that $G$ contains at least one node of degree $d$.\\
-Here $d\in\nats$ is a threshold parameter \& $c\in(0,1]$ is an approximation parameter.
+Here $d\in\nats$ is a threshold parameter and $c>1$ is an approximation parameter.
 }}\end{adjustwidth}\vspace{.3cm}
@@
 \par\texttt{Neighbourhood Detection} can be solved trivially by storing the neighbourhood of every vertex and then returning any subset of size $\frac{d}c$.
-This trivial solution is very space inefficient and, with the movement towards Big Data, would quickly overflow a computer's RAM.
-This results in the use of virtual data which significantly slows down data access time and thus the whole algorithm.
+But, this solution is space inefficient, requiring $\mathcal{O}(n^2)$ space, and, with the movement towards Big Data, would quickly overflow a computer's RAM.
 This motivates the need for space efficient solutions to \texttt{Neighbourhood Detection}.
+
+\par In this paper I shall discuss two streaming algorithms which solving \texttt{Neighbourhood Detection} in $polylog(n)$ space.
+This is achieved by using a series of samplers in order to reduce the set of edges which need to be analysed.
+This means that there is a possibility of failure and this possibility increases as less edges are sampled, but there is a trade-off here with space efficiency.
+%TODO move this sentence & rewrite
+Using graph streams is good here as it allows for more space efficient algorithms but due to time-ordered nature of most graph streams their is an added difficulty that instructions do not arrive in the most optimal order for solving \texttt{Neighbourhood Detection}.
 
@@
 \begin{itemize}
-	\item[-] Given a list of connections in a social network, identify \textit{social influencers} and determine the demographics of their audience in order to decide who should promote a particular product.
-	\item[-] Given a list of receipts, identify which items are commonly sold together and use this information to determine how to stack shelves in order to increase sales.
-	\item[-] Given a log of traffic within a network identify which resources are being accessed most often, and by whom.
+	\item Given a list of connections in a social network, identify \textit{social influencers} and determine the demographics of their audience in order to decide who should promote a particular product.
+	\item Given a list of receipts, identify which items are commonly sold together and use this information to determine how to stack shelves in order to increase sales.
+	\item Given a log of traffic within a network identify which resources are being accessed most often, and by whom.
 This information can be used to determine what upgrades should be made to the network and to identify potential attacks on the network.
@@
 \section{Objectives}
+%TODO expand
 The objectives of this project are to:
 \begin{itemize}
+  \item Understand the problem posed in \cite{orig} and its context in practical situtations.
 	\item Implement and understand the algorithms proposed in \cite{orig}.
+  \item Implement na\"ive algorithms and test their performance against the proposed.
+  \item Show that na\"ive algorithms have a limit to the size of graph they can work on and that it is much lower than for the proposed algorithms.
+  \item Investigate the practicalities of the samplers used in these algorithms.
   \item Discuss and test alterations to the logic and implementation of the proposed algorithms, and the subroutines they rely upon, in order to improve the performance of the proposed algorithm.
+  \item Discuss the trade-offs in practicalities these alterations have, namely between speed and space usage.
 	\item Use large graphs from real-world scenarios to evaluate the practicalities of these implementations on real-world data sets.
+  \item Discuss how graphs from different real-world sources (\ie not social networks) might perform on these implementation and how the implementations could be changed to accomodate for them.
 \end{itemize}
+
+\section{Related Works}
+
+%\begin{itemize}
+%  \item Insertion-Only
+%  \begin{itemize}
+%    \item Counting triangles (ie subgraphs to assess clustering coefficient)
+%    \item Connected Graphs
+%    \item Random Walk
+%  \end{itemize}
+%  \item Insertion-Deletion
+%  \begin{itemize}
+%    \item Linear Sketches
+%    \item Sparsification via sampling
+%    \item $k$-Connecitivty
+%  \end{itemize}
+%\end{itemize}
+
+%TODO provide examples for how they could be used to analyse a graph of social-media connections in order to solidify their relevance
+
+Since the formalisation of the data streaming model in 1996 much research has been done into solving problems using this model. Here I give an overview of previous works in this field.\newline
+
+\par\noindent\textbf{Clustering coefficient}
+\par The clustering coefficient of a graph measures how likely it is that two connected nodes are part of a larger cluster.
+The equation for this measure is $C=3\times\frac{\text{number of triangles}}{\text{number of wedges}}$ where a triangle any sub-graph which contains three vertices and three edges (\ie any 3-cycles $C_3$ or 3-complete $K_3$ sub-graphs) and a wedge is any sub-graph which contains three vertices and only two edges (\ie any paths of length 2 which are not cycles).
+\par Jha et al.
+\cite{clusteringCofficient} present a space-efficient single-pass streaming algorithm which approximates the clustering coefficent (and triangle count) in $\mathcal{O}(m/T)$ space where $m$ is the number of edges and $T$ is the number of triangles.
+Their approach works for undirected insertion-only graph streams and produced relative errors of less than $5\%$.\newline
+
+\par\noindent\textbf{Minimum Distance between two nodes}
+\par A $\alpha$-\textit{spanner} subgraph $H$ of a graph $G=(V,E)$ is a subgraph where $\forall\ u,v\in V$ $d_G(u,v)\leq d_H(u,v)\leq\alpha\cdot d_G(u,v)$ where $d_G(\cdot,\cdot)$ is the length of the shortest path between the given nodes in $G$.
+Spanners are a space-efficient structure for evaluating the length of shortest path between any two nodes in a graph.
+A $(2t-1)$-spanner requires at most $\mathcal{O}(n^{1+\frac1t})$ edges need to be stored \cite{edgesOfSpanner}.
+\par Baswana \cite{spanner} presents an single-pass streaming algorithm for unweighted insertion-only graph streams which runs in $\mathcal{O}(m)$ time.
+This is an amortised run-time of $\mathcal{O}(1)$ per edge, a natural limit for performance.\newline
+
+\par\noindent\textbf{Random walks}
+\par A random walk from $v_0\in V$ in a graph is any random sequence of vertices $v_0,v_1,v_2,\dots$ where $v_i$ is randomly selected from the neighbourhood of $v_{i-1}$.
+The distribution of $v_t$ can be assessed.
+\par Das Sarma et al.
+\cite{RandomWalks} present an algorithm for approximating the distribution of $v_t$ in the context of estmating the \textit{PageRank} of a vertex.
+They present an algorithm which can perform $\frac{n}{l}$ indendepent random walks of length $l$ in $o(n)$ space and $o(l)$ passes.
+They use this algorithm to sample random walks of different lengths in order to estimate the distribution of $v_t$ and thus estimate \textit{PageRank}.\newline
+
+\par\noindent\textbf{Sketchs}
+\par A sketch of a graph is random linear projection of it.
+Streaming algoritms which run on sketchs are a popular area of reasearch at the moment as sketchs are of lower dimension than the true graph, reducing the space requirements.
+In his survey of graph stream algorithms \cite{GraphStreamSurvey} McGregor presents two types of sketch (linear and homomorphic) and discusses how they can be used to assess the connectiveness of a graph.
+The types of sketchs discussed in the survey are very versitile as they can be generated for both insertion-only and insertion-deletion graphs easily.
 
@@
 \renewcommand\thesection{\thechapter.\arabic{section}}
+\renewcommand\thesubsection{\thesection.\arabic{subsection}}
 \setcounter{chapter}{0}
@@
 These are all interconnected and come with their own trade-offs.
-Typically, the changes to space-used \& time-taken when different strategies are implemented occur in the same direction, but at different rates.
+Typically, the changes to space-used and time-taken when different strategies are implemented occur in the same direction, but at different rates.
 During most tests the success-rate was near perfect, but the other metrics were significantly higher as it took longer for a solution to be found.
@@
 \texttt{Degree-Based Reservoir Sampling} is a technique for uniformly sampling a vertex from the set of vertices whose degrees are greater than a specified minimum bound $d_1$ along with $d_2$ of its neighbours.
-This is achieved by maintaining a sample of this set of vertices, known as the \textit{reservoir}, and storing the first $d_2$ of the edges incident to a vertex after it is sampled.
-The size of the reservoir is a parameter of the algorithm.
+The sampler maintains a subset of the vertices of degree $d_1$, known as the \textit{reservoir} \cite{degResSampling}, and stores the first $d_2$ of the edges incident to a vertex after it is added to the reservoir.
+The size of this reservoir is a parameter of the algorithm.
 \textbf{Algorithm 1} outlines pseudocode for performing \texttt{Degree-Based Reservoir Sampling} on a bipartite graph using an insertion-only graph stream.
 
-\begin{algorithm}
+\begin{algorithm}[H]
 	\caption{\texttt{Degree-Based Reservoir Sampling}$(d_1,d_2,s)$}
@@
 	\If{$\exists\ a\in R$ with $D[a]\geq d_1+d_2-1$} {
-		\Return{Uniform random vertex \& neighbourhood from those of size $d_2$}
+		\Return{Uniform random vertex and neighbourhood from those of size $d_2$}
 	} \lElse { \Return{\texttt{FAIL}}}
@@
 
-\par Three data stores are used in \texttt{Degree-Based Reservoir Sampling}: a map of the degree of every node in the graph $D$; a set for the reservoir $R$; and a set for edges incident to the vertices in the reservoir $E$.
-
-\par The reservoir has an invariant that at any point in time it contains a uniform sample of the vertices whose degrees are known to be greater than $d_1$.
+\par Three data stores are used in \texttt{Degree-Based Reservoir Sampling}: a \texttt{map} of the degree of every node in the graph $D$; a \texttt{set} for the reservoir $R$; and a \texttt{set} for edges incident to the vertices in the reservoir $E$.
+
+\par The reservoir $R$ has an invariant that at any point in time it contains a uniform sample of the vertices whose degrees are known to be greater than $d_1$.
 This invariant is proved as \textbf{Lemma 2.1}, first I shall describe how the reservoir is maintained by controlling how vertices are inserted.
@@
 	\item If the reservoir is \underline{not} full (\ie $|R|<s$) then $v$ is inserted into the reservoir.
-	\item Otherwise, if the reservoir is already full, a Bernoulli random variable with probability $p=\frac{x}s$ is used, where $x$ is the number of vertices known to have degree at least $d_1$.
+	\item Otherwise, a Bernoulli random variable with probability $p=\frac{x}s$ is used, where $x$ is the number of vertices known to have degree at least $d_1$.
 If this random variable succeeds then pick, uniformly at random, a vertex $u$ currently in the reservoir and replace it with $v$.
@@
 
-\textbf{Lemma 2.1} At any time $t$, $R$ contains a uniform sample of the vertices whose degrees are known to be at least $d_1$.\\
+\textbf{Lemma 2.1} At time $t$, $R$ contains a uniform sample of the vertices whose degrees are known to be at least $d_1$.\\
 
@@
 Note that in the case that the reservoir is full and $v$ is inserted into it, $v$ is replacing an element that was in the reservoir meaning any progress made towards finding a neighbourhood for that vertex is annulled.
-The amount of progress is not taken into account \& thus if there is a high turn-over of elements in a reservoir then less progress is made towards finding a neighbourhood for each of them.
+The amount of progress is not taken into account and thus if there is a high turn-over of elements in a reservoir then less progress is made towards finding a neighbourhood for each of them.
 This is discussed further during evaluation.
@@
 \par \texttt{Degree-Based Reservoir Sampling} only works on insertion-only streams as it does not account for when the degree of a vertex falls below the threshold, after being sampled.
-This means the reservoir is not necessarily a uniform random sample of the vertices with degree $\geq d_1$.
-%TODO is this sufficient
+This means the reservoir may contain vertices whose degree is less than $d_1$.
 
@@
 
-The \texttt{Neighbourhood-Detection Problem} can be solved using \texttt{degree-based reservoir sampling} sampling by setting $d_2=\frac{d}c$.
+\texttt{Neighbourhood Detection} can be solved using \texttt{degree-based reservoir sampling} by setting $d_2=\frac{d}c$.
 The question remains as to what to set $d_1$ and $s$ to in order to achieve a high probability of success.
-%in little time and space
-
-\par \textbf{Algorithm 2} is the algorithm proposed in \cite{orig} for solving the \texttt{Neighbourhood-Detection Problem} for insertion-only streams.
+\textbf{Algorithm 2} is the space-efficient algorithm proposed in \cite{orig} for solving \texttt{Neighbourhood-Detection} for insertion-only streams.
 
@@
 
-\par In \cite{orig} it is proven that setting $s=\lceil\ln(n)\cdot n^{\frac1c}\rceil$ and using $c$ \texttt{degree-based reservoir samplers} each with a different lower bound, incremented from $\frac{d}c$ to $d$ stepping by $\frac{d}c$ each time, results in a high probability of success.
+\par In \cite{orig} it is proven that setting $s=\left\lceil\ln(n)\cdot n^{\frac1c}\right\rceil$ and using $c$ \texttt{degree-based reservoir samplers} each with a different lower bound, incremented from $\frac{d}c$ to $d$ stepping by $\frac{d}c$ each time, results in a high probability of success.
 The samplers are run in parallel so that the algorithm only requires a single pass of the stream.
@@
 \par Each of the $c$ samplers requires $\mathcal{O}(n\log n+sd_2\log n)$ space.
-Since the degree map can be shared between samplers the total space is $\mathcal{O}(n\log n+c\cdot sd_2\log n)=\mathcal{O}(n\log n+c\lceil\log(n)n^{\frac1c}\rceil\frac{d}c\log n)=\mathcal{O}(n\log n+n^{\frac1c}d\log^2n)$.
+Since the degree map can be shared between samplers the total space is $\mathcal{O}(n\log n+c\cdot sd_2\log n)=\mathcal{O}(n\log n+n^{\frac1c}d\log^2n)$.
 The proposed algorithm does require you to know the number of vertices in the graph stream before running the algorithm, or at least the $\log$ of it.
-If this is unknown then it is quick to run through the whole stream \& build a set of vertices in the graph.
+If this is unknown then it is quick to run through the whole stream and build a set of vertices in the graph.
 This requires $\mathcal{O}(n\log n)$ space, assuming $O(\log n)$ space is required to store a vertex.
@@
 \section{Implementation}
-%TODO justify changes wrt gplus graph only, include plots as I go
-\par I centred my implementation around working for the graph streams from \textit{SNAP} \cite{SNAP} as they came from real-world sources \& thus are very important for testing the real-world practicalities of the proposed algorithm.
-These graph streams are not formatted in a strictly bipartite way.
-I adjusted \texttt{degree-based reservoir sampling} to account for this by by repeating \texttt{lines 6-17} (and the early termination clause) of \textbf{Algorithm 1} after the end of the \texttt{for} loop, with all occurrences of $a_i$ replaced by $b_i$.
-%TODO something about implementing in parallel. Nah
-\par I tested changes to the implementation on multiple graphs, for brevity and succinctness I shall only show the results for \texttt{gplus}.
-\texttt{Figure 2.3.1} shows the results of testing the initial implementation of \textbf{Algorithm 2}.
-These results shows that the space requirements of \textbf{Algorithm 2} are significantly less than the size of file stream and that the space requirements decrease as the approximation factor increases but there is an undesirable result that the time taken increases as the approximation factor increases.
+
+\par The graphs from \cite{SNAP} are not formatted in a strictly bipartite way.
+I adjusted \texttt{degree-based reservoir sampling} to account for this by repeating \texttt{lines 6-17} (and the early termination clause) of \textbf{Algorithm 1} after the end of the \texttt{for} loop, with all occurrences of $a_i$ replaced by $b_i$.
+
 \begin{figure}[H]
@@
 
-\noindent After this initial implementation I considered two alterations which could be made:
+\par During this section I tested changes to the implementation on multiple graphs, for brevity and succinctness I shall only show the results for \texttt{gplus}.
+\texttt{Figure 2.3.1} shows the results of testing the initial implementation of \textbf{Algorithm 2}.
+These results shows that the space requirements of \textbf{Algorithm 2} are significantly less than the size of file stream and that the space requirements decrease as the approximation factor increases but there is an undesirable result that the time taken increases as the approximation factor increases.
+
+\par After this initial implementation I considered two alterations to the implementation:
 \begin{enumerate}
@@
 \end{enumerate}
+
+\subsection{Early Termination}
 
@@
 The space-requirements are down in every case and are down more than $90\%$ for reasonable values of $c$.
-Both metrics now converge quickly, to $\smallsim\!\!1\frac1.5$ seconds and $3.3$MBs respectively, meaning we can make predictions on the performance of the algorithm for greater values of $c$.
+Both metrics now converge quickly, to $\smallsim\!\!1.5$ seconds and $3.3$MBs respectively, meaning we can make predictions on the performance of the algorithm for greater values of $c$.
 These are really good results, so I implemented this change going forward.
@@
 
+\subsection{Shared Edge-Set}
+
 \vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
@@
 \textit{Proof (by contradiction)} Let $v$ be a vertex in the graph stream.
-Suppose, without loss of generality, that $v$ is a member of the reservoirs of both \texttt{Deg-Res-Sampling}$(\frac{d}{c},\frac{d}{c},s)$ \& \texttt{Deg-Res-Sampling}$(\frac{2d}{c},\frac{d}{c},s)$.
+Suppose, without loss of generality, that $v$ is a member of the reservoirs of both \texttt{Deg-Res-Sampling}$(\frac{d}{c},\frac{d}{c},s)$ and \texttt{Deg-Res-Sampling}$(\frac{2d}{c},\frac{d}{c},s)$.
 This means that $v$ was sampled twice, after encountering the $\frac{d}c^\text{th}$ \underline{and} $\frac{2d}c^\text{th}$ edge incident to $v$ in the stream.
@@
 This means that introducing a shared edge-set for all samplers will have little effect on the space requirements of \texttt{Algorithm 2}.
-\par However, the alterations to the implementation which would have to made to allow for a shared edge-set could introduce time savings.
-Namely, I would need a data-structure which tells me \underline{precisely} which reservoir each vertex is in that so I know the appropriate value of $d_1$ in \texttt{Algorithm 1:Line 18}, otherwise improper termination could occur.
+\par However, the alterations to the implementation, which would have to made to allow for a shared edge-set, could introduce time savings.
+Namely, a data-structure is required to store \underline{precisely} which reservoir each vertex is in so that the appropriate value of $d_1$ is used in \texttt{Algorithm 1:Line 18}, otherwise improper termination could occur.
 An appropriate data-structure for this would be a \texttt{map} from each vertex to a bitstring whose indices indicate whether the vertex lies in an associated reservoir.
-With this implementation, inserting or deleting a vertex to/from a reservoir and querying whether a vertex is in a given reservoir becomes querying a map \& then a known bit in a bitstring which both take $\mathcal{O}(1)$ time.
-This \texttt{map} replaces the \texttt{ordered set}s currently used to store the reservoirs of each sampler and thus reduce these action times from $O(\log n)$ to $\mathcal{O}(1)$ time.
+With this implementation, inserting or deleting a vertex to/from a reservoir and querying whether a vertex is in a given reservoir is achieved with a single query of a \texttt{map} and then a known bit in a bitstring which both take $\mathcal{O}(1)$ time.
+This \texttt{map} replaces the \texttt{ordered set}s currently used to store the reservoirs of each sampler and thus reduces these action times from $O(\log n)$ to $\mathcal{O}(1)$ time.
 As these actions occur a lot this change should result in a reasonable reduction in run-time.
@@
 Implementing a shared-edge set does not change the logic of the algorithm so does not affect its success rate.
-Due to the reduction in mean run-time \& no noteworthy changes in space requirements, I recommend implementing this change.
+Due to the reduction in mean run-time and no noteworthy changes in space requirements, I recommend implementing this change.
 
@@
 
-\par The changes discussed above have been to the technical implementation of the proposed algorithm, rather than to the its logic.
+\par The changes discussed in \textbf{Section 2.4} have been to the technical implementation of \textbf{Algorithm 2}, rather than to the its logic.
 This means these changes have had no affect on the success rate of the algorithm.
 I will now discuss potential changes to the logic of the proposed algorithm.
-These changes will affect on the success rate of the algorithm.
-It is possible that no change in the success-rate will be seen \& rather the change will occur in execution time as more instructions have to be read before a solution is found.
-I am investigating the practicalities of the algorithm so keeping the run-times usable is a high priority.
-In \texttt{Algorithm 2} there are two parameters which can be played around with: the number of samplers; and, the size of the reservoirs, $s$.
-
-\par The total reservoir size, $\left\lceil\ln(n)n^{\frac1c}\right\rceil\times\text{\#\ Samplers}$, is the main factor effecting the space-used by the algorithm as this is the number of vertices it needs to store edges for.
-Noting that \texttt{Algorithm 2} sets the number of samplers to be $c$ and that $cp\left\lceil\ln(n)n^{\frac1c}\right\rceil\leq c\left\lceil p\ln(n)n^{\frac1c}\right\rceil\ \forall\ p\in[0,1]$ it is apparent that reducing the number of samplers has a greater effect on total reservoir size, and thus space-used, than reducing reservoir size by the same proportion.
+These changes will affect the success rate of the algorithm.
+As I am investigating the practicalities of the algorithm keeping the run-times usable is a high priority.
+In \textbf{Algorithm 2} there are two parameters to scrutinise: the number of samplers; and, the size of the reservoirs $s$.
+
+\par The total reservoir size, $\left\lceil\ln(n)n^{\frac1c}\right\rceil\times\text{\#\ Samplers}$, is the main factor affecting the space-used by the algorithm as this is the number of vertices it needs to store edges for.
+Noting that \textbf{Algorithm 2} sets the number of samplers to be $c$ and that $cp\left\lceil\ln(n)n^{\frac1c}\right\rceil\leq c\left\lceil p\ln(n)n^{\frac1c}\right\rceil\ \forall\ p\in[0,1]$ it is apparent that reducing the number of samplers has a greater effect on total reservoir size, and thus space-used, than reducing reservoir size by the same proportion.
 For this reason I investigated the number of samplers first.
@@
 
-\par When changing the number of samplers we need to consider their lower bounds, $d_1$, too.
-If the range of vertex degrees which individual samplers sample, $[d_1,d_1+d_2]$, at overlap then duplicate sampling of vertices can occur which offers no advantages \& leads to a reservoir space being wasted.
+\par When changing the number of samplers we need to consider their lower bounds $d_1$ too.
+If the range of vertex degrees $[d_1,d_1+d_2]$ which individual samplers sample at overlap then duplicate sampling of vertices can occur which offers no advantages and leads to a reservoir space being wasted.
 And, if there are gaps between these ranges then there are vertices which may not be sampled.
@@
 Thus, the best set up is to have the sampling ranges be sequential with no gaps and so when varying the number of samplers we should only add or remove from the endpoints of these ranges.
-The samplers with lower lower-bounds, $d_1$, start sampling earlier \& from a larger pool.
+
+\par The samplers with lower lower-bounds $d_1$ start sampling earlier and from a larger pool.
 I expect these early samplers to be more important wrt performance gains, than later samplers.
 It is not necessarily true that the first sampler is the most important as it is sampling from the largest pool and so has the greatest turn-over rate in its reservoir.
-
-\par I ran the following tests to investigate what had a greater effect on performance, removing later samplers or earlier samplers.
-\begin{enumerate}
-	\item Removing the \underline{first} $x$ samplers from \texttt{Algorithm 2} for $x\in[1,c)$, for $c\in[1,20]$.\\
-	\textit{i.e.} Change the \texttt{for} loop in \texttt{Algorithm 2:Line 2} to run for $i\in[x,c-1]$.
-	\item Removing the \underline{last} $x$ samplers from \texttt{Algorithm 2} for $x\in[1,c)$, for $c\in[1,20]$.\\
-	\textit{i.e.} Change the \texttt{for} loop in \texttt{Algorithm 2:Line 2} to run for $i\in[0,c-x]$.
-\end{enumerate}
 
@@
 
-\par \texttt{Figure 2.4.4} shows the results of tests, with the results normalised to make them comparable.
+\par Two tests where run to investigate what had a greater effect on performance, removing later samplers or earlier samplers.
+\begin{enumerate}
+	\item Removing the \underline{first} $x$ samplers from \texttt{Algorithm 2} for $x\in[1,c)$, for $c\in[1,20]$.\\
+	\textit{i.e.} Change the \texttt{for} loop in \texttt{Algorithm 2:Line 2} to run for $i\in[x,c-1]$.
+	\item Removing the \underline{last} $x$ samplers from \texttt{Algorithm 2} for $x\in[1,c)$, for $c\in[1,20]$.\\
+	\textit{i.e.} Change the \texttt{for} loop in \texttt{Algorithm 2:Line 2} to run for $i\in[0,c-x]$.
+\end{enumerate}
+
+\par \texttt{Figure 2.4.4} shows the results of these tests, with the results normalised to make them comparable.
 These results confirm the hypothesis that the lower a sampler's lower-bound is, the greater the effect of its inclusion on performance.
@@
 Comparing the normalised results shows that removing early samplers is more detrimental to performance than including late samplers.
-This is due to late samplers only starting to sample later \& they may not have begun sampling by the time one of the previous samplers has succeed, terminating the algorithm.
+This is due to late samplers only starting to sample later and they may not have begun sampling by the time one of the previous samplers has succeed, terminating the algorithm.
 When the same proportion of samplers were removed for different values of the  $c$, the relative performance was worse for greater values of $c$.
@@
 	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/gplusRemovedSamplersSpace.png}}}
-	\caption{Results when using a reduced number of samplers and the technical improvements, tested on \texttt{gplus} graph for different approximation factors, against just using the technical improvements.}
+	\caption{Results when using a reduced number of samplers for different approximation factors, against just using the technical improvements. Results from graph \texttt{gplus}}
 	%"Final Evaluation - insertion-only"/Master
@@
 	}
-	\Return{\small Uniform random neighbourhood $(a_i,S_i)$ \text{from successful runs}}
+	\Return{\small First neighbourhood of sufficient degree and its root}
 \end{algorithm}
@@
 % Identify which change had what effect (to space, time, success)
-%TODO 1000star is useful as there is only one valid solution
 
@@
 	\item Initial Implementation - No early termination, shared edge-set or sampler removal. %insertionStreamsNotQuitEarly
-	\item After Technical Optimisation - Implements early termination \& a shared edge-set. %insertionStreamsSharedEdgeSet
+	\item After Technical Optimisation - Implements early termination and a shared edge-set. %insertionStreamsSharedEdgeSet
 	\item After Algorithmic Optimisation - Implements early termination, shared edge-set and sampler removal. %insertionStreamsRemovedSamplers
@@
 
-% TODO a table of results??
 \begin{figure}[H]
 	\label{Figure 12}
-	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusTime.png}}}
-	\subfloat[Space Used]{{\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusSpace.png}}}
-	\caption{Time taken and space used by different implementations of the proposed algorithm for insertion-only streams when tested on \texttt{gplus}}
+	\subfloat[\texttt{gplus} - Time Taken]{{\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusTime.png}}}
+	\subfloat[\texttt{gplus} - Space Used]{{\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusSpace.png}}}\newline
+  \subfloat[\texttt{gplus\_large} - Time Taken]{{\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusLargeTime.png}}}
+	\subfloat[\texttt{gplus\_large} - Space Used]{{\includegraphics[width=.5\textwidth]{img/insertionOnlyGplusLargeSpace.png}}}
+	\caption{Time taken and space used by different implementations of the proposed algorithm for insertion-only streams.}
 	%"Final Evaluation - insertion-only"/Master
@@
 \textbf{Lemma 3.1} proves that $\iota^2\equiv\phi\cdot\tau$ is a sufficient test for whether the described vector is exactly $1$-sparse.
-If the vector is exactly $1$-sparse then $\iota/\phi$ is the index of the non-zero element \& $\phi$ is its value.
+If the vector is exactly $1$-sparse then $\iota/\phi$ is the index of the non-zero element and $\phi$ is its value.
 The only data stores are three counters each requiring $\mathcal{O}(\log n)$ space, thus \texttt{Perfect 1-Sparse Recovery} requires $\mathcal{O}(\log n)$ total space.
@@
 	}
-	$x\leftarrow\log_2(r)$\tcp*{$s$-Sparse recovery to recovery \& sample from}
+	$x\leftarrow\log_2(r)$\tcp*{$s$-Sparse recovery to recovery and sample from}
 	$a\leftarrow$returned from $S_x$\tcp*{Set of returned non-zero indices}
@@
 \textit{Proof} In an unweighted insertion-deletion streams $\Delta=-1$ for a deletion instruction and $\Delta=1$ for an insertion instruction.
-In the case where a vector is exactly 0-sparse (\ie is the zero vector) the result trivially fails since there is no non-zero index for $\iota$ to hold.
+In the case where a vector is exactly 0-sparse (\textit{i.e.} is the zero vector) the result trivially fails since there is no non-zero index for $\iota$ to hold.
 Consider a vector $v$ with non-zero index $i$.
@@
 Since each deletion cancels out an insertion there is only one insertion left.
-This means $\phi=1$ and $\iota=1\cdoti=i$, the result holds.\\
+This means $\phi=1$ and $\iota=1\cdot i=i$, the result holds.\\
 Suppose $v$ is exactly $s$-sparse with $s>1$.
@@
 \begin{figure}[H]
-	\label{Figure 14}
+	\label{Figure 15}
 	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/facebookDeletionInitialTime.png}}}
@@
 \begin{figure}[H]
-	\label{Figure 15}
+	\label{Figure 16}
 	\subfloat[Time Taken]{{\includegraphics[width=.5\textwidth]{img/facebookDeletionSamplersPerVertexTime.png}}}
@@
 
+\par I will briefly discuss the na\"ive algorithm as it may be useful for future projects.
+\textbf{Algorithm 11} is very similar to the na\"ive implementaiton given for insertion-only streams \textbf{Algorithm 4} in that it records the neighbourhood for every vertex.
+However, it does not have the efficiency of terminating after find the first sufficient neighbourhood as it is possible for that neighbourhood to shrink later in the stream.
+This method will see no significant variation in performance when $c$ is varied as it will produce the same set of neighbourhoods in every case.
+This method uses $\mathcal{O}(n^2)$ space for all values of $c$.
+
 \chapter{Conclusions}
@@
 The run-times of my final implementation are very usable with solutions being found for a graph of over 30 million edges in a couple of minutes for close approximations $c\in[2,4]$.
-\par Due to difficulties in generating functions from a $k$-wise independent hashing family the proposed algorithm for insertion-deletion graphs could not be evaluated on graphs which were sufficiently large as to be interesting.
+\par My main contribution to the proposed insertion-deletion algorithm is the demonstration of how limiting a poor implementation of $k$-wise independent hash family is to the practicalities of this algorithm.
+Research into possible implementations of these hashing families would consistute a significant and useful future project.
+\par Due to the difficulties in generating functions from a $k$-wise independent hashing family the proposed algorithm for insertion-deletion graphs could not be evaluated on graphs which were sufficiently large as to be interesting.
 However, I still managed to discuss several improvements to the implementation.
@@
 The probabilistic approach taken to reduce the number of $L_0$ samplers run for each vertex significantly improved performance by reducing redundancy, without compromising success.
-The indexing method which mapped edges to unique values in $[0,\frac12n(n-1))$ was a really good improvement as it is the optimal indexing method for undirected graphs.
-Although, the introduction of a mathematical hashing function rather than a look up table would negate this improvement.
+The indexing method which mapped edges to unique values in $\left[0,\frac12n(n-1)\right)$ was a really good improvement as it is the optimal indexing method for undirected graphs.
 
@@
 	\addcontentsline{toc}{chapter}{Bibliography}
-	\bibitem{orig} Christian Konrad \textit{Streaming Frequent Items with Timestamps and Detecting Large Neighborhoods in Graph Streams.} November 2019
+	\bibitem{orig} Christian Konrad \textit{Streaming Frequent Items with Timestamps and Detecting Large Neighborhoods in Graph Streams.} November 2019.
 	\bibitem{SNAP}http://snap.stanford.edu/data/
-	\bibitem{L0Framework} Graham Cormode, Donatella Firmani \textit{A Unifying Frameworkd for $\ell_0$-Sampling Algorithms.} September 2014
-  \bibitem{socialNetworkDistribution} Tushti Rastogi \textit{A power law approach to estimating fake social network accounts.} May 2016
+	\bibitem{L0Framework} Graham Cormode, Donatella Firmani \textit{A Unifying Frameworkd for $\ell_0$-Sampling Algorithms.} September 2014.
+  \bibitem{socialNetworkDistribution} Tushti Rastogi \textit{A power law approach to estimating fake social network accounts.} May 2016.
+  \bibitem{GraphStreamSurvey} Andrew McGregor \textit{Graph Stream Algorithms: A Survey.} May 2014.
+  \bibitem{clusteringCofficient} Madhav Jha, C.
+Seshadhri, and Ali Pinar \textit{A space-efficient streaming algorithm for estimating transitivity and triangle counts using the birthday paradox.} February 2015.
+  \bibitem{edgesOfSpanner} B\'ela Bollob\'as \textit{Extremal Graph Theory.} 1978
+  \bibitem{spanner} Surender Baswana \textit{Streaming Algorithm for Graph Spanners - Single Pass and Constant Processing Time per Edge.} April 2008.
+  \bibitem{RandomWalks} Atish Das Sarma, Sreenivas Gollapudi, and Rina Panigraphy \textit{Estimating PageRank on Graph Streams.} May 2011.
+  \bibitem{degResSampling} Jeffery Vitter \textit{Random Sampling with a Reservoir.} Match 1985.
 \end{thebibliography}

### 2020-05-13 paper/src/paper.tex
Submitted

@@
 
+\usepackage{adjustbox}
 \usepackage[ruled, vlined, linesnumbered]{algorithm2e}
@@
 
+\renewcommand\thetable{\arabic{table}} % figure numbering strategy
+\renewcommand{\thefigure}{\arabic{chapter}.\arabic{section}.\arabic{figure}}
+\pagenumbering{roman}
+
 \renewcommand{\headrulewidth}{0pt}
@@
 
-\renewcommand\thetable{\arabic{table}} % figure numbering strategy
-\renewcommand{\thefigure}{\arabic{chapter}.\arabic{section}.\arabic{figure}}
-
-\pagenumbering{roman}
-
 % Header
@@
 % Title
-\title{Implementing and Evaluating Space Efficient Algorithms for Detecting Large Neighbourhoods in Graph Streams}
-\author{Dom Hutchinson}
-\date{\today}
-\maketitle
+
 \newpage
+\thispagestyle{empty}
+
+\adjustbox{padding={5pt},frame={1pt},right}{Dissertation Type: Research}
+
+\vspace*{\fill}
+\begin{center}
+  \includegraphics[scale=1]{img/logo_uob_colour.PNG}\\
+  \vspace*{1.0cm}
+  DEPARTMENT OF COMPUTER SCIENCE\\
+  \vspace*{2.0cm}
+  {\LARGE Implementing and Evaluating Space Efficient Algorithms for Detecting Large Neighbourhoods in Graph Streams} \\
+  \vspace*{1.0cm}
+  {\Large Dominic Hutchinson}\\
+
+  \vspace*{1.0cm}
+  \rule{0.5\textwidth}{0.5pt}
+  \vspace*{1.0cm}
+
+  A dissertation submitted to the University of Bristol in accordance with the requirements of the degree of Master of Engineering in the Faculty of Engineering.
+
+  \vspace*{1.0cm}
+  \rule{0.5\textwidth}{0.5pt}
+  \vspace*{1.0cm}
+
+  May 9, 2020
+\end{center}
+\vspace*{\fill}
 
 \chapter*{Declaration}
+\par This dissertation is submitted to the University of Bristol in accordance with the requirements of a third year project for the degree of MEng in the Faculty of Engineering.
+It has not been submitted for any other degree or diploma of any examining body.
+Except where specifically acknowledged, it is all the work of the Author.
+\par\vspace{5cm}
+\par\noindent Dominic Hutchinson, Saturday $9^\text{th}$ May 2020.
+
+\section*{Ethics}
+\par This project does not require ethics approval.
 
@@
 
-\par The \texttt{Neighbourhood Detection Problem} is a problem in graph theory which tasks one with finding a vertex in the graph which has a certain number of neighbours and to then return a subset of these neighbours.
+\par \texttt{Neighbourhood Detection} is a problem in graph theory which tasks one with finding a vertex in the graph which has a certain number of neighbours and to then return a subset of these neighbours.
 This is a trivial problem to solve in theory, but in practice is hard to solve within a time frame which makes the implementation practical.
-A major reason for this is that many approaches to solving the problem require a lot of space and so quickly overflow a computer's RAM, slowing the execution significantly down.
+A major reason for this is that many approaches to solving the problem require a lot of memory and so quickly overflow a computer's RAM, slowing the execution significantly down.
 This common space inefficiency is becoming increasingly problematic since the move towards Big Data.
 
-\par In this paper I discuss and implement the two algorithms presented in \cite{orig} which offer space efficient solutions for \texttt{neighbourhood detection} for two types of graph stream.
-During the implementation I discuss and evaluate several methods for improving performance of these algorithms, focusing on space-efficiency in order to maximise the potential size of graphs the implementations are effective on.
+\par In this paper I discuss and implement the two algorithms presented in \cite{orig} which offer space efficient solutions for \texttt{Neighbourhood Detection} for two types of graph stream.
+During the implementation I discuss and evaluate several methods for improving the practical performance of these algorithms, focusing on space-efficiency in order to maximize the potential size of graphs the implementations are effective on.
 
 \par The final implementation of the proposed algorithm for insertion-only streams proves very promising with it able to find large neighbourhoods in a couple of minutes for a graph with 30 million edges.
-The algorithm for insertion-deletion streams proves much more difficult to evaluate due to restrictions incurred from the implementation of $L_0$ samplers.
+The algorithm for insertion-deletion streams proves much more difficult to evaluate due to restrictions incurred from the implementation hash functions for $L_0$ samplers.
 
 \chapter*{Acknowledgments}
+
+I would like to thank Dr Christian Konrad for being an enthusiastic and engaging supervisor throughtout this project.
+For the help and guidance he offered throughout.
+In particular, I would like to thank him for being willing to supervise this project after my original supervisor was unable to do so.
 
@@
 % The commercial movement towards data science and the desire to find relationships has lead to big data....
-\par The movement towards Big Data has lead to ever larger data sets which traditional offline algorithms cannot work with efficiently due to their high memory-requirements.
-As the size of data-sets is increasing at a much greater rate than standard RAM capacities means that virtual memory has to be used for these algorithms, significantly increasing their run-times.
+\par The movement towards Big Data has lead to ever larger data sets, which traditional offline algorithms cannot work with efficiently due to their high memory-requirements.
+Moreover, the size of data-sets is increasing at a much greater rate than standard RAM capacities meaning that virtual memory has to be used for these algorithms, significantly increasing their run-times.
+
 \par The \textit{Data Stream Model} was introduced as a method for coping with these large data sets.
-In the \textit{Data Stream Model} instructions are received sequentially, as a `stream', and generally random access is not allowed.
-\textit{Graph Streams} are a common implementation of this model, where instructions describe updates to the edge set of a graph and are typically ordered as a time-series.
+In the data stream model instructions are received sequentially, as a \textit{`stream'}, and generally random access is not allowed.
+\textit{Graph Streams} are a popular implementation of this model, where instructions describe updates to the edge set of a graph and are typically ordered as a time-series.
 This allows for a dynamic representation of a graph, which typical \texttt{adjacency list} implementations do not.
-%TODO citation=https://en.wikipedia.org/wiki/Streaming_algorithm#Data_stream_model
-\par Streaming algorithms are a class of algorithms, first formalised in 1996 \cite{TODO}, to work within the \textit{Data Stream Model}.
+Graph streams are ideal for representing changing networks such as a social or computer network where objects can connect and then disconnect.
+
+\par Streaming algorithms are a class of algorithms, first formalised in 1996 \cite{graphStreams}, to work within the data stream model.
 Streaming algorithms are typically designed to work in $polylog(n)$, where $n$ is the number of vertices in the case of graph streams.
 Streaming algorithms are designed to work in as few passes of the stream as possible, preferable only a single-pass is used as this means the stream can easily be extended without the algorithm having to start over.
-Due to the limited space streaming algorithms are allocated, they are often produce an approximation of a solution in order to improve space efficiency.
-When implementing a streaming algorithm the trade-off between accuracy and space-usage needs to be considered.
+Due to the limited space streaming algorithms are allocated, they often produce an approximation of a solution in order to improve space efficiency.
+When implementing a streaming algorithm the trade-off between accuracy, number of passes and space-usage needs to be considered.
 
 \par Graphs are used to represent relationships between objects and are popular in data science.
-A graph stream is a sequence of instructions which describe how to construct a graph.
-The sequential nature of a graph stream allows it to represent the evolution of a graph over time, which adjacency matrices cannot.
-This makes graph streams ideal for representing changing networks such as a social or computer network where objects can connect and then disconnect.
-
-\par A basic problem in graph theory is \texttt{High Degree Detection} where you want to find the vertices with the greatest degree.
+A basic problem in graph theory is \texttt{High Degree Detection} where you want to find the vertices with the greatest degree.
 This problem can be trivially and efficiently solved by counting the number of edges incident to each vertex and then returning any vertices of sufficient degree.
@@
 
-\par\texttt{Neighbourhood Detection} is the natural set up from \texttt{High Degree Detection}, it tasks you with finding vertices of high degree \underline{and} a subset of their neighbours.
+\par\texttt{Neighbourhood Detection} is the natural set on from \texttt{High Degree Detection}, it tasks you with finding vertices of high degree \underline{and} a subset of their neighbours.
 Solving \texttt{Neighbourhood Detection} allows for inferences to be made about the influence a vertex has.
@@
 
-\par In this paper I shall discuss two streaming algorithms which solving \texttt{Neighbourhood Detection} in $polylog(n)$ space.
+\par The temporal-ordering of instructions in a graph stream is a hindrance for some problems, including \texttt{Neighbourhood Detection}, as it means edges likely do not arrive in the most useful order and very few assumptions can be made about what the next instruction will be.
+Suppose instructions were ordered by one of their incident vertex names, then we could infer when the last occurence of a vertex in the stream has occured and discard unnecessary information about that vertex.
+
+\par In this paper I discuss two streaming algorithms which solve \texttt{Neighbourhood Detection} in $polylog(n)$ space.
 This is achieved by using a series of samplers in order to reduce the set of edges which need to be analysed.
-This means that there is a possibility of failure and this possibility increases as less edges are sampled, but there is a trade-off here with space efficiency.
-%TODO move this sentence & rewrite
-Using graph streams is good here as it allows for more space efficient algorithms but due to time-ordered nature of most graph streams their is an added difficulty that instructions do not arrive in the most optimal order for solving \texttt{Neighbourhood Detection}.
+This means that there is a possibility of failure and this possibility increases as fewer edges are sampled, but there is a trade-off here with space efficiency.
 
@@
 \section{Objectives}
-%TODO expand
 The objectives of this project are to:
@@
 	\item Use large graphs from real-world scenarios to evaluate the practicalities of these implementations on real-world data sets.
-  \item Discuss how graphs from different real-world sources (\ie not social networks) might perform on these implementation and how the implementations could be changed to accomodate for them.
+  \item Discuss how graphs from different real-world sources (\ie not social networks) might perform on these implementations and how the implementations should be changed to accommodate for them.
 \end{itemize}
@@
 
-Since the formalisation of the data streaming model in 1996 much research has been done into solving problems using this model. Here I give an overview of previous works in this field.\newline
+Since the formalisation of the data streaming model in 1996 much research has been done into solving problems using this model.
+Here I give an overview of some of these problems and the work which has been done on them.\newline
 
@@
 \par\noindent\textbf{Minimum Distance between two nodes}
-\par A $\alpha$-\textit{spanner} subgraph $H$ of a graph $G=(V,E)$ is a subgraph where $\forall\ u,v\in V$ $d_G(u,v)\leq d_H(u,v)\leq\alpha\cdot d_G(u,v)$ where $d_G(\cdot,\cdot)$ is the length of the shortest path between the given nodes in $G$.
+\par An $\alpha$-\textit{spanner} subgraph $H$ of a graph $G=(V,E)$ is a subgraph where $\forall\ u,v\in V$ $d_G(u,v)\leq d_H(u,v)\leq\alpha\cdot d_G(u,v)$ where $d_G(\cdot,\cdot)$ is the length of the shortest path between the given nodes in $G$.
 Spanners are a space-efficient structure for evaluating the length of shortest path between any two nodes in a graph.
@@
 \par A random walk from $v_0\in V$ in a graph is any random sequence of vertices $v_0,v_1,v_2,\dots$ where $v_i$ is randomly selected from the neighbourhood of $v_{i-1}$.
-The distribution of $v_t$ can be assessed.
+The distribution of $v_t$ can be assessed to make inferences about the structure of the graph.
 \par Das Sarma et al.
@@
 In \texttt{Chapter 3} I discuss, implement and evaluate the algorithm proposed in \cite{orig} for insertion-deletion graph streams.
-And, in \texttt{Chapter 4} review what was achieved in this project and provide some ideas for possible future work to build upon those achievements.
+And, in \texttt{Chapter 4} I review what was achieved in this project and provide some ideas for possible future work to build upon those achievements.
+
+\section{Code}
+A git repository is available containing the code for this project.\\
+\texttt{https://github.com/dajhutchinson/Detecting-Large-Neighbourhoods-in-Graph-Streams}
 
@@
 
-\par A graph $G=(V,E)$ is said to be a \textit{Bipartite Graph} if its vertex set $V$ can be partitioned into two disjoint subsets $A,B$ and the every edges connects a vertex in $A$ to a vertex in $B$.
+\par A graph $G=(V,E)$ is said to be a \textit{Bipartite Graph} if its vertex set $V$ can be partitioned into two disjoint subsets $A,B$ such that every edge connects a vertex in $A$ to a vertex in $B$.
 Formally, $\exists\ A,B\subseteq V$ st $A\cup B=V,\ A\cap B=\emptyset$ and $\forall\ (a,b)\in E$ we have that $a\in A,b\in B$.
@@
 
-\par The \textit{Neighbourhood} of a vertex $v$ is the set of vertices which share an edge with, $N_v=\{u:(u,v)\in E\}$.
-The \textit{Degree} of a vertex $v$ is the size of its neighbourhood, $\delta_v=|N_v|$.
+\par The \textit{Neighbourhood} of a vertex $v$ is the set of vertices which share an edge with $N_v=\{u:(u,v)\in E\}$.
+The \textit{Degree} of a vertex $v$ is the size of its neighbourhood $\delta_v=|N_v|$.
 A vertex $v$ is \textit{$s$-sparse} if its degree is less than, or equal to, $s$.
-\par The \textit{Edge Vector} of a vertex $v$ of length $|V|$ where each index gives the weight of the edge from $v$ to the vertex associated with that index.
+The \textit{Degree Distribution} of a graph is a distribution of the degrees of the vertices in a graph.
+\par The \textit{Edge Vector} of a vertex $v$ is a vector of length $|V|$ where each index gives the weight of the edge from $v$ to the vertex associated with that index.
 Typically a weight of 0 is stored if no edge exists between $v$ and a specific vertex.
 The graphs used in this project are unweighted so the values in the edge vector are booleans describing whether, or not, an edge exists.
-These edge vectors can be stored as bitstrings.
+These edge vectors can be stored as \texttt{bitstrings}.
 \par \textit{Graph Streams} are an unbounded sequence of instructions $\{e_1,e_2,\dots\}$ which describe how to construct the graph.
-These instructions $e_i$ describe modifications to the edges of a graph.
+These instructions $e_i$ describe modifications to the edge set of the graph.
 The vertex set of the described graph is inferred from the instructions.
 In this project we are only concerned with two types of graph stream.
-%TODO something about the unbounded part??
 \begin{itemize}
@@
 \par An \textit{$\alpha$-Approximation Algorithm} is an algorithm which returns a result within an $\alpha$ factor to the optimal solution to the problem.
-Algorithms which solve the \texttt{Neighbourhood Detection Problem} are $\frac1c$-approximation algorithms since they only return $d/c$ of the neighbours of a vertex which we know to have at least $d$ neighbours.
+Algorithms which solve \texttt{Neighbourhood Detection} are $\frac1c$-approximation algorithms since they only return $d/c$ of the neighbours of a vertex which we know to have at least $d$ neighbours.
 
@@
 For the implementations in this project I chose to use C++, specifically \texttt{C++11}.
-I initially considered using either python or C++, as they both work in the object-orientated paradigm and I am proficient in both.
+I initially considered using either Python or C++, as they both work in the object-orientated paradigm and I am proficient in both.
 I chose C++ as it allows for more control with memory management and typically has faster run-times.
@@
 
-During evaluation three performance metrics were measured: success-rate, time-taken, and, space-used.
+During evaluation three performance metrics were measured: success-rate, time-taken, and space-used.
 These are all interconnected and come with their own trade-offs.
@@
 For this reason $d$ is set to the maximum degree in the graph during all tests.
-Further, the greater the proportion of a neighbourhood returned the better the inferences made are so I focussed testing on low values of $c$, namely $c\in[1,20]$.
+Further, the greater the proportion of a neighbourhood returned the better the inferences made are so I focussed testing on low values of $c$, namely $c\in[2,20]$.
 
@@
 I chose to use the graphs generated from social networks as I found that application most motivating, but this choice is essentially arbitrary.
-The graph streams from \cite{SNAP} were all insertion-only so I wrote a utility program which created insertion-deletion streams from them by using a Bernoulli random variable to decide whether to delete an edge immediately after inserting it.
+The graph streams from \cite{SNAP} were all insertion-only so I wrote a function in \texttt{utility.cpp} which created insertion-deletion streams from them by using a Bernoulli random variable to decide whether to delete an edge immediately after inserting it.
 \textbf{Table 1} provides details about the graphs used during evaluation.
@@
 \end{center}
+
+\par I wrote a function \texttt{utility.cpp} which took the returned vertex $v$ and set of neighbours $A$ and validated whether the final degree of $v$ was at least $d$ and that $A$ was in fact a subset of the final neighbourhood of $v$.
 
@@
 This invariant is proved as \textbf{Lemma 2.1}, first I shall describe how the reservoir is maintained by controlling how vertices are inserted.
-The first time the degree counter for a vertex $v$ surpasses $d_1$, $v$ is considered for insertion into the \textit{reservoir}.
-There are two cases
+
+\par The first time the degree counter for a vertex $v$ surpasses $d_1$, $v$ is considered for insertion into the \textit{reservoir}. There are two cases
 \begin{itemize}
@@
 If this random variable succeeds then pick, uniformly at random, a vertex $u$ currently in the reservoir and replace it with $v$.
-All the edges in $E$ which are incident to $u$, and no other vertex in the reservoir, are removed from $E$.
+All the edges in $E$ which are incident to $u$, and no other vertex in $R$, are removed from $E$.
 \end{itemize}
 
-Whenever an edge $(u,v)$ is encountered and $u$ is in the reservoir and the number of edges stored in $E$ incident to $u$ is less than $d_2$ then $(u,v)$ is inserted into $(u,v)$. %TODO rewrite this
-\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+\noindent Whenever an edge $(u,v)$ is encountered and $u$ is in the reservoir and the number of edges stored in $E$ incident to $u$ is less than $d_2$ then $(u,v)$ is inserted into $E$.\newline
+
+\noindent\fbox{\parbox{\textwidth}{
 
@@
 Hence by mathematical induction, $R$ contains a uniform sample of the vertices of sufficient degree at all points in time.$\hfill\square$
-}}\end{adjustwidth}\vspace{.3cm}
-%TODO adding termination condition? Maybe add during implementation
+}}\newline
 
@@
 
-\texttt{Neighbourhood Detection} can be solved using \texttt{degree-based reservoir sampling} by setting $d_2=\frac{d}c$.
+\texttt{Neighbourhood Detection} can be solved using \texttt{Degree-Based Reservoir Sampling} by setting $d_2=\frac{d}c$.
 The question remains as to what to set $d_1$ and $s$ to in order to achieve a high probability of success.
-\textbf{Algorithm 2} is the space-efficient algorithm proposed in \cite{orig} for solving \texttt{Neighbourhood-Detection} for insertion-only streams.
+\textbf{Algorithm 2} is the space-efficient algorithm proposed in \cite{orig} for solving \texttt{Neighbourhood Detection} for insertion-only streams.
 
 \begin{algorithm}
-	\caption{One-pass $c$-Approximation Insertion-Only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
+	\caption{One-pass $c$-Approximation Insertion-only Streaming Algorithm for $\mathtt{Neighbourhood\ Detection}$}
 	\SetKwInOut{Require}{require}
@@
 	\For{$i\in[0,c-1]$\text{\textbf{ in parallel}}} {
-		$(a_i,S_i)\leftarrow\mathtt{Deg}\mbox{-}\mathtt{Res}\mbox{-}\mathtt{Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
-	}
-	\Return{\small Uniform random neighbourhood $(a_i,S_i)$ \text{from successful runs}}
+		$(a_i,S_i)\leftarrow$\texttt{Degree-Based Reservoir Sampling}$\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
+	}
+	\Return{Uniform random neighbourhood $(a_i,S_i)$ from successful runs}
 \end{algorithm}
 
-\par In \cite{orig} it is proven that setting $s=\left\lceil\ln(n)\cdot n^{\frac1c}\right\rceil$ and using $c$ \texttt{degree-based reservoir samplers} each with a different lower bound, incremented from $\frac{d}c$ to $d$ stepping by $\frac{d}c$ each time, results in a high probability of success.
+\par In \cite{orig} it is proven that setting $s=\left\lceil\ln(n)\cdot n^{\frac1c}\right\rceil$ and using $c$ \texttt{Degree-Based Reservoir Samplers} each with a different lower bound, incremented from $\frac{d}c$ to $d$ stepping by $\frac{d}c$ each time, results in a high probability of success.
 The samplers are run in parallel so that the algorithm only requires a single pass of the stream.
-Implementing parallel running of the samplers just requires passing the same instruction to each sample before fetching the next instruction.
+Implementing parallel running of the samplers requires passing the same instruction to each sample before fetching the next instruction.
 
 \par Each of the $c$ samplers requires $\mathcal{O}(n\log n+sd_2\log n)$ space.
-Since the degree map can be shared between samplers the total space is $\mathcal{O}(n\log n+c\cdot sd_2\log n)=\mathcal{O}(n\log n+n^{\frac1c}d\log^2n)$.
-The proposed algorithm does require you to know the number of vertices in the graph stream before running the algorithm, or at least the $\log$ of it.
+Since $D$ can be shared between samplers the total space is $\mathcal{O}(n\log n+c\cdot sd_2\log n)=\mathcal{O}(n\log n+n^{\frac1c}d\log^2n)$.
+The \textbf{Algorithm 2} does require you to know the number of vertices in the graph stream before running the algorithm, or at least the $\log$ of it.
 If this is unknown then it is quick to run through the whole stream and build a set of vertices in the graph.
@@
 \par The graphs from \cite{SNAP} are not formatted in a strictly bipartite way.
-I adjusted \texttt{degree-based reservoir sampling} to account for this by repeating \texttt{lines 6-17} (and the early termination clause) of \textbf{Algorithm 1} after the end of the \texttt{for} loop, with all occurrences of $a_i$ replaced by $b_i$.
+I adjusted \texttt{Degree-Based Reservoir Sampling} to account for this by repeating \texttt{lines 6-17} of \textbf{Algorithm 1} after the end of the \texttt{for} loop, with all occurrences of $a_i$ replaced by $b_i$.
 
@@
 \texttt{Figure 2.3.1} shows the results of testing the initial implementation of \textbf{Algorithm 2}.
-These results shows that the space requirements of \textbf{Algorithm 2} are significantly less than the size of file stream and that the space requirements decrease as the approximation factor increases but there is an undesirable result that the time taken increases as the approximation factor increases.
+These results show that the space requirements of \textbf{Algorithm 2} are significantly less than the file size of the stream (12MB for \texttt{gplus}) and that the space requirements decrease as the approximation factor increases but there is an undesirable result that the time taken increases as the approximation factor increases.
 
@@
 \textbf{Algorithm 1} states that the whole stream should be run through before the sampler returns a result uniformly from those which succeeded.
-This is unnecessary to fulfil \texttt{neighbourhood detection} as returning the first successful result would be sufficient.
+For insertion-only streams this is unnecessary to fulfil \texttt{Neighbourhood Detection} as returning the first successful result would be sufficient.
 Similarly \textbf{Algorithm 2} states that all samplers should be run to completion and then a result is returned from those which succeeded.
-Again, this is unnecessary as return the first encountered neighbourhood of sufficient size would suffice for solving \texttt{neighbourhood detection}.
+Again, this is unnecessary as returning the first encountered neighbourhood of sufficient size would suffice for solving \texttt{Neighbourhood Detection}.
 These early terminations can be implemented by terminating the algorithm after it finds a vertex of degree $d_1+d_2-1$ which is in a reservoir (remembering that $d_1$ is different for each reservoir).
@@
 
-\par \texttt{Figure 2.3.2} shows the results implementing early termination, against not implementing it, for graph \texttt{gplus} at different approximation factors.
+\par \texttt{Figure 2.3.2} shows the results of implementing early termination, against not implementing it, for graph \texttt{gplus} at different approximation factors.
 The time taken now decreases as $c$ increases and is more than halved in all cases.
@@
 Both metrics now converge quickly, to $\smallsim\!\!1.5$ seconds and $3.3$MBs respectively, meaning we can make predictions on the performance of the algorithm for greater values of $c$.
-These are really good results, so I implemented this change going forward.
-%TODO explain reasons behind these results
+These are really good results, so I implemented early termination going forward.
 %NOTE space converged to is that used by degree map
-%TODO comment about effect on space used by reservoirs vs degree map
-%TODO this change means that time taken is now a measure of success as failure is (basically) the only time the whole stream is run through
 
@@
 
-\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
+\fbox{\parbox{\textwidth}{
 
@@
 % TODO a proof by contradiction
-\textit{Proof (by contradiction)} Let $v$ be a vertex in the graph stream.
-Suppose, without loss of generality, that $v$ is a member of the reservoirs of both \texttt{Deg-Res-Sampling}$(\frac{d}{c},\frac{d}{c},s)$ and \texttt{Deg-Res-Sampling}$(\frac{2d}{c},\frac{d}{c},s)$.
+\textit{Proof (by contradiction)} Let $v\in V$.
+Suppose, without loss of generality, that $v$ is a member of the reservoirs of both \texttt{Degree-Based Reservoir Sampling}$(\frac{d}{c},\frac{d}{c},s)$ and \texttt{Degree-Based Resservoir Sampling}$(\frac{2d}{c},\frac{d}{c},s)$.
 This means that $v$ was sampled twice, after encountering the $\frac{d}c^\text{th}$ \underline{and} $\frac{2d}c^\text{th}$ edge incident to $v$ in the stream.
@@
 
-}}\end{adjustwidth}\vspace{.3cm} %TODO refine
+}}
 
@@
 Namely, a data-structure is required to store \underline{precisely} which reservoir each vertex is in so that the appropriate value of $d_1$ is used in \texttt{Algorithm 1:Line 18}, otherwise improper termination could occur.
-An appropriate data-structure for this would be a \texttt{map} from each vertex to a bitstring whose indices indicate whether the vertex lies in an associated reservoir.
-With this implementation, inserting or deleting a vertex to/from a reservoir and querying whether a vertex is in a given reservoir is achieved with a single query of a \texttt{map} and then a known bit in a bitstring which both take $\mathcal{O}(1)$ time.
+An appropriate data-structure for this would be a \texttt{map} from each vertex to a \texttt{bitstring} whose indices indicate whether the vertex lies in an associated reservoir.
+With this implementation, inserting or deleting a vertex to/from a reservoir and querying whether a vertex is in a given reservoir is achieved with a single query of a \texttt{map} and then a known bit in a \texttt{bitstring} which both take $\mathcal{O}(1)$ time.
 This \texttt{map} replaces the \texttt{ordered set}s currently used to store the reservoirs of each sampler and thus reduces these action times from $O(\log n)$ to $\mathcal{O}(1)$ time.
@@
 there is a slight increase in space usage for low values of $c$ but this difference disappears very quickly so is of little concern.
-Implementing a shared-edge set does not change the logic of the algorithm so does not affect its success rate.
-Due to the reduction in mean run-time and no noteworthy changes in space requirements, I recommend implementing this change.
+Implementing a shared-edge set did not change the logic of the algorithm so does not affect its success rate.
+Due to the reduction in mean run-time and no noteworthy changes in space requirements, I recommend implementing a shared-edge set.
 
@@
 
-\par The changes discussed in \textbf{Section 2.4} have been to the technical implementation of \textbf{Algorithm 2}, rather than to the its logic.
-This means these changes have had no affect on the success rate of the algorithm.
+\par The changes discussed in \textbf{Section 2.3} addressed the technical implementation of \textbf{Algorithm 2}, rather than to the its logic.
+This meant these changes had no affect on the success rate of the algorithm.
 I will now discuss potential changes to the logic of the proposed algorithm.
@@
 
-\par When changing the number of samplers we need to consider their lower bounds $d_1$ too.
-If the range of vertex degrees $[d_1,d_1+d_2]$ which individual samplers sample at overlap then duplicate sampling of vertices can occur which offers no advantages and leads to a reservoir space being wasted.
-And, if there are gaps between these ranges then there are vertices which may not be sampled.
-%TODO flesh this out more.
-Thus, the best set up is to have the sampling ranges be sequential with no gaps and so when varying the number of samplers we should only add or remove from the endpoints of these ranges.
+\par When changing the number of samplers their lower bounds $d_1$ needs to be considered too.
+If the range of vertex degrees $[d_1,d_1+d_2]$ which individual samplers sample at overlap then duplicate sampling of vertices can occur which offers no advantages and leads to reservoir space being wasted.
+And, if there are gaps between these ranges then there are vertices which may not be sampled, which offers no advantages as each vertex is assumed to be equally likely to have a sufficient neighbourhood.
+Thus, the best set up is to have the sampling ranges be sequential with no gaps and so when varying the number of samplers, samplers should only be added or removde from the endpoints of these ranges.
 
@@
 	\includegraphics[width=.5\textwidth]{img/gplusSuccessRate.png}
-	\caption{Success rate when only the first $n$ samplers are run on \texttt{gplus}.}
+	\caption{Success rate when only the first $n$ samplers are run on \texttt{gplus}, for different approximation factors $c\in[2,20]$.}
 	%"Final Evaluation - insertion-only"/Master
@@
 \par \texttt{Figure 2.4.6} shows the results when using the first $\mathtt{min}\left(c,\mathtt{max}\left(2,\left\lceil\frac15\ln n\right\rceil\right)\right)$ samplers against just using the technical improvements discussed in \texttt{Section 2.3}.
-As seen in \texttt{Figure 2.4.4}, the space usage is unchanged while the time taken is down in all cases.
+As seen in \texttt{Figure 2.4.4}, the space usage is unchanged while the time taken is decreased in all cases.
 Run time is reduced by over $50\%$ for $c\geq 12$ and is below 1 second for larger values of $c$.
@@
 
-\par Decreasing the reservoir size $s$ means less edges are stored, reducing space requirements, but increases the turn-over rate of the elements in the reservoir as the probability a vertex is sampled is the same but the reservoir size is smaller.
+\par Decreasing the reservoir size $s$ means fewer edges are stored, reducing space requirements, but increases the turn-over rate of the elements in the reservoir as the probability a vertex is sampled is the same but the reservoir size is smaller.
 The greater turn-over rate is detrimental to the success of the algorithm as it more likely for a vertex to be removed when it is close to succeeding.
@@
 	\For{$i\in[0,n_s]$\text{\textbf{ in parallel}}} {
-		$(a_i,S_i)\leftarrow\mathtt{Deg}\mbox{-}\mathtt{Res}\mbox{-}\mathtt{Sampling}\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
-	}
-	\Return{\small First neighbourhood of sufficient degree and its root}
+		$(a_i,S_i)\leftarrow$\texttt{Degree-Based Reservoir Sampling}$\left(\max\left\{1,i\cdot\frac{d}{c}\right\},\frac{d}c,s\right)$
+	}
+	\Return{First neighbourhood of sufficient degree and its root}
 \end{algorithm}
@@
 	\Require{Stream $\{(s_0,t_0)\dots(s_n,t_n)\}$, degree bound $d$, precision bound $c$}
-	$N\leftarrow\{\{\}\}\ \{\text{neighbourhoods}\}$\\
+	$N\leftarrow\{\{\}\}$\tcp*{Neighbourhoods}
 	\For{$i=0\dots n$} {
@@
 
-\textbf{Algorithm 4} is a naïve algorithm for solving \texttt{neighbourhood detection} for insertion-only graph streams.
+\textbf{Algorithm 4} is a naïve algorithm for solving \texttt{Neighbourhood Detection} for insertion-only graph streams.
 This approach records the neighbourhood of every encountered vertex and returns the first one which surpasses the $\frac{d}c$ threshold.
@@
 As the naïve algorithm terminates at the first neighbourhood of sufficient size and uses all edges in the stream up to that point, it is guaranteed to find the first vertex with a sufficient neighbourhood.
-This means the naïve algorithm is very quick, provided its space-usage is within the memory allocated to the algorithm.
+This means the naïve algorithm is very quick, provided its space-usage remains within the memory allocated to the algorithm.
 In the worst case, where all elements have degree $\frac{d}c-1$ except for the last which has degree $\frac{d}c$, $\frac{d}{c}+(n-1)(\frac{d}c-1)\in \mathcal{O}(n\frac{d}c)$ space is required.
@@
 However, after three hours of computation the naïve algorithm had failed to return a result for the largest graph \texttt{gplus\_large} due to its high space usage.
-For the largest graph tested, \texttt{gplus\_large}, $13.7\%$ of the stream file size was required to find a 2-approximation using the initial implementation and only $1.39\%$ using the final implementation.
-This is a notable decrease in space requirements, meaning the algorithm can run efficiently on even large graphs.
-
-\par The run time of the final implementation is useably quick: when run on \texttt{gplus\_large} it took $\sim\!\!150$ seconds to find a solution when $c=2$ and less than a minute for $c\geq5$.
+For the largest graph tested, \texttt{gplus\_large}, $13.7\%$ of the stream's file size was required to find a 2-approximation when using the initial implementation and only $1.39\%$ when using the final implementation.
+This is a notable decrease in space requirements, meaning the final implementation can run efficiently on very large graphs.
+
+\par The run time of the final implementation is useably quick: when run on \texttt{gplus\_large} it took $\smallsim\!\!150$ seconds to find a solution when $c=2$ and less than a minute for $c\geq5$.
 The naïve algorithm being quicker is irrelevant for this evaluation as it fails to terminate for larger files due to its space inefficiency.
@@
 %TODO citations
-\par\texttt{$L_0$ Sampling} is a technique for sampling, approximately uniformly, from the non-zero indices of a vector when given a stream of updates to the weights of the indices.
+\par\texttt{$L_0$ Sampling} \cite{L0Framework,LpSurvey} is a technique for sampling, approximately uniformly, from the non-zero indices of a vector when given a stream of updates to the weights of the indices.
 Here I shall describe one space-efficient technique for \texttt{$L_0$ Sampling} which succeeds with probability $1-\delta$ and requires $\mathcal{O}\left(\frac1\delta\log(1/\delta\gamma)\log^2(n)\right)$ space, this is significantly less than the $\mathcal{O}(n)$ space required to store the whole vector.
 This technique relies on two subroutines: \texttt{Perfect 1-Sparse Recovery}; and \texttt{Exact $s$-Sparse Recovery}.
-%TODO check space requirements (do I need to mention the two different success rates)
 
@@
 \par \texttt{Exact $s$-Sparse Recovery} recovers a uniform sample of upto $s$ of the non-zero indicies in a vector described by a stream with probability $1-\gamma$.
-\textbf{Algorithm 6} defines a space efficient implementation of \texttt{Exact $s$-Sparse Recovery}.
+\textbf{Algorithm 6} defines a space-efficient implementation of \texttt{Exact $s$-Sparse Recovery}.
 
@@
 %TODO is it a subvector or is it that the described vector is exactly s-sparse
-\par \textbf{Algorithm 6} uses a two-dimensional matrix $M$ of \texttt{Perfect $1$-Sparse Recoveries} and assigns a hash function $g_r$ to each row.
-When an update $(i,\Delta)$ is received, it is passed to the $g_r(i)^\text{th}$ \texttt{$1$-Sparse Recovery} of each row.
-Once the updates have been exhausted the returned elements of the $1$-sparse recoveries are collected into a set $A$ and returned.
+\par \textbf{Algorithm 6} uses a two-dimensional matrix $M$ of \texttt{$1$-Sparse Recoveries} and assigns a hash function $g_r$ to each row.
+When an update $(i,\Delta)$ is received, it is passed to the $g_r(i)^\text{th}$ recovery of each row.
+Once the updates have been exhausted the returned elements of the \texttt{$1$-Sparse Recoveries} are collected into a set $A$ and returned.
 \par $g_r$ maps, on average, $\frac{n}c$ indices to each $1$-sparse recovery.
@@
 $j$ \texttt{Exact $s$-Sparse Recoveries} are instantiated and a $k$-wise independent hash function $h(\cdot)$ (with $k\in\mathcal{O}(s)$) is used to determine which recovery to pass an update to.
-When an update $(i,\Delta)$ is received we update an estimate of the sparsity $r$ of the vector being sampled and pass the update to all $S_k$ where $h(i)\leq\frac{n^3}{2^k}$.
+When an update $(i,\Delta)$ is received an estimate of the sparsity $r$ of the vector being sampled is updated and the update is passed to all $S_k$ where $h(i)\leq\frac{n^3}{2^k}$.
 Once the update stream is consumed we consider the set returned from the $\log(r)^\text{th}$ recovery and return the index which produces the lowest hash-value.
@@
 
-In \cite{orig} two single-pass algorithms are proposed for \texttt{neighbourhood detection} in an insertion-deletion stream, both of which use $L_0$ sampling to acquire edges from the described graph.
+In \cite{orig} two single-pass algorithms are proposed for \texttt{Neighbourhood Detection} in an insertion-deletion stream, both of which use $L_0$ sampling to acquire edges from the described graph.
 
@@
 
-\par As incident edges are sampled uniformly, using $10\frac{d}c\ln n\gg\frac{d}c$ samplers results in a high probability that a neighbourhood size $\frac{d}c$ is found for vertices of degree greater $\frac{d}c$.
-Thus the success of \textbf{Algorithm 8} relies on it sampling a a vertex of sufficient degree.
+\par As incident edges are sampled uniformly, using $10\frac{d}c\ln n\gg\frac{d}c$ samplers results in a high probability that a neighbourhood of size $\frac{d}c$ is found even for vertices with degrees slightly greater than $\frac{d}c$.
+Thus the success of \textbf{Algorithm 8} relies on sampling a vertex of sufficient degree.
 This means \textbf{Algorithm 8} is not effective when there are very few vertices of sufficient degree.
@@
 
-\par During these implementations I made the assumption that vertices were labelled with unique values in $[1,n]$.
+\par During these implementations I made the assumption that vertices were labelled with unique values in $[n]$.
 This simplifies many parts of the implementation as each vertex aligns to an index in the edge vector.
@@
 Suppose $v$ is exactly $s$-sparse with $s>1$.
-Then there exists $s-1$ other indices $j_1,\dots,j_{s-1}\neq i$ which are inserted exactly one more time than it is deleted.
-After letting the deletions cancel the insertions there are $s$ insertion instructions left, this means $\phi=s$ and the result does not hold.
+Then there exists $s-1$ unique indices $j_1\neq\dots\neq j_{s-1}\neq i$ which are inserted exactly one more time than they are deleted.
+After letting the deletions cancel out the insertions there are $s$ insertion instructions left, this means $\phi=s$ and the result does not hold.
 Thus the result holds \underline{iff} $v$ is exactly $1$-sparse.$\hfill\square$
-
-}}\end{adjustwidth}\vspace{.3cm}
+}}\newline
 
 The insertion-deletion streams used in this project are unweighted.
-Thus, by \textbf{Lemma 3.2}, the exact 1-sparse recovery process describded in \textbf{Algorithm 5} can be simplified to the implementation described in \textbf{Algorithm 10}.
-This implementation removes the $\tau$ counter meaning the total space used by the algorithm is reduced by a third which is significant as lots of these processes are run during \textbf{Algorithm 8} and \textbf{Algorithm 9}.
+Thus, by \textbf{Lemma 3.2}, the \texttt{Exact 1-Sparse Recovery} process described in \textbf{Algorithm 5} can be simplified to the implementation described in \textbf{Algorithm 10}.
+This implementation removes the $\tau$ counter meaning the total space used by the algorithm is reduced by a third which is significant as lots of these recoveries are run during \textbf{Algorithm 8} and \textbf{Algorithm 9}.
 Implementing \textbf{Algorithm 10} is trivial.
@@
 
-\par When implementing \texttt{Exact $s$-Sparse recovery} the main decision is how to implement the family of pairwise hash functions efficiently. I used the set of functions
+\par When implementing \texttt{Exact $s$-Sparse Recovery} the main decision is how to implement the family of pairwise hash functions efficiently. I used the family of functions
 $$\mathcal{H}_2:=\{h_{a,b}(x)=((ax+b)\text{mod }p)\text{mod }2s|a,b\in[1,P-1]\}$$
@@
 These hash functions map to the desired range $[0,2s)$.
-This family of hash functions are ideal for this project as can be stored in $O(\log P)$ bits as only the values of $a,b$ need to stored for each function.
-Similarly it is easy to generate a function in this family by sampling twice from $[1,P-1]$.
+This family of hash functions are ideal for this project as each function can be stored in $O(\log P)$ bits as only the values of $a,b$ need to stored.
+Similarly, it is easy to generate a function in this family by sampling twice from $[1,P-1]$.
 
 \par Finding primes is hard so I hard coded $P=1,073,741,789>2^{29}$.
-This is an upper bound on number of columns the \texttt{$s$-Sparse Recovery} algorithm can handle, which is equivalent to placing a lower bound on the acceptable failure rate $\delta$ of the \texttt{$L_0$ Sampling} algorithm.
+This is an upper bound on number of columns the \textbf{Algorithm 6} can handle, which is equivalent to placing a lower bound on the acceptable failure rate $\delta$ of the \texttt{$L_0$ Sampling} algorithm.
 $2\cdot\frac1\delta<P\implies\delta>\frac2P\approx2\times10^{-9}$.
@@
 %NOTE maybe move to implementation
-For this project $L_0$ sampling is used on two different vectors: the edge vector of a specified vertex; and the edge set of the whole graph.
+For this project \texttt{$L_0$ Sampling} is used on two different vectors: the edge vector of a specified vertex; and the edge set of the whole graph.
 The edge set of a graph is often defined as an $n\times n$ adjacency matrix, to apply $L_0$ sampling to this matrix we flatten it into a $n^2$ dimensional vector by reading the rows sequentially.
@@
 These results pass a $\chi^2$ test for a uniform distribution so I implemented this \texttt{map} as the $k$-wise independent hash function.
-As the vertices are assumed to be labelled in $[n]$ this \texttt{map} can be implemented as an array with the hash values stored in each index.
-Using an array is more space-efficient than a tradition \texttt{map}.
+As the vertices are assumed to be labelled in $[n]$ this \texttt{map} can be implemented as an \texttt{array} with the hash values stored in each index.
+Using an \texttt{array} is more space-efficient than a tradition \texttt{map} (I shall refer to this structure as a \texttt{hash array}).
 
 \par The maximum hash value is a constraint on the size of the domain and thus the number of vertices which can be handled.
-In theory this is not a problem, but in practice the limit of this value depends on the data type used.
-I implemented the hash array as an array of \texttt{uint64\_t} (8 bytes) which has the greatest maximum value of the standard c++ numeric data type.
-The maximum domain is the cube root of this value, thus \textbf{Algorithm 8} will not work for graphs with more than $(2^{64}-1)^{1/3}\approx2.6\times10^6$ vertices and \textbf{Algorithm 9} will not work for graphs with more than $(2^{64}-1)^{1/6}\approx1,625$ vertices.
+In theory this is not a problem, but in practice the limit of this value depends on the numeric data type used.
+I implemented the \texttt{hash arrayx} as an \texttt{array} of \texttt{uint64\_t} (8 bytes) which has the greatest maximum value of the standard c++ numeric data type.
+The maximum domain is the cube root of this value, thus \textbf{Algorithm 8} will not work for graphs with more than $(2^{64}-1)^{1/3}\approx2.6\times10^6$ vertices and \textbf{Algorithm 9} will not work for graphs with more than $(2^{64}-1)^{1/6}\approx1,625$ vertices, due to overflow.
 This constraints can be reduced by reducing the relative range of the hash function.
@@
 \texttt{Figure 3.3.2 a)} plots success-rate and time taken against space used when $\gamma=.1$ and $\delta$ is varied.
-The time taken is constant throughout, but the success rate plateaus at $\smallsim\!90\%$ when $d\geq.2$.
+The time taken is constant throughout, but the success rate plateaus at $\smallsim\!90\%$ when $\delta\leq.2$.
 \texttt{Figure 3.3.2 b)} shows that fixing $\delta=.2$ and varying $\gamma$ has no effect on performance.
@@
 \par The vertex sample $A$ can be acquired by uniformly sampling from $[n]$ until a sufficient number of unique values is sampled.
-For a sample of size $m$, each time a vertex is sampled the probability of a collision is at most $\frac{m}n$ meaning the expected number of total collisions is at most $m$.
-Thus, in the worst case a sample is generated in $\mathcal{O}(mn)$ time.
+For a sample of size $m$, each time a vertex is sampled the probability of a collision is at most $\frac{m}n$ meaning the expected number of total collisions is at most $\frac{m^2}{n}$.
+Thus, a sample is generated in $\mathcal{O}(m^2/n)$ time.
 
 \par The sampled edges form an insertion-only graph stream so a sufficient neighbourhood can be found among them using the algorithms discussed in \textbf{Chapter 2}.
-Due to the limitations from the $L_0$ sampler implementation I could not run this implementation on particular large graphs, this meant that the number of sampled edges is very small and thus the naive \textbf{Algorithm 4} provided the best results as it always succeeds.
+Due to the limitations from the \texttt{$L_0$ Sampler} implementation I could not run this implementation on particular large graphs, this meant that the number of sampled edges is very small and thus the naive \textbf{Algorithm 4} provided the best results as it always succeeds.
 If the implementation was improved to run on larger graphs then \textbf{Algorithm 3} should be used for its overall performance improvements.
@@
 These failures do not have a significant effect the performance of the algorithm as checking for sufficient neighbourhoods accounts for a small amount of the time and space usage.
-As expected, the hashing functions used in $L_0$ sampling account for the majority of space and time usage.
+As expected, the \texttt{hash arrays} used in \texttt{$L_0$ Sampling} account for the majority of space and time usage.
 Reducing the number of samplers will reduce the time and space usage of the implementation but will increase the failure rate if the number of unique edges sampled is reduced.
@@
 
-\vspace{.3cm}\begin{adjustwidth}{.3cm}{}\fbox{\parbox{\textwidth}{
-
+\subsubsection{Number of $L_0$ Samplers}
+
+\fbox{\parbox{\textwidth}{
 \textbf{Lemma 3.3} $\frac{\ln(1-p)}{\ln\left(1-\frac1m\right)}$ $L_0$ samplers are expected to find $p\delta$ unique neighbours of a vertex of degree $\delta$.\\
@@
 
-}}\end{adjustwidth}\vspace{.3cm}
-
-\par \textbf{Lemma 3.3} tells us the minimum number of $L_0$ samplers required for us to expect a sufficient neighbourhood to a vertex to be found and by tuning the value of $p$ the number of samplers and achieved success rate can be optimised.
+}}
+
+\par \textbf{Lemma 3.3} states the minimum number of \texttt{$L_0$ Samplers} required for us to expect a sufficient neighbourhood to a vertex to be found and, by tuning the value of $p$, the number of samplers and achieved success rate can be optimised.
 When a vertex is sampled it is appropriate to assume it has at least degree $\frac{d}c$.
@@
 Testing showed that setting $p=.9$ resulted in a sufficient neighbourhood being found $23\%$ of the time for a vertex of degree $\frac{11d}{10c}$ and $97\%$ of the time for a vertex of degree $\frac{12d}{10c}$.
-The implementation of $L_0$ samplers being used only successfully samples $90\%$ of the time, to account for this the expected number of samplers is multiplied by $\frac{1}{.9}$.
+The implementation of \texttt{$L_0$ Samplers} being used only successfully samples $90\%$ of the time, to account for this the expected number of samplers is multiplied by $\frac{1}{.9}$.
 
@@
 \end{figure}
-\texttt{Figure 3.3.4} shows the results when $\frac1{.9}\cdot\frac{\ln(1-.9)}{\ln(1-\frac{c}d)}$ samplers are run for each sampled vertex, rather than the suggested $10\frac{d}{c}\ln(n)$.
-The total space usage more than halved in all cases, and the space not used by the hash maps reduced by over $80\%$.
-The time taken by the algorithm (excluding that spent producing the hash maps) decreased by over $40\%$ in all cases.
+\texttt{Figure 3.3.4} shows the results when $\frac1{.9}\cdot\frac{\ln(1-.9)}{\ln(1-\frac{c}d)}$ \texttt{$L_0$ Samplers} are run for each sampled vertex, rather than the suggested $10\frac{d}{c}\ln(n)$.
+The total space usage more than halved in all cases, and the space not used by the \texttt{hash arrays} reduced by over $80\%$.
+The time taken by the algorithm (excluding that spent producing the \texttt{hash arrays}) decreased by over $40\%$ in all cases.
 Both of these reductions are due to the total number of $L_0$ samplers being reduced.
@@
 
+\subsubsection{Vertex Sample Size}
 \par The optimal vertex sample size depends on the degree distribution of the graph as this describes how many vertices are of sufficient degree.
@@
 \texttt{Figure 3.3.5} shows the performance results when a vertex sample size of $\frac{12n}{10c}$ was taken compared to the initial $\sqrt{n}$.
-As predicted, the algorithm succeed in all cases where the new sample size was used however it's performance is significantly worse in all cases where $\frac{12n}{10c}\geq n$.
-The small size \texttt{facebook\_deletion} means that Zipf's Law is unlikely to fit well and it is theorised that the degree distribution of social networks have a fatter tail than a standard Power-law distribution.
+As predicted, the algorithm succeed in all cases where the new sample size was used however it's performance is significantly worse in all cases where $\frac{12n}{10c}>n$.
+The small size of \texttt{facebook\_deletion} means that Zipf's Law is unlikely to fit well and it is theorised that the degree distributions of social networks have a fatter tail than a standard Power-law distribution.
 It is clear that further research is required into a performance optimal sample size for different scenarios.
@@
 
-Running an $L_0$ sampler on the set of edges means the domain of the $k$-wise independent hash function is the edge set and thus we require a way of indexing the edges such that each has a unique value which can then be hashed.
+Running an \texttt{$L_0$ Sampler} on the set of edges means the domain of the $k$-wise independent hash function is the edge set and thus we require a way of indexing the edges such that each has a unique value which can then be hashed.
 This indexing needs to be bijective so that the recovered index can be converted back into an edge.
 My initial implementation assigned the edge $(u,v)$ to $i=(\mathtt{min}(u,v)-1)\cdot n+\mathtt{max}(u,v)-1$ which maps edges to values in $[0,n^2)$.
-The use of \texttt{max} and min means that $(u,v)$ and $(v,u)$ map to the same value which is desirable since the edges are undirected.
+The use of \texttt{max} and \texttt{min} means that $(u,v)$ and $(v,u)$ map to the same value which is desirable since the edges are undirected.
 An edge can be recovered from this mapping as $\mathtt{min}(u,v)=(i\text{ mod }n)+1$ and $\mathtt{max}(u,v)=\mathtt{floor}(1/n)+1$.
-This mapping means the hash maps used in $L_0$ sampling now have $\mathcal{O}(n^2)$ space and time complexity.
+This mapping means the \texttt{hash arrays} used in \texttt{$L_0$ Sampling} now have $\mathcal{O}(n^2)$ space and time complexity.
 %TODO extracting edge from the key
@@
 \texttt{Figure 3.3.6} shows the results from the initial implementation of \textbf{Algorithm 9}.
-As with the implementation of the vertex sampling algorithm, the majority of the space and time requirements are spent in generating the hash arrays for the $L_0$ samplers.
+As with the implementation of the vertex sampling algorithm, the majority of the space and time requirements are spent in generating the \texttt{hash arrays}.
 The requirement for generating these is so high that I could not run the algorithm on \texttt{facebook\_deletion} as the space requirements were over $1$ GB and it took several minutes to generate a single hash array.
@@
 \par To reduce the size of the hash maps I changed how the edges were indexed.
-The graphs used in this project are undirected so have at most $\frac12n(n-1)$ unique edges.
+The graphs used in this project are undirected so have at most ${n\choose2}$ unique edges.
 These can be visualised as the upper triangle of an adjacency matrix.
-The following formula map between an edge $(u,v)$ to key $k\in[\frac{1}{2}n(n-1)]$.
+The following formula map between an edge $(u,v)$ to key $k\in\left[{n\choose2}\right]$.
 In the implementation I ensured that $v>u$ so that the edges $(u,v)$ and $(v,u)$ mapped to the same $k$ and that the equations would succeed.
@@
 \end{array}\]
-Implementing this method of indexing edges optimises the hash arrays but they still impact performance too much for the implementation to be usable on anything except the smallest of graphs.\texttt{Figure 3.3.7} shows the performance improvements from this implementation.
-
-\par This implementation succeeded in every case due to large number of samplers successfully sampling every edge, this is inefficient as most of these edges went unused while checking for a sufficient neighbourhood.
-For larger graphs not all edges would be sampled as the relative difference between the number of edges and number of samplers decreases.
+Implementing this method of indexing edges optimises the \texttt{hash arrays} but they still impact performance too much for the implementation to be usable on anything except the smallest of graphs.
+\texttt{Figure 3.3.7} shows the performance improvements from this implementation.
+
+\par This implementation succeeded in every case due to large number of samplers managing to successfully sample every edge, this is inefficient as most of these edges went unused while checking for a sufficient neighbourhood.
+For larger graphs not all edges would be sampled as the relative difference between the number of edges and number of \texttt{$L_0$ Samplers} increases.
 However, this implementation is not usable on reasonably sized graphs so I cannot effectively test alterations to the number of samplers run.
-The edge sampling method should be designed to work in cases where the vertex sampling method fails.
+
+\par The edge sampling method should be designed to work in cases where the vertex sampling method fails.
 Primarily this is when it is unlikely that a sufficient vertex is sampled, this occurs in graphs where only a handful of vertices are of high degree and the rest are of very low degree.
-In these cases there are very few edges which do not involve a vertex of high degree and by estimating how many high degree vertices there are you can calculate the expected number of edges which need to be sampled before a sufficient neighbourhood is found.
+In these cases there are very few edges which are not incident to a vertex of high degree and by estimating how many high degree vertices there are you can calculate the expected number of edges which need to be sampled before a sufficient neighbourhood is found.
 
@@
 
-The performance of the implementations of \textbf{Algorithm 8} and \textbf{Algorithm 9} are dominated by the hash functions used by the $L_0$ samplers.
+The performance of the implementations of \textbf{Algorithm 8} and \textbf{Algorithm 9} are dominated by the \texttt{hash arrays} implemented for \texttt{$L_0$ Sampling}.
 This meant that the implementations could not be run on graphs of sufficient size to evaluate certain aspects of the implementations, namely the vertex sample size in \textbf{Algorithm 8} and number of samples in \textbf{Algorithm 9}.
@@
 
-The algorithm proposed in \cite{orig} for insertion-only graphs proves very practical for solving \texttt{neighbourhood detection} on real world graphs.
-The changes to the implementations made reduced the time and space used by the algorithm significantly, while not compromising on the success rate.
+\textbf{Algorithm 2} proves very practical for solving \texttt{Neighbourhood Detection} on large real-world graphs.
+The changes to the implementations reduced the time and space used by the algorithm significantly, while not compromising on the success rate.
 The final evaluation shows that the proposed algorithm is much more space-efficient than a naïve approach.
-Even for close approximations of large streams the final implementation uses less than $6\%$ of the file size and for $c\geq5$ it uses $\leq1\%$.
+Even for close approximations of large graph streams the final implementation uses less than $6\%$ of the file size and for $c\geq5$ it uses less than $1\%$.
 This space efficiency increases the maximum file size at which the algorithm is usable, without requiring virtual memory.
@@
 However, I still managed to discuss several improvements to the implementation.
-The simplification of \texttt{1-sparse recovery} to only require two counters is a great improvement as, when the hash function is discounted, the \texttt{1-sparse recoveries} account for the majority of the space used by the $L_0$ samplers and thus both proposed algorithms.
-The probabilistic approach taken to reduce the number of $L_0$ samplers run for each vertex significantly improved performance by reducing redundancy, without compromising success.
-The indexing method which mapped edges to unique values in $\left[0,\frac12n(n-1)\right)$ was a really good improvement as it is the optimal indexing method for undirected graphs.
+The simplification of \texttt{Exact 1-Sparse Recovery} to only require two counters is a great improvement as, when the \texttt{hash arrays} are discounted, the \texttt{Exact 1-sparse recoveries} account for the majority of the space used by the \texttt{$L_0$ Samplers} and thus both proposed algorithms.
+The probabilistic approach taken to reduce the number of \texttt{$L_0$ Samplers} run for each vertex significantly improved performance by reducing redundancy, without compromising success.
+The indexing method which mapped edges to unique values in $\left[0,{n\choose2}\right)$ is the optimal indexing method for undirected graphs.
 
@@
 \subsection*{Complete Real World Evaluation}
-In this project I soley used graphs from social networks as they were readily available and an area where \texttt{neighbourhood detection} is an interesting problem.
+In this project I soley used graphs from social networks as they were readily available and an area where \texttt{Neighbourhood Detection} is an interesting problem.
 It would be interesting to test the implementations produced in this project on data sets from other fields.
@@
 \subsection*{$k$-Wise Independent Hashing Family}
-An alternative family of $k$-wise independent hash function is ${\displaystyle h(x)=\left(\left(\sum_{i=0}^{k-1}a_ix^i\right)\text{mod }P\right)\text{mod }m}$ where $a_0>0$ and $a_i<P\ \forall\ i$.
+An alternative family of $k$-wise independent hash function is
+$${\displaystyle h(x)=\left(\left(\sum_{i=0}^{k-1}a_ix^i\right)\text{mod }P\right)\text{mod }m}$$
+where $a_0>0$ and $a_i<P\ \forall\ i$.
 A function from this family is quick to generate as only the $k$ $a_i$ values need to be random generated and can be stored as $k+2$ integers (the values of $a_i$, $k$ and $m$).
 Both of these are dramatic improvements from the implementation of hash arrays used in the implementation of an $L_0$ sampler.
-The difficulty with implementation functions from this family is that the values of $x^i$ quickly overflow causing inaccurate calculations.
+The difficulty with implementating functions from this family is that the values of $x^i$ quickly overflow causing inaccurate calculations.
 I am sure an implementation of this family of hash functions could be worked out, but I ran out of time to do so in this project.
@@
 I discussed how the size of the vertex sample used in \textbf{Algorithm 8} should depend on the degree distribution of the graph, and investigated one line of thought for how this could be done for graphs of social network connections.
-Investigating this further was outside the scope of this project but would be an interesting feature to research further.
+Investigating this further was outside the scope of this project but would be an interesting area to research further.
 \par The edge sampling approach proposed in \textbf{Algorithm 9} should be considered whilst evaluating the size of the vertex sample, as there will be cases where this approach is more efficient.
 The most obvious example of this would be a star graph where the vertex sampling approach would need to sample every vertex in order to guarantee find a sufficient neighbourhood, whereas the edge sampling approach only needs to sample $\frac{d}c$ edges.
-In more general scenarios it is likely that the edge sampling approach is better for close approximations and the vertex sampling approach better for loser approximations.
+In more general scenarios it is likely that the edge sampling approach is better for tight approximations $c\leq5$ and the vertex sampling approach better for loser approximations.
 
 \subsection*{Reservoir Sample Size}
-I tested changing the reservoir sample size used in \textbf{Algorithm 2} but no strategies where found which consistently improved performance.
+I tested changing the reservoir sample size used in \textbf{Algorithm 2} but no strategies which consistently improved performance presented themselves.
 These tests changed the sample size of every sampler used, but it may be that reducing the sample size for later samplers produces good results as those samplers are sampling for a smaller pool.
-This would be the main feature to look at when further trying to optimised the implementation from \textbf{Chapter 2}.
+This would be the main area to look at when further trying to optimised the implementation from \textbf{Chapter 2}.
 
@@
 	\addcontentsline{toc}{chapter}{Bibliography}
-	\bibitem{orig} Christian Konrad \textit{Streaming Frequent Items with Timestamps and Detecting Large Neighborhoods in Graph Streams.} November 2019.
   \bibitem{SNAP}http://snap.stanford.edu/data/
+  \bibitem{spanner} Surender Baswana \textit{Streaming Algorithm for Graph Spanners - Single Pass and Constant Processing Time per Edge.} April 2008.
+  \bibitem{edgesOfSpanner} B\'ela Bollob\'as \textit{Extremal Graph Theory.} 1978
   \bibitem{L0Framework} Graham Cormode, Donatella Firmani \textit{A Unifying Frameworkd for $\ell_0$-Sampling Algorithms.} September 2014.
-  \bibitem{socialNetworkDistribution} Tushti Rastogi \textit{A power law approach to estimating fake social network accounts.} May 2016.
-  \bibitem{GraphStreamSurvey} Andrew McGregor \textit{Graph Stream Algorithms: A Survey.} May 2014.
+  \bibitem{LpSurvey} Graham Cormode, Hossein Jowhari \textit{$L_p$ Samplers and Their Applications: A Survey.} November 2018.
   \bibitem{clusteringCofficient} Madhav Jha, C.
 Seshadhri, and Ali Pinar \textit{A space-efficient streaming algorithm for estimating transitivity and triangle counts using the birthday paradox.} February 2015.
-  \bibitem{edgesOfSpanner} B\'ela Bollob\'as \textit{Extremal Graph Theory.} 1978
-  \bibitem{spanner} Surender Baswana \textit{Streaming Algorithm for Graph Spanners - Single Pass and Constant Processing Time per Edge.} April 2008.
+	\bibitem{orig} Christian Konrad \textit{Streaming Frequent Items with Timestamps and Detecting Large Neighborhoods in Graph Streams.} November 2019.
+  \bibitem{GraphStreamSurvey} Andrew McGregor \textit{Graph Stream Algorithms: A Survey.} May 2014.
+  \bibitem{graphStreams}Alon Noga, Mario Szegedy, Matias Yossi \textit{The space complexity of approximating the frequency moments.} July 1996.
+  \bibitem{socialNetworkDistribution} Tushti Rastogi \textit{A power law approach to estimating fake social network accounts.} May 2016.
   \bibitem{RandomWalks} Atish Das Sarma, Sreenivas Gollapudi, and Rina Panigraphy \textit{Estimating PageRank on Graph Streams.} May 2011.

### 2020-05-13 src/insertionDeletionStreams/insertionDeletionEdgeSampling.cpp

@@ -87,10 +87,11 @@ int main() {
   //execute_test(2,20,1,reps,d,num_vertices,edge_file_path,vertex_file_path,out_file);
 
   // details of graph to perform on
-  //edge_file_path="../../data/facebook_deletion.edges"; vertex_file_path="../../data/facebook_deletion.vertices"; num_vertices=747; d=267; reps=10;
-  edge_file_path="../../data/facebook_small_deletion.edges"; vertex_file_path="../../data/facebook_deletion.vertices"; num_vertices=52; d=33; reps=10;
-  out_file="edge_sampled_better_id.csv";
-  execute_test(2,d,1,reps,d,num_vertices,edge_file_path,vertex_file_path,out_file);
+  //edge_file_path="../../data/facebook_deletion.edges"; vertex_file_path="../../data/facebook_deletion.vertices"; num_vertices=747; d=267; reps=1;
+  edge_file_path="../../data/facebook_small_deletion.edges"; vertex_file_path="../../data/facebook_deletion.vertices"; num_vertices=52; d=16; reps=10;
+  //out_file="facebook_deletion_edge_sampled_better_id.csv";
+  out_file="edge_sampled_better_id_2.csv";
+  execute_test(2,2,1,reps,d,num_vertices,edge_file_path,vertex_file_path,out_file);
 
   //set<vertex> neighbourhood; vertex root; // variables for returned values
   //int c=2;
@@ -185,7 +186,7 @@ void single_pass_insertion_deletion_stream(int c, int d, int num_vertices, strin
   uint64_t n3=pow(possible_edges,3);
   cout<<"possible_edges="<<possible_edges<<", n3="<<n3<<endl;
   for (int i=0; i<total_samplers; i++) {
-    if (i%10==0) cout<<"\r"<<i;
+    if (i%1==0) cout<<"\r"<<i;
     uint64_t* hash_map=generate_random_hash(possible_edges,n3);
     unique_hash_maps[i]=hash_map;
   }

### 2020-05-13 src/insertionDeletionStreams/insertionDeletionStreamsVertexSampling.cpp

@@ -94,10 +94,10 @@ int main() {
   //execute_test(2,20,1,reps,d,num_vertices,edge_file_path,vertex_file_path,out_file);
 
   // details of graph to perform on
-  edge_file_path="../../data/facebook_deletion.edges"; vertex_file_path="../../data/facebook_deletion.vertices";
-  out_file="no_tau_test.csv";
-  num_vertices=747; d=267; reps=200;
-  execute_test(10,100,1,reps,d,num_vertices,edge_file_path,vertex_file_path,out_file);
+  //edge_file_path="../../data/facebook_deletion.edges"; vertex_file_path="../../data/facebook_deletion.vertices";num_vertices=747; d=267; reps=2;
+  edge_file_path="../../data/gplus.edges"; vertex_file_path="../../data/gplus.vertices";num_vertices=12417; d=5948; reps=10;
+  out_file="gplus_insertion_with_idv_algorithm.csv";
+  execute_test(5,20,1,reps,d,num_vertices,edge_file_path,vertex_file_path,out_file);
 
   /*set<vertex> neighbourhood; vertex root; // variables for returned values
   int c=10;
@@ -110,13 +110,13 @@ int main() {
 
 void execute_test(int c_min, int c_max, int c_step, int reps, int d, int n, string edge_file_path, string vertex_file_path, string out_file) {
   ofstream outfile(out_file);
-  outfile<<"name,"<<vertex_file_path<<endl<<"n,"<<n<<endl<<"d,"<<d<<endl<<"repetitions,"<<reps<<endl<<"delta,0.2"<<endl<<"gamma,0.3"<<endl<<"vertex sample size,log(n)"<<endl<<"L0 per vertex,ceil((1/success_rate)*log(1-.9)/log(1-((c-1)/(double)d)))"<<endl; // test details
-  outfile<<"c,time (microseconds), generating l0 hash time (microseconds) ,mean max space (bytes), l0 hash space (bytes), variance time, variance hash time, variance max space, variance hash space,successes"<<endl; // headers
[…]

### 2020-05-13 src/insertionDeletionStreams/l0Samplers/sSparseRecovery.cpp

@@ -69,7 +69,7 @@ int main() {
 
   // details of graph to perform on
   string file_path="../../../data/facebook_deletion.edges";
-  vertex target=2290;
+  vertex target=0;
   int num_vertices=747;
 
   set<vertex> recovered_neighbourhood=s_sparse_recovery(file_path, target, num_vertices, s, delta);
